---
title: "Binary 2D Segmentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Binary 2D Segmentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This experimental contract remains unpromoted (`vetted=FALSE`) pending reviewer
acceptance of the executed evidence. The `pytorch_resnet18_segmentation` contract
fits a binary mask on the canonical
128 by 128 grid. Each custodian must select the **patient privacy unit** and
configure its image and mask roots. The analyst supplies schema column names
and asset aliases, never filesystem roots or privacy parameters.

The current evidence is **protocol v5: public BUSI decoder pretraining followed
by federated DP fine-tuning**. Random-initialization v3 and v4 results remain
separately labelled boundary evidence; v2 remains invalidated by its schedule
defect. The September 20 reconciliation established that the previously
packaged current files were stale v3 copies. The current files now contain the
verified v5 results. Mechanism checks remain separately recorded; execution
and a passed utility floor do not constitute reviewer acceptance.

## Paired image and mask schema

A samples table declares a stable subject identifier, `image_id`, an image path
(default `relative_path`), and the mask path column passed as `target`. PNG/JPEG
images and single-channel PNG masks are supported. Declare `mask_values` as
`"0,255"` or `"0,1"`; masks are never intensity normalized. An optional
`mask_empty_col` explicitly declares empty annotations. Without this declaration,
a missing annotation is invalid, rather than a background example.

For each subject, the node selects the lexicographically first canonical image
ID (trimmed Unicode NFC). Masks for that image are unioned. Conflicting image
paths for the same selected ID invalidate the subject instead of introducing an
order-dependent tie. Invalid paths, images, masks or pairings become fixed safe
tensors with zero loss validity. Those subjects remain in the sampling census;
no completeness counts are released.

Direct samples tables use custodian options `dsflower.image_data_root` and
`dsflower.mask_data_root`. Imaging resources use admitted dsImaging image and
`mask_root` descriptors. Both routes enforce the same asset containment checks.

## Submit the fixed spatial contract

```r
model <- ds.flower.model.pytorch_resnet18_segmentation(
  sample_id_col = "image_id",
  image_path_col = "relative_path",
  mask_empty_col = "mask_empty",
  mask_values = "0,255",
  batch_size = 8L,
  local_epochs = 1L,
  learning_rate = 0.001
)

fit <- ds.flower.fit(
  conns, symbol = "samples", target = "mask_path",
  model = model, task = "segmentation", data_kind = "image",
  rounds = 5L, output_dir = "segmentation-model"
)
```

`ds.flower.model("pytorch_resnet18_segmentation")`, `ds.flower.recipe()` and
`ds.flower.submit()` use the same contract. The target is a **mask path role**,
so do not supply class levels, target bounds or tabular feature columns.

The resident encoder is ResNet18 through layer2, pinned to
`ResNet18_Weights.IMAGENET1K_V1` and profile `resnet18_layer2_128_v1`. The full
checkpoint SHA256 is
`f37072fd47e89c5e827621c5baffa7500819f7896bbacec160b1a16c560e07ec`.
Custodians provision these exact weights before training. Missing or mismatched
weights fail preflight. The encoder is frozen, in evaluation mode, outside the
released module; its BatchNorm statistics never learn from private images.

The default decoder has three convolutions (128 to 32, 32 to 16, 16 to 1),
ReLU activations and nearest upsampling by 2 then 4, producing `[1,128,128]`.
The loss mixes per-image mean BCE and smoothed Dice with `alpha=0.5` and
smoothing 1. `alpha=1` is the declared BCE ablation. There is no batch-wide Dice,
trainable encoder, mixed precision, or image augmentation in this first profile.

The constructor defaults to `decoder_init="random"` and the 41,537-parameter
`decoder="current"`. It also supports `decoder="narrow"` (9,521 parameters;
convolution channels 128 to 8, 8 to 4, 4 to 1) and `decoder="pointwise"` (129
parameters). Selecting the narrow architecture alone retains random
initialization.

## Public decoder initialisation

The two 0.7.0 routes share a complete digest-bound bundle including the encoder,
decoder and evidence. Use researcher-local material for an analyst declaration:

```r
model <- ds.flower.model.pytorch_resnet18_segmentation(
  decoder = "narrow", decoder_init = "client:/analyst/public/busi_bundle.zip"
)
fit <- ds.flower.fit(
  conns, symbol = "samples", target = "mask_path", model = model,
  task = "segmentation", data_kind = "image", rounds = 2L,
  output_dir = "segmentation-public-model"
)
```

For custodian-registered material, first assign and admit the authorized resource:

```r
DSI::datashield.assign.resource(
  conns, symbol = "CKPT_R", resource = "PublicModels.busi"
)
DSI::datashield.assign.expr(
  conns, symbol = "CKPT", expr = quote(flowerCheckpointInitDS("CKPT_R"))
)
model <- ds.flower.model.pytorch_resnet18_segmentation(
  decoder = "narrow", decoder_init = "resource:CKPT"
)
fit <- ds.flower.fit(
  conns, symbol = "samples", target = "mask_path", model = model,
  public_checkpoint_file = "/analyst/public/checkpoint.npz",
  task = "segmentation", data_kind = "image", rounds = 2L,
  output_dir = "segmentation-resource-model"
)
```

The local coordinator copy must match every node's admitted identity; it grants
no resource authority. Nodes never return checkpoint bytes. A typed handle binds
the snapshot to its session; raw URLs, paths, lists and unassigned symbols are
refused. The custodian option `dsflower.public_initialisation` accepts
`analyst_or_resource` (default), `resource_only`, or `none`, with an optional
`.pytorch_resnet18_segmentation` suffix. Both admission and execution fail closed
on missing, altered or incompatible artifacts. The first round compares received
public tensors with the admitted checkpoint; later rounds use federated updates.

The node package's `PUBLIC_INITIALISATION.md` documents exact Opal, Armadillo and
DSLite registration commands. Its `inst/extdata/segmentation-public-checkpoints/`
contains the unchanged BUSI reference evidence and recovery instructions; original
NPZ binaries remain unavailable. No 0.6.0 registry installation authorizes the new
routes. Request-selection v2 retains its mechanism and binds canonical content
identity and origin, excluding resource aliases and archive packaging. The
accountant, clipping, sampler and private training path are unchanged.

The historical v5 benchmark applied the same public BUSI checkpoints through
campaign tooling. This registry route makes custodian enablement independent
of that tooling; selecting an initialization alone does not reproduce the
benchmark's full schedule or establish a new utility claim.

## Local prediction and public evaluation

```r
probabilities <- ds.flower.predict(
  "segmentation-model", c("public/image-1.png", "public/image-2.png"),
  type = "prob"
)
masks <- ds.flower.predict(
  "segmentation-model", c("public/image-1.png", "public/image-2.png"),
  type = "response"
)
# Both arrays use [image, channel, row, column] = [2, 1, 128, 128].
# Responses threshold probabilities at >= 0.5.
```

Prediction verifies the decoder artifact and the frozen encoder checkpoint
before opening image paths. It shares the node's stored-pixel orientation, RGB
conversion, bilinear 128 by 128 resize and ImageNet channel normalization.
Invalid local images receive fixed zero features. Outputs use the canonical
grid. To map a binary prediction back to the stored native image of height `H`
and width `W`, threshold at 0.5 on the canonical grid first, then use nearest
resampling with the following zero-based coordinate rule (exact ties choose
the larger canonical index):

```text
native[r, c] = canonical[min(127, floor((r + 0.5) * 128 / H)),
                         min(127, floor((c + 0.5) * 128 / W))]
```

This retains stored-pixel orientation, with no EXIF rotation, cropping or
padding. It restores the coordinate grid, not detail discarded by resizing.
For known dimensions from an authorized local image, the same rule in R is:

```r
canonical <- masks[1L, 1L, , ]
native_rows <- pmin(128L, floor((2 * seq_len(native_height) - 1) *
                                128 / (2 * native_height)) + 1L)
native_cols <- pmin(128L, floor((2 * seq_len(native_width) - 1) *
                                128 / (2 * native_width)) + 1L)
native_mask <- canonical[native_rows, native_cols, drop = FALSE]
```

Use these indices explicitly: resize libraries may differ at exact tie pixels
through floating-point coordinate accumulation. Declare and label any
native-resolution secondary metric with this mapping.
The primary benchmark metric remains on the canonical 128 by 128 grid.

Channel-B evaluation uses public or independently authorized analyst-local
held-out subjects. Report mean subject Dice and IoU at threshold 0.5, and report
foreground-positive and empty-mask strata separately. Both-empty Dice is 1;
exactly one empty mask gives Dice 0. Neither images nor predictions are returned
from a private node for evaluation.

Private `ds.flower.validate()`, holdout and cross-validation release pooled
foreground Dice from bounded per-patient intersection/prediction/reference
statistics. This differs from mean subject Dice on a public split. Both admitted
public-initialisation routes support holdout and fresh starts for each CV fold.
See [private validation and CV](../PRIVATE_VALIDATION_CV.md) for calls and sensitivities.
Private-score HPO remains unsupported; the local HPO callback does not authorize it. Any public
benchmark utility claim must cite an executed subject-disjoint protocol, exact
central twins and the recorded gate results; the constructor alone establishes
no utility claim.

## Public benchmark evidence

The current preregistered v5 campaign uses BrEaST and BUS-BRA with three DSLite sites,
subject-disjoint 80/20 splits, three seeds, epsilon 1, 4 and 8, delta `1e-5`,
clipping 1 and separate nominal batch-size 16 and 64 arms. Protocol v5 fixes 128 by 128
inputs and selects the narrow decoder, 60 public pretraining epochs, batch16
and 20 rounds from three-seed BUS-BRA inner validation. Batch64 is the
prespecified sensitivity arm. Fine-tuning uses three local epochs per round
(60 total), round-local Adam at learning rate 0.001, and no augmentation or
mixed precision. BUS-BRA also has the nested
192-subject, BCE-only and heterogeneous-site extensions. The released
federated DP model is compared with pooled DP, pooled nonprivate and
federated nonprivate twins using the same seed-specific public-pretrained decoder, effective tensors,
loss, optimizer and subject splits. Nonprivate twins are public benchmark
diagnostics outside the released contract.
The trivial comparison selects the strongest of four fixed masks: empty,
full, a centered disk of radius 32, and a centered 64 by 64 square.

Public pretraining used 780 BUSI images from the
[Arya Shah Kaggle mirror, version 1](https://www.kaggle.com/datasets/aryashah2k/breast-ultrasound-images-dataset),
attributed to [Al-Dhabyani et al. (2020)](https://doi.org/10.1016/j.dib.2019.104863).
It is a disclosed nonprivate stage outside fine-tuning's DP accounting. BUSI
has no patient mapping for this protocol and supplies no held-out utility
claim; within-BUSI duplicates were retained and cross-dataset near-duplicate
screening was not claimed. The mirror's CC0 declaration is preserved as
provenance. `provenance/busi-v1/provenance.json`,
`public-pretraining-manifest.json` and `public-pretraining/audit.json` bind the
dataset, preprocessing, encoder and seed-specific checkpoint digests.
The dataset ZIP SHA256 is
`7fffc86a517934da55021f66641d81383058d26a30adb0ffab3780ca7e52d57d`;
the pretraining audit SHA256 is
`d59334151e074de56afaaec81e92930ddaad8371f7f2665577096ce4ff7d9d57`.
Dataset images are not distributed with the package. The node's public
checkpoint bundle instructions describe the separately installed decoder tensors
and preserve the mirror's licence declaration and rights-attribution limitation.

Selection excluded outer-test subjects, but confirmation reused previously
inspected public outer splits. These results describe public benchmark
confirmation after development, not an unseen clinical test. Repeated
benchmark runs are separately recorded; their privacy budgets must not be
portrayed as one non-composed private release. The model remains `vetted=FALSE`.

The tables below read `campaign-status.json`, `breast-evidence.json` and
`busbra-evidence.json` in each arm directory under
`inst/extdata/campaign/segmentation/` at render time. Arms are reported separately;
batch16 is the selected primary arm and batch64 is sensitivity, with no
averaging across arms. Current `protocol.md` is byte-identical to
`protocol-v5.md`, SHA256
`28a713f8ffe6f7bb7b74e7d59d13d1be265d5738b43503c60e6581a67e0c5820`.
The frozen protocol, dataset release hashes, licence texts, subject-mapping
audits and split hashes are archived alongside the evidence. The
archival assembler validates the complete JSON schemas, captured accountant
steps, effective tensors, released artifacts and twin provenance. Rendering
checks the schema identity, contract and protocol hash again. An incomplete
cohort cannot produce a completed-matrix summary. Execution status is separate
from utility acceptance and reviewer promotion.

The earlier negative results remain in `retained/v3/` and `retained/v4/`, each
with its original protocol. V4 uses the same selected narrow decoder and
20-round schedule with random initialization; v3 uses the original decoder
and ten rounds. The following comparison retains those distinctions.
The reconciliation report and `audit-results.json` under
`reconciliation-v5-20260920/` document the stale-path correction. Independent
forward scoring reproduced all six v5 batch16 full-cohort epsilon8 results;
the captured initialization matched the public checkpoints. Batch64 was not
independently forward-scored in that audit.

```{r segmentation-evidence, eval=TRUE, echo=FALSE, error=FALSE, results="asis"}
evidence_dir <- file.path("..", "inst", "extdata", "campaign", "segmentation")
if (!dir.exists(evidence_dir)) {
  evidence_dir <- system.file("extdata", "campaign", "segmentation",
                              package = "dsFlowerClient")
}
archive_dir <- evidence_dir
boundary_rows <- list()
for (version in c("v3", "v4", "v5")) {
  version_dir <- if (version == "v5") archive_dir else
    file.path(archive_dir, "retained", version)
  evidence <- jsonlite::read_json(file.path(version_dir, "batch16", "busbra-evidence.json"))
  stopifnot(identical(evidence$protocol_sha256,
    digest::digest(file = file.path(version_dir, "protocol.md"), algo = "sha256")))
  for (epsilon in c(1, 4, 8)) {
    replicates <- Filter(function(value) value$epsilon == epsilon, evidence$replicates)
    stopifnot(length(replicates) == 3L,
      setequal(vapply(replicates, function(value) value$seed, integer(1)),
               c(20260919L, 20260920L, 20260921L)))
    boundary_rows[[length(boundary_rows) + 1L]] <- data.frame(
      Protocol = version,
      Initialization = if (version == "v5") "Public BUSI" else "Random",
      Epsilon = epsilon,
      `Foreground Dice` = sprintf("%.3f", mean(vapply(replicates,
        function(value) value$federated_dp$foreground_positive$dice, numeric(1)))),
      check.names = FALSE)
  }
}
print(knitr::kable(do.call(rbind, boundary_rows), row.names = FALSE,
  caption = "BUS-BRA batch16 federated-DP foreground Dice: mean over three seeds, with earlier boundaries retained."))
cat("\n\n")
render_arm <- function(evidence_dir) {
  if (basename(evidence_dir) %in% c("batch16", "batch64")) {
    cat("## Nominal ", basename(evidence_dir), " arm\n\n", sep = "")
  }
  protocol_hash <- digest::digest(file = file.path(archive_dir, "protocol-v5.md"),
                                  algo = "sha256")
  read_evidence <- function(filename, schema_file) {
    value <- jsonlite::read_json(file.path(evidence_dir, filename))
    schema <- jsonlite::read_json(file.path(archive_dir, schema_file))
    stopifnot(identical(value$schema, schema$`$id`),
              all(unlist(schema$required) %in% names(value)),
              identical(value$contract, "pytorch_resnet18_segmentation"),
              identical(value$task, "segmentation"),
              identical(value$protocol_sha256, protocol_hash),
              value$status %in% c("not_executed", "failed", "executed"))
    if (value$status != "executed") {
      stopifnot(is.character(value$reason), is.null(value$replicates))
    }
    value
  }
  status_schema <- jsonlite::read_json(file.path(evidence_dir, "campaign-status.json"))$schema
  status <- read_evidence("campaign-status.json",
    if (status_schema == "dsflower-segmentation-campaign-summary-v1")
      "campaign-summary-schema.json" else "evidence-schema.json")
  cat("**Campaign status: ", status$status, ".** ", status$reason, "\n\n", sep = "")
  cat("Protocol SHA256: `", protocol_hash, "`.\n\n", sep = "")
  cohort_files <- intersect(c("breast-evidence.json", "busbra-evidence.json"),
                            list.files(evidence_dir))
  cohorts <- lapply(cohort_files, read_evidence, schema_file = "evidence-schema.json")
  names(cohorts) <- sub("-evidence.json", "", cohort_files, fixed = TRUE)
  labels <- c(breast = "BrEaST", busbra = "BUS-BRA")
  variants <- c(full = "Primary", small192 = "192 subjects", bce = "BCE only",
                heterogeneous = "Heterogeneous sites")
  show_table <- function(rows, caption) {
    print(knitr::kable(do.call(rbind, rows), row.names = FALSE, caption = caption))
    cat("\n\n")
  }
  format_metric <- function(value) {
    if (is.null(value)) return("Not applicable")
    stopifnot(value$n_replicates >= 3L, length(value$ci95) == 2L,
              all(is.finite(c(value$mean, value$sd, unlist(value$ci95)))))
    sprintf("%.3f [%.3f, %.3f]", value$mean, value$ci95[[1L]], value$ci95[[2L]])
  }
  if (!length(cohorts)) {
    cat("No completed cohort evidence is archived. Utility floors and envelopes remain **not evaluated**.\n\n")
  }
  for (dataset in names(cohorts)) {
    cohort <- cohorts[[dataset]]
    cat("### ", labels[[dataset]], "\n\n**Cohort status: ", cohort$status, ".**\n\n", sep = "")
    failures <- Filter(function(cell) identical(cell$status, "failed"), cohort$cells)
    if (length(failures)) {
      show_table(lapply(failures, function(cell) data.frame(
        Setting = variants[[cell$variant]], Epsilon = cell$epsilon,
        Seed = cell$seed, Failure = cell$reason)), "Failed cells retained in the archive.")
    }
    if (cohort$status != "executed") {
      cat(cohort$reason, ". Utility floors and envelopes remain **not evaluated**.\n\n", sep = "")
      observed <- list()
      scalar_metric <- function(value) if (is.null(value)) "Not applicable" else sprintf("%.3f", value)
      branches <- c(federated_dp = "Federated DP", pooled_dp = "Pooled DP",
                    pooled_nonprivate = "Pooled nonprivate", federated_nonprivate = "Federated nonprivate")
      for (replicate in cohort$completed_cells) {
        for (branch in names(branches)) {
          values <- replicate[[branch]]
          observed[[length(observed) + 1L]] <- data.frame(
            Epsilon = replicate$epsilon, Seed = replicate$seed, Branch = branches[[branch]],
            Dice = scalar_metric(values$all$dice), IoU = scalar_metric(values$all$iou),
            `Foreground Dice` = scalar_metric(values$foreground_positive$dice),
            `Foreground IoU` = scalar_metric(values$foreground_positive$iou),
            `Empty Dice` = scalar_metric(values$empty_reference$dice),
            `Empty IoU` = scalar_metric(values$empty_reference$iou), check.names = FALSE)
        }
      }
      if (length(observed)) show_table(observed,
        "Observed successful cells from an incomplete matrix; these are not complete three-seed summaries or utility acceptance.")
      next
    }
    stopifnot(length(cohort$replicates) == 9L,
              !is.null(cohort$envelopes), !is.null(cohort$provenance),
              !is.null(cohort$runtime), identical(cohort$dataset, dataset))
    rows <- strata <- list()
    groups <- c(list(full = cohort$summaries), cohort$additional_summaries)
    for (variant in names(groups)) {
      for (epsilon in names(groups[[variant]])) {
        metrics <- groups[[variant]][[epsilon]]
        label <- paste(variants[[variant]], "epsilon", epsilon)
        replicates <- Filter(function(value) value$variant == variant &&
          value$epsilon == as.numeric(epsilon),
          c(cohort$replicates, cohort$additional_replicates))
        stopifnot(length(replicates) == 3L,
          setequal(vapply(replicates, function(value) value$seed, integer(1)),
                   c(20260919L, 20260920L, 20260921L)))
        rows[[length(rows) + 1L]] <- data.frame(
          Setting = label,
          `Federated DP Dice` = format_metric(metrics$federated_dp$all$dice),
          `Pooled DP Dice` = format_metric(metrics$pooled_dp$all$dice),
          `Pooled nonprivate Dice` = format_metric(metrics$pooled_nonprivate$all$dice),
          `Federated nonprivate Dice` = format_metric(metrics$federated_nonprivate$all$dice),
          `Strongest trivial Dice` = sprintf("%.3f", mean(vapply(replicates,
            function(value) value$trivial$strongest_dice, numeric(1)))),
          check.names = FALSE)
        for (stratum in c("all", "foreground_positive", "empty_reference")) {
          values <- metrics$federated_dp[[stratum]]
          strata[[length(strata) + 1L]] <- data.frame(
            Setting = label, Stratum = gsub("_", " ", stratum),
            Dice = format_metric(values$dice), IoU = format_metric(values$iou))
        }
      }
    }
    show_table(rows, "Mean subject Dice with 95% t intervals across three seeds.")
    show_table(strata, "Released-model Dice and IoU at threshold 0.5; absent strata are not applicable.")
    checks <- cohort$envelopes
    floor <- checks$utility_floor
    cat("**Epsilon 8 utility floor: ", if (isTRUE(floor$pass)) "PASS" else "FAILED", ".** ",
        if (dataset == "busbra") "This is the preregistered primary floor. " else
          "This is a descriptive application of the BUS-BRA floor. ",
        sprintf("Mean Dice %.3f; strongest trivial mask %.3f; margin %.3f. ",
                floor$federated_mean_dice, floor$strongest_trivial_mean_dice,
                floor$federated_mean_dice - floor$strongest_trivial_mean_dice),
        "Required: Dice >= 0.50 and margin >= 0.10.\n\n", sep = "")
    envelope_rows <- lapply(checks$epsilon_envelope, function(check) data.frame(
      Check = paste("Epsilon", check$previous, "to", check[["next"]]),
      Result = if (isTRUE(check$flag)) "FLAGGED" else "Not flagged",
      Detail = sprintf("Shortfall increase %.4f; seed-SD tolerance %.4f",
                       check$gap_increase, check$tolerance)))
    trend <- checks$small_n_noise_trend
    envelope_rows[[length(envelope_rows) + 1L]] <- data.frame(
      Check = "Small-cohort noise trend",
      Result = if (trend$status != "executed") "Not evaluated" else
        if (isTRUE(trend$flag)) "FLAGGED" else "Not flagged",
      Detail = if (trend$status != "executed") "Complete matched small-cohort matrix unavailable" else
        sprintf("Shortfall SD: epsilon 1 %.4f; epsilon 8 %.4f", trend$sd_epsilon1, trend$sd_epsilon8))
    near_groups <- c(list(full = checks$near_central), cohort$additional_near_central)
    if (!is.null(trend$near_central)) {
      if (!is.null(near_groups$small192)) {
        stopifnot(identical(near_groups$small192, trend$near_central))
      }
      near_groups$small192 <- trend$near_central
    }
    near_variants <- if (dataset == "busbra") names(variants) else "full"
    for (variant in near_variants) {
      near <- Filter(function(check) isTRUE(check$flag), near_groups[[variant]])
      expected_near <- if (variant %in% c("full", "small192")) 9L else 3L
      observed_near <- length(near_groups[[variant]])
      incomplete_near <- observed_near < expected_near
      envelope_rows[[length(envelope_rows) + 1L]] <- data.frame(
        Check = paste("Unexpected near-central agreement:", variants[[variant]]),
        Result = if (is.null(near_groups[[variant]])) "Not evaluated" else
          if (incomplete_near) "Incomplete" else
            if (length(near)) "FLAGGED" else "Not flagged",
        Detail = if (is.null(near_groups[[variant]])) "Complete matched setting unavailable" else
          paste0("Observed ", observed_near, " of ", expected_near, "; ",
            if (length(near)) paste(vapply(near, function(check)
              paste("FLAGGED epsilon", check$epsilon, "seed", check$seed), character(1)), collapse = "; ") else
              "no observed replicate met the investigation rule"))
    }
    show_table(envelope_rows, "Envelope checks; flags require investigation and are retained.")
  }
}
arm_dirs <- file.path(archive_dir, c("batch16", "batch64"))
arm_dirs <- arm_dirs[file.exists(file.path(arm_dirs, "campaign-status.json"))]
if (!length(arm_dirs)) arm_dirs <- archive_dir
invisible(lapply(arm_dirs, render_arm))

```

The shortfall used by the envelope is **pooled nonprivate Dice minus
federated DP Dice**. Positive values indicate lower utility for the released
model. Epsilon-envelope flags compare changes against seed dispersion;
near-central flags identify unexpectedly close agreement under the
preregistered small-cohort rule. An unflagged envelope does not establish
privacy by itself. Failed utility floors remain failed and do not trigger
changes to epsilon, delta, clipping, subject retention or the preregistered
protocol. Foreground-positive results remain visible even when empty-mask
performance changes the overall mean.
