---
title: "Binary 2D Segmentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Binary 2D Segmentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This experimental contract remains unpromoted (`vetted=FALSE`) pending reviewer
acceptance of the executed evidence. The `pytorch_resnet18_segmentation` contract
fits a binary mask on the canonical
128 by 128 grid. Each custodian must select the **patient privacy unit** and
configure its image and mask roots. The analyst supplies schema column names
and asset aliases, never filesystem roots or privacy parameters.

## Paired image and mask schema

A samples table declares a stable subject identifier, `image_id`, an image path
(default `relative_path`), and the mask path column passed as `target`. PNG/JPEG
images and single-channel PNG masks are supported. Declare `mask_values` as
`"0,255"` or `"0,1"`; masks are never intensity normalized. An optional
`mask_empty_col` explicitly declares empty annotations. Without this declaration,
a missing annotation is invalid, rather than a background example.

For each subject, the node selects the lexicographically first canonical image
ID (trimmed Unicode NFC). Masks for that image are unioned. Conflicting image
paths for the same selected ID invalidate the subject instead of introducing an
order-dependent tie. Invalid paths, images, masks or pairings become fixed safe
tensors with zero loss validity. Those subjects remain in the sampling census;
no completeness counts are released.

Direct samples tables use custodian options `dsflower.image_data_root` and
`dsflower.mask_data_root`. Imaging resources use admitted dsImaging image and
`mask_root` descriptors. Both routes enforce the same asset containment checks.

## Submit the fixed spatial contract

```r
model <- ds.flower.model.pytorch_resnet18_segmentation(
  sample_id_col = "image_id",
  image_path_col = "relative_path",
  mask_empty_col = "mask_empty",
  mask_values = "0,255",
  batch_size = 8L,
  local_epochs = 1L,
  learning_rate = 0.001
)

fit <- ds.flower.fit(
  conns, symbol = "samples", target = "mask_path",
  model = model, task = "segmentation", data_kind = "image",
  rounds = 5L, output_dir = "segmentation-model"
)
```

`ds.flower.model("pytorch_resnet18_segmentation")`, `ds.flower.recipe()` and
`ds.flower.submit()` use the same contract. The target is a **mask path role**,
so do not supply class levels, target bounds or tabular feature columns.

The resident encoder is ResNet18 through layer2, pinned to
`ResNet18_Weights.IMAGENET1K_V1` and profile `resnet18_layer2_128_v1`. The full
checkpoint SHA256 is
`f37072fd47e89c5e827621c5baffa7500819f7896bbacec160b1a16c560e07ec`.
Custodians provision these exact weights before training. Missing or mismatched
weights fail preflight. The encoder is frozen, in evaluation mode, outside the
released module; its BatchNorm statistics never learn from private images.

The released decoder has three convolutions (128 to 32, 32 to 16, 16 to 1),
ReLU activations and nearest upsampling by 2 then 4, producing `[1,128,128]`.
The loss mixes per-image mean BCE and smoothed Dice with `alpha=0.5` and
smoothing 1. `alpha=1` is the declared BCE ablation. There is no batch-wide Dice,
trainable encoder, mixed precision, or image augmentation in this first profile.

## Local prediction and public evaluation

```r
probabilities <- ds.flower.predict(
  "segmentation-model", c("public/image-1.png", "public/image-2.png"),
  type = "prob"
)
masks <- ds.flower.predict(
  "segmentation-model", c("public/image-1.png", "public/image-2.png"),
  type = "response"
)
# Both arrays use [image, channel, row, column] = [2, 1, 128, 128].
# Responses threshold probabilities at >= 0.5.
```

Prediction verifies the decoder artifact and the frozen encoder checkpoint
before opening image paths. It shares the node's stored-pixel orientation, RGB
conversion, bilinear 128 by 128 resize and ImageNet channel normalization.
Invalid local images receive fixed zero features. Outputs use the canonical
grid. To map a binary prediction back to the stored native image of height `H`
and width `W`, threshold at 0.5 on the canonical grid first, then use nearest
resampling with the following zero-based coordinate rule (exact ties choose
the larger canonical index):

```text
native[r, c] = canonical[min(127, floor((r + 0.5) * 128 / H)),
                         min(127, floor((c + 0.5) * 128 / W))]
```

This retains stored-pixel orientation, with no EXIF rotation, cropping or
padding. It restores the coordinate grid, not detail discarded by resizing.
For known dimensions from an authorized local image, the same rule in R is:

```r
canonical <- masks[1L, 1L, , ]
native_rows <- pmin(128L, floor((2 * seq_len(native_height) - 1) *
                                128 / (2 * native_height)) + 1L)
native_cols <- pmin(128L, floor((2 * seq_len(native_width) - 1) *
                                128 / (2 * native_width)) + 1L)
native_mask <- canonical[native_rows, native_cols, drop = FALSE]
```

Use these indices explicitly: resize libraries may differ at exact tie pixels
through floating-point coordinate accumulation. Declare and label any
native-resolution secondary metric with this mapping.
The primary benchmark metric remains on the canonical 128 by 128 grid.

Channel-B evaluation uses public or independently authorized analyst-local
held-out subjects. Report mean subject Dice and IoU at threshold 0.5, and report
foreground-positive and empty-mask strata separately. Both-empty Dice is 1;
exactly one empty mask gives Dice 0. Neither images nor predictions are returned
from a private node for evaluation.

Private `ds.flower.validate()`, holdout, cross-validation and private-score HPO
are unsupported for this contract and fail public preflight. The local HPO
callback API does not create a private segmentation scoring route. Any public
benchmark utility claim must cite an executed subject-disjoint protocol, exact
central twins and the recorded gate results; the constructor alone establishes
no utility claim.

## Public benchmark evidence

The preregistered campaign uses BrEaST and BUS-BRA with three DSLite sites,
subject-disjoint 80/20 splits, three seeds, epsilon 1, 4 and 8, delta `1e-5`,
clipping 1 and separate nominal batch-size 16 and 64 arms. Protocol v2 fixes 128 by 128
inputs, five rounds with two local epochs per round, learning rate 0.01,
and no augmentation or mixed precision. BUS-BRA also has the nested
192-subject, BCE-only and heterogeneous-site extensions. The released
federated DP model is compared with pooled DP, pooled nonprivate and
federated nonprivate twins using the same initial decoder, effective tensors,
loss, optimizer and subject splits. Nonprivate twins are public benchmark
diagnostics outside the released contract.
The trivial comparison selects the strongest of four fixed masks: empty,
full, a centered disk of radius 32, and a centered 64 by 64 square.

The tables below read `campaign-status.json`, `breast-evidence.json` and
`busbra-evidence.json` in each arm directory under
`inst/extdata/campaign/segmentation/` at render time. Arms are reported separately;
no result is selected or averaged across arms.
The frozen protocol, dataset release hashes, licence texts, subject-mapping
audits and split hashes are archived alongside the evidence. The
archival assembler validates the complete JSON schemas, captured accountant
steps, effective tensors, released artifacts and twin provenance. Rendering
checks the schema identity, contract and protocol hash again. An incomplete
cohort cannot produce a completed-matrix summary. Execution status is separate
from utility acceptance and reviewer promotion.

```{r segmentation-evidence, eval=TRUE, echo=FALSE, error=FALSE, results="asis"}
evidence_dir <- file.path("..", "inst", "extdata", "campaign", "segmentation")
if (!dir.exists(evidence_dir)) {
  evidence_dir <- system.file("extdata", "campaign", "segmentation",
                              package = "dsFlowerClient")
}
archive_dir <- evidence_dir
render_arm <- function(evidence_dir) {
  if (basename(evidence_dir) %in% c("batch16", "batch64")) {
    cat("## Nominal ", basename(evidence_dir), " arm\n\n", sep = "")
  }
  protocol_hash <- digest::digest(file = file.path(archive_dir, "protocol.md"),
                                  algo = "sha256")
  read_evidence <- function(filename, schema_file) {
    value <- jsonlite::read_json(file.path(evidence_dir, filename))
    schema <- jsonlite::read_json(file.path(archive_dir, schema_file))
    stopifnot(identical(value$schema, schema$`$id`),
              all(unlist(schema$required) %in% names(value)),
              identical(value$contract, "pytorch_resnet18_segmentation"),
              identical(value$task, "segmentation"),
              identical(value$protocol_sha256, protocol_hash),
              value$status %in% c("not_executed", "failed", "executed"))
    if (value$status != "executed") {
      stopifnot(is.character(value$reason), is.null(value$replicates))
    }
    value
  }
  status_schema <- jsonlite::read_json(file.path(evidence_dir, "campaign-status.json"))$schema
  status <- read_evidence("campaign-status.json",
    if (status_schema == "dsflower-segmentation-campaign-summary-v1")
      "campaign-summary-schema.json" else "evidence-schema.json")
  cat("**Campaign status: ", status$status, ".** ", status$reason, "\n\n", sep = "")
  cat("Protocol SHA256: `", protocol_hash, "`.\n\n", sep = "")
  cohort_files <- intersect(c("breast-evidence.json", "busbra-evidence.json"),
                            list.files(evidence_dir))
  cohorts <- lapply(cohort_files, read_evidence, schema_file = "evidence-schema.json")
  names(cohorts) <- sub("-evidence.json", "", cohort_files, fixed = TRUE)
  labels <- c(breast = "BrEaST", busbra = "BUS-BRA")
  variants <- c(full = "Primary", small192 = "192 subjects", bce = "BCE only",
                heterogeneous = "Heterogeneous sites")
  show_table <- function(rows, caption) {
    print(knitr::kable(do.call(rbind, rows), row.names = FALSE, caption = caption))
    cat("\n\n")
  }
  format_metric <- function(value) {
    if (is.null(value)) return("Not applicable")
    stopifnot(value$n_replicates >= 3L, length(value$ci95) == 2L,
              all(is.finite(c(value$mean, value$sd, unlist(value$ci95)))))
    sprintf("%.3f [%.3f, %.3f]", value$mean, value$ci95[[1L]], value$ci95[[2L]])
  }
  if (!length(cohorts)) {
    cat("No completed cohort evidence is archived. Utility floors and envelopes remain **not evaluated**.\n\n")
  }
  for (dataset in names(cohorts)) {
    cohort <- cohorts[[dataset]]
    cat("### ", labels[[dataset]], "\n\n**Cohort status: ", cohort$status, ".**\n\n", sep = "")
    failures <- Filter(function(cell) identical(cell$status, "failed"), cohort$cells)
    if (length(failures)) {
      show_table(lapply(failures, function(cell) data.frame(
        Setting = variants[[cell$variant]], Epsilon = cell$epsilon,
        Seed = cell$seed, Failure = cell$reason)), "Failed cells retained in the archive.")
    }
    if (cohort$status != "executed") {
      cat(cohort$reason, ". Utility floors and envelopes remain **not evaluated**.\n\n", sep = "")
      observed <- list()
      scalar_metric <- function(value) if (is.null(value)) "Not applicable" else sprintf("%.3f", value)
      branches <- c(federated_dp = "Federated DP", pooled_dp = "Pooled DP",
                    pooled_nonprivate = "Pooled nonprivate", federated_nonprivate = "Federated nonprivate")
      for (replicate in cohort$completed_cells) {
        for (branch in names(branches)) {
          values <- replicate[[branch]]
          observed[[length(observed) + 1L]] <- data.frame(
            Epsilon = replicate$epsilon, Seed = replicate$seed, Branch = branches[[branch]],
            Dice = scalar_metric(values$all$dice), IoU = scalar_metric(values$all$iou),
            `Foreground Dice` = scalar_metric(values$foreground_positive$dice),
            `Foreground IoU` = scalar_metric(values$foreground_positive$iou),
            `Empty Dice` = scalar_metric(values$empty_reference$dice),
            `Empty IoU` = scalar_metric(values$empty_reference$iou), check.names = FALSE)
        }
      }
      if (length(observed)) show_table(observed,
        "Observed successful cells from an incomplete matrix; these are not complete three-seed summaries or utility acceptance.")
      next
    }
    stopifnot(length(cohort$replicates) == 9L,
              !is.null(cohort$envelopes), !is.null(cohort$provenance),
              !is.null(cohort$runtime), identical(cohort$dataset, dataset))
    rows <- strata <- list()
    groups <- c(list(full = cohort$summaries), cohort$additional_summaries)
    for (variant in names(groups)) {
      for (epsilon in names(groups[[variant]])) {
        metrics <- groups[[variant]][[epsilon]]
        label <- paste(variants[[variant]], "epsilon", epsilon)
        replicates <- Filter(function(value) value$variant == variant &&
          value$epsilon == as.numeric(epsilon),
          c(cohort$replicates, cohort$additional_replicates))
        stopifnot(length(replicates) == 3L,
          setequal(vapply(replicates, function(value) value$seed, integer(1)),
                   c(20260919L, 20260920L, 20260921L)))
        rows[[length(rows) + 1L]] <- data.frame(
          Setting = label,
          `Federated DP Dice` = format_metric(metrics$federated_dp$all$dice),
          `Pooled DP Dice` = format_metric(metrics$pooled_dp$all$dice),
          `Pooled nonprivate Dice` = format_metric(metrics$pooled_nonprivate$all$dice),
          `Federated nonprivate Dice` = format_metric(metrics$federated_nonprivate$all$dice),
          `Strongest trivial Dice` = sprintf("%.3f", mean(vapply(replicates,
            function(value) value$trivial$strongest_dice, numeric(1)))),
          check.names = FALSE)
        for (stratum in c("all", "foreground_positive", "empty_reference")) {
          values <- metrics$federated_dp[[stratum]]
          strata[[length(strata) + 1L]] <- data.frame(
            Setting = label, Stratum = gsub("_", " ", stratum),
            Dice = format_metric(values$dice), IoU = format_metric(values$iou))
        }
      }
    }
    show_table(rows, "Mean subject Dice with 95% t intervals across three seeds.")
    show_table(strata, "Released-model Dice and IoU at threshold 0.5; absent strata are not applicable.")
    checks <- cohort$envelopes
    floor <- checks$utility_floor
    cat("**Epsilon 8 utility floor: ", if (isTRUE(floor$pass)) "PASS" else "FAILED", ".** ",
        if (dataset == "busbra") "This is the preregistered primary floor. " else
          "This is a descriptive application of the BUS-BRA floor. ",
        sprintf("Mean Dice %.3f; strongest trivial mask %.3f; margin %.3f. ",
                floor$federated_mean_dice, floor$strongest_trivial_mean_dice,
                floor$federated_mean_dice - floor$strongest_trivial_mean_dice),
        "Required: Dice >= 0.50 and margin >= 0.10.\n\n", sep = "")
    envelope_rows <- lapply(checks$epsilon_envelope, function(check) data.frame(
      Check = paste("Epsilon", check$previous, "to", check[["next"]]),
      Result = if (isTRUE(check$flag)) "FLAGGED" else "Not flagged",
      Detail = sprintf("Shortfall increase %.4f; seed-SD tolerance %.4f",
                       check$gap_increase, check$tolerance)))
    trend <- checks$small_n_noise_trend
    envelope_rows[[length(envelope_rows) + 1L]] <- data.frame(
      Check = "Small-cohort noise trend",
      Result = if (trend$status != "executed") "Not evaluated" else
        if (isTRUE(trend$flag)) "FLAGGED" else "Not flagged",
      Detail = if (trend$status != "executed") "Complete matched small-cohort matrix unavailable" else
        sprintf("Shortfall SD: epsilon 1 %.4f; epsilon 8 %.4f", trend$sd_epsilon1, trend$sd_epsilon8))
    near_groups <- c(list(full = checks$near_central), cohort$additional_near_central)
    if (!is.null(trend$near_central)) {
      if (!is.null(near_groups$small192)) {
        stopifnot(identical(near_groups$small192, trend$near_central))
      }
      near_groups$small192 <- trend$near_central
    }
    near_variants <- if (dataset == "busbra") names(variants) else "full"
    for (variant in near_variants) {
      near <- Filter(function(check) isTRUE(check$flag), near_groups[[variant]])
      expected_near <- if (variant %in% c("full", "small192")) 9L else 3L
      observed_near <- length(near_groups[[variant]])
      incomplete_near <- observed_near < expected_near
      envelope_rows[[length(envelope_rows) + 1L]] <- data.frame(
        Check = paste("Unexpected near-central agreement:", variants[[variant]]),
        Result = if (is.null(near_groups[[variant]])) "Not evaluated" else
          if (incomplete_near) "Incomplete" else
            if (length(near)) "FLAGGED" else "Not flagged",
        Detail = if (is.null(near_groups[[variant]])) "Complete matched setting unavailable" else
          paste0("Observed ", observed_near, " of ", expected_near, "; ",
            if (length(near)) paste(vapply(near, function(check)
              paste("FLAGGED epsilon", check$epsilon, "seed", check$seed), character(1)), collapse = "; ") else
              "no observed replicate met the investigation rule"))
    }
    show_table(envelope_rows, "Envelope checks; flags require investigation and are retained.")
  }
}
arm_dirs <- file.path(archive_dir, c("batch16", "batch64"))
arm_dirs <- arm_dirs[file.exists(file.path(arm_dirs, "campaign-status.json"))]
if (!length(arm_dirs)) arm_dirs <- archive_dir
invisible(lapply(arm_dirs, render_arm))

```

The shortfall used by the envelope is **pooled nonprivate Dice minus
federated DP Dice**. Positive values indicate lower utility for the released
model. Epsilon-envelope flags compare changes against seed dispersion;
near-central flags identify unexpectedly close agreement under the
preregistered small-cohort rule. An unflagged envelope does not establish
privacy by itself. Failed utility floors remain failed and do not trigger
changes to epsilon, delta, clipping, subject retention or the preregistered
protocol. Foreground-positive results remain visible even when empty-mask
performance changes the overall mean.
