---
title: "Clinical Privacy Benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical Privacy Benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
find_evidence <- function(file) {
  candidates <- c(
    file.path("..", "inst", "extdata", file),
    file.path("inst", "extdata", file),
    system.file("extdata", file, package = "dsFlowerClient")
  )
  candidates[nzchar(candidates) & file.exists(candidates)][1]
}
read_evidence <- function(file) {
  path <- find_evidence(file)
  if (is.na(path)) stop("Evidence file not found: ", file, call. = FALSE)
  jsonlite::fromJSON(path, simplifyVector = FALSE)
}
flat_results <- function(x, label) {
  rows <- lapply(names(x$results), function(id) {
    r <- x$results[[id]]
    history <- r$history %||% list()
    failures <- sum(vapply(history, function(h) h$n_failures %||% 0L, integer(1)))
    policy <- r$privacy_policy %||% x$privacy_policy %||% list()
    data.frame(
      evidence = label,
      privacy_profile = x$privacy_profile,
      privacy_mechanism = policy$privacy_mechanism %||% NA_character_,
      secagg_required = policy$secure_aggregation_required %||% NA,
      dp_scope = policy$dp_scope %||% "none",
      epsilon = r$curve_epsilon %||% r$privacy_parameters$epsilon %||% NA_real_,
      dataset = r$dataset_label,
      model = r$model_label,
      status = r$status,
      n = r$n_total,
      train_test = paste(r$n_train, r$n_test, sep = "/"),
      sites = paste(unlist(r$site_train_n), collapse = "/"),
      rounds = r$rounds %||% NA_integer_,
      central_auc = r$central_metrics$auc,
      federated_auc = r$federated_metrics$auc,
      delta_auc = r$metric_delta$auc,
      central_accuracy = r$central_metrics$accuracy,
      federated_accuracy = r$federated_metrics$accuracy,
      delta_accuracy = r$metric_delta$accuracy,
      failures = failures,
      stringsAsFactors = FALSE
    )
  })
  do.call(rbind, rows)
}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || (length(x) == 1 && is.na(x))) y else x
}
trusted <- read_evidence("dsflower_clinical_algorithm_results.json")
secagg <- read_evidence("dsflower_clinical_secagg_results.json")
dp <- read_evidence("dsflower_clinical_dp_results.json")
dp_curve <- read_evidence("dsflower_clinical_dp_curve_results.json")
results <- rbind(
  flat_results(trusted, "trusted_internal"),
  flat_results(secagg, "clinical_default_secagg"),
  flat_results(dp, "high_sensitivity_dp")
)
curve_results <- flat_results(dp_curve, "dp_epsilon_curve")
curve_results$epsilon <- vapply(dp_curve$results, `[[`, numeric(1), "curve_epsilon")
```

This vignette reports public clinical benchmark runs executed through the same
DataSHIELD/Flower path used by dsFlower applications. Each run partitions the
training data over three Opal/Rock servers, launches the selected then-authorised
template through `ds.flower.fit()`, trains a centralised baseline on the same
training split, evaluates both models on the same held-out test set, and records
the approved output metrics.

The datasets are public health or biomedical classification datasets:
Breast Cancer Wisconsin (Original), UCI Heart Disease processed Cleveland,
Pima Indians Diabetes, and CDC Diabetes Health Indicators. The benchmark matrix
matches each privacy profile to dataset/model combinations that satisfy the
active server-side row policy. Heart Disease contributes to the trusted-internal
evidence, while the stricter clinical and DP profiles use larger datasets so
that the row policy is satisfied before training starts.

```{r dataset-map}
dataset_map <- data.frame(
  dataset = c(
    "Breast Cancer Wisconsin (Original)",
    "UCI Heart Disease processed Cleveland",
    "Pima Indians Diabetes",
    "CDC Diabetes Health Indicators"
  ),
  role = c(
    "Small cytology classification task used for trusted and SecAgg checks.",
    "Cardiology classification task retained in trusted-internal evidence; the clinical row policy intentionally excludes it.",
    "Medium-size diabetes-onset task used for clinical SecAgg logistic regression.",
    "Larger population-health diabetes task used for the broader clinical and DP-SGD evidence."
  ),
  validated_profiles = c(
    "trusted_internal; clinical_default",
    "trusted_internal",
    "trusted_internal; clinical_default",
    "trusted_internal; clinical_default; high_sensitivity_dp"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(dataset_map)
```

It is laid out in two parts: first the commands you would run to reproduce the
benchmark, then the recorded results loaded from the committed evidence
artifacts. No live Opal servers are required to render this page; the command
chunks are shown for reference and are not executed.

## Running This Benchmark

Each dataset is partitioned across three Opal nodes that are already set up and
serving the cohort table; each benchmark row is one `ds.flower.fit()` call over
the server-side partitions. Only aggregate training history and model parameters
are returned to the researcher.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that hold the partitioned cohort (the CDC Diabetes Health Indicators
table is shown).

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = "D")
```

**2. Run a benchmark row.** Historically, train one then-authorised template through dsFlower with
FedAvg over the server-side table, choosing the privacy profile for the posture
under test. The clinical Secure Aggregation profile is shown here.

```{r fit, eval=FALSE}
fit <- ds.flower.fit(
  conns,
  symbol   = "D",
  target   = "outcome",
  model    = "sklearn_logreg",
  strategy = "fedavg",
  privacy  = "clinical_default",
  rounds   = 5
)

DSI::datashield.logout(conns)
```

Swapping the `privacy` argument moves between postures: `"trusted_internal"` for
ordinary federated training, `"clinical_default"` for the Secure Aggregation
profile over the `sklearn_logreg`, `sklearn_ridge`, `sklearn_sgd`,
`pytorch_logreg` and `pytorch_mlp` templates, and `"high_sensitivity_dp"` for
Opacus DP-SGD on the compatible PyTorch templates. The epsilon curves below
repeat the DP rows across a sweep of requested budgets.

## Evidence map

The committed evidence is split by privacy posture. The `trusted_internal`
profile exercises ordinary federated training on public datasets. The
`clinical_default` profile requires Secure Aggregation, fixed three-client
participation and suppressed per-node metrics. `high_sensitivity_dp` adds
patient-level DP-SGD for compatible PyTorch losses.

```{r evidence-map}
evidence_map <- data.frame(
  evidence = c(
    "trusted_internal",
    "clinical_default_secagg",
    "high_sensitivity_dp"
  ),
  datasets = c(
    paste(unique(subset(results, evidence == "trusted_internal")$dataset),
          collapse = "; "),
    paste(unique(subset(results, evidence == "clinical_default_secagg")$dataset),
          collapse = "; "),
    paste(unique(subset(results, evidence == "high_sensitivity_dp")$dataset),
          collapse = "; ")
  ),
  templates = c(
    paste(unique(subset(results, evidence == "trusted_internal")$model),
          collapse = "; "),
    paste(unique(subset(results, evidence == "clinical_default_secagg")$model),
          collapse = "; "),
    paste(unique(subset(results, evidence == "high_sensitivity_dp")$model),
          collapse = "; ")
  ),
  n_runs = c(
    nrow(subset(results, evidence == "trusted_internal")),
    nrow(subset(results, evidence == "clinical_default_secagg")),
    nrow(subset(results, evidence == "high_sensitivity_dp"))
  ),
  failures = c(
    sum(subset(results, evidence == "trusted_internal")$failures),
    sum(subset(results, evidence == "clinical_default_secagg")$failures),
    sum(subset(results, evidence == "high_sensitivity_dp")$failures)
  ),
  stringsAsFactors = FALSE
)
knitr::kable(evidence_map, row.names = FALSE)
```

## Result tables

```{r policy-table}
policy_table <- unique(
  results[, c("evidence", "privacy_profile", "privacy_mechanism",
              "secagg_required", "dp_scope")]
)
knitr::kable(policy_table, row.names = FALSE)
```

```{r trusted-table}
knitr::kable(
  subset(results, evidence == "trusted_internal"),
  digits = 4,
  row.names = FALSE
)
```

```{r secagg-table}
knitr::kable(
  subset(results, evidence == "clinical_default_secagg"),
  digits = 4,
  row.names = FALSE
)
```

```{r dp-table}
knitr::kable(
  subset(results, evidence == "high_sensitivity_dp"),
  digits = 4,
  row.names = FALSE
)
```

The `clinical_default` runs require Secure Aggregation, suppress per-node
metrics and avoid exact per-site count release. This evidence covers logistic
regression, ridge, SGD, PyTorch MLP and PyTorch logistic regression. The
`high_sensitivity_dp` evidence uses Opacus DP-SGD with `delta = 1e-5` and
clipping norm `1` for both PyTorch MLP and PyTorch logistic regression, in
addition to the Secure Aggregation requirement.

## DP Utility Curve

```{r dp-curve-table}
knitr::kable(
  curve_results[, c("epsilon", "dataset", "model", "rounds", "central_auc",
                    "federated_auc", "delta_auc", "failures")],
  digits = 4,
  row.names = FALSE
)
```

```{r dp-curve-plot, fig.width=6.4, fig.height=3.6, fig.alt="Differential-privacy utility curves for PyTorch logistic regression and MLP on CDC Diabetes Health Indicators."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  central_lines <- unique(curve_results[, c("model", "central_auc")])
  ggplot2::ggplot(
    curve_results,
    ggplot2::aes(x = epsilon, y = federated_auc, color = model, group = model)
  ) +
    ggplot2::geom_hline(
      data = central_lines,
      ggplot2::aes(yintercept = central_auc, color = model),
      linewidth = 0.3,
      linetype = "dashed",
      inherit.aes = FALSE
    ) +
    ggplot2::geom_line(linewidth = 0.5) +
    ggplot2::geom_point(size = 2.2) +
    ggplot2::scale_x_continuous(breaks = c(2, 4, 8)) +
    ggplot2::labs(
      x = "Requested epsilon",
      y = "Held-out AUC",
      color = "Template",
      caption = "Dashed lines: centralised non-DP baselines on the same split."
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}
```

## Centralised versus federated AUC

```{r auc-plot, fig.width=8, fig.height=5, fig.alt="Federated minus centralised AUC deltas by dataset, model and privacy profile."}
plot_df <- results
plot_df$label <- paste(plot_df$privacy_profile, plot_df$dataset, plot_df$model, sep = " / ")
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    plot_df,
    ggplot2::aes(x = stats::reorder(label, delta_auc), y = delta_auc,
                 fill = privacy_profile)
  ) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = "grey35") +
    ggplot2::geom_col(width = 0.72) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = "Federated AUC - centralised AUC") +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}
```

All reported retained runs completed with status `pass` and zero recorded
client failures.
