---
title: "Clinical Privacy Benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical Privacy Benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
find_evidence <- function(file) {
  candidates <- c(
    file.path("..", "inst", "extdata", file),
    file.path("inst", "extdata", file),
    system.file("extdata", file, package = "dsFlowerClient")
  )
  candidates[nzchar(candidates) & file.exists(candidates)][1]
}
read_evidence <- function(file) {
  path <- find_evidence(file)
  if (is.na(path)) stop("Evidence file not found: ", file, call. = FALSE)
  jsonlite::fromJSON(path, simplifyVector = FALSE)
}
flat_results <- function(x, label) {
  rows <- lapply(names(x$results), function(id) {
    r <- x$results[[id]]
    history <- r$history %||% list()
    failures <- sum(vapply(history, function(h) h$n_failures %||% 0L, integer(1)))
    policy <- r$privacy_policy %||% x$privacy_policy %||% list()
    data.frame(
      evidence = label,
      privacy_profile = x$privacy_profile,
      privacy_mechanism = policy$privacy_mechanism %||% NA_character_,
      secagg_required = policy$secure_aggregation_required %||% NA,
      dp_scope = policy$dp_scope %||% "none",
      epsilon = r$curve_epsilon %||% r$privacy_parameters$epsilon %||% NA_real_,
      dataset = r$dataset_label,
      model = r$model_label,
      status = r$status,
      n = r$n_total,
      train_test = paste(r$n_train, r$n_test, sep = "/"),
      sites = paste(unlist(r$site_train_n), collapse = "/"),
      rounds = r$rounds %||% NA_integer_,
      central_auc = r$central_metrics$auc,
      federated_auc = r$federated_metrics$auc,
      delta_auc = r$metric_delta$auc,
      central_accuracy = r$central_metrics$accuracy,
      federated_accuracy = r$federated_metrics$accuracy,
      delta_accuracy = r$metric_delta$accuracy,
      failures = failures,
      stringsAsFactors = FALSE
    )
  })
  do.call(rbind, rows)
}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || (length(x) == 1 && is.na(x))) y else x
}
trusted <- read_evidence("dsflower_clinical_algorithm_results.json")
secagg <- read_evidence("dsflower_clinical_secagg_results.json")
dp <- read_evidence("dsflower_clinical_dp_results.json")
xgb_noise <- read_evidence("dsflower_xgboost_histogram_privacy_results.json")
dp_curve <- read_evidence("dsflower_clinical_dp_curve_results.json")
xgb_curve <- read_evidence("dsflower_xgboost_histogram_privacy_curve_results.json")
results <- rbind(
  flat_results(trusted, "trusted_internal"),
  flat_results(secagg, "clinical_default_secagg"),
  flat_results(dp, "high_sensitivity_dp"),
  flat_results(xgb_noise, "xgboost_histogram_update_noise")
)
curve_results <- flat_results(dp_curve, "dp_epsilon_curve")
curve_results$epsilon <- vapply(dp_curve$results, `[[`, numeric(1), "curve_epsilon")
xgb_curve_results <- flat_results(xgb_curve, "xgboost_update_noise_curve")
xgb_curve_results$epsilon <- vapply(xgb_curve$results, `[[`, numeric(1), "curve_epsilon")
```

This vignette reports public clinical benchmark runs executed through the same
DataSHIELD/Flower path used by dsFlower applications. Each run partitions the
training data over three Opal/Rock servers, launches the selected authorised
template through `ds.flower.fit()`, trains a centralised baseline on the same
training split, evaluates both models on the same held-out test set, and records
the approved output metrics.

The datasets are public health or biomedical classification datasets:
Breast Cancer Wisconsin (Original), UCI Heart Disease processed Cleveland,
Pima Indians Diabetes, and CDC Diabetes Health Indicators. The benchmark matrix
matches each privacy profile to dataset/model combinations that satisfy the
active server-side row policy. Heart Disease contributes to the trusted-internal
evidence, while the stricter clinical and DP profiles use larger datasets so
that the row policy is satisfied before training starts.

```{r dataset-map}
dataset_map <- data.frame(
  dataset = c(
    "Breast Cancer Wisconsin (Original)",
    "UCI Heart Disease processed Cleveland",
    "Pima Indians Diabetes",
    "CDC Diabetes Health Indicators"
  ),
  role = c(
    "Small cytology classification task used for trusted and SecAgg checks.",
    "Cardiology classification task retained in trusted-internal evidence; the clinical row policy intentionally excludes it.",
    "Medium-size diabetes-onset task used for clinical SecAgg logistic regression.",
    "Larger population-health diabetes task used for the broader clinical, DP-SGD and XGBoost privacy-profile evidence."
  ),
  validated_profiles = c(
    "trusted_internal; clinical_default",
    "trusted_internal",
    "trusted_internal; clinical_default",
    "trusted_internal; clinical_default; clinical_update_noise; high_sensitivity_dp"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(dataset_map)
```

It is laid out in two parts: first the commands you would run to reproduce the
benchmark, then the recorded results loaded from the committed evidence
artifacts. No live Opal servers are required to render this page; the command
chunks are shown for reference and are not executed.

## Running This Benchmark

Each dataset is partitioned across three Opal nodes that are already set up and
serving the cohort table; each benchmark row is one `ds.flower.fit()` call over
the server-side partitions. Only aggregate training history and model parameters
are returned to the researcher.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that hold the partitioned cohort (the CDC Diabetes Health Indicators
table is shown).

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.cdc_diabetes_cohort")
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = "D")
```

**2. Run a benchmark row.** Train one authorised template through dsFlower with
FedAvg over the server-side table, choosing the privacy profile for the posture
under test. The clinical Secure Aggregation profile is shown here.

```{r fit, eval=FALSE}
fit <- ds.flower.fit(
  conns,
  symbol   = "D",
  target   = "outcome",
  model    = "sklearn_logreg",
  strategy = "fedavg",
  privacy  = "clinical_default",
  rounds   = 5
)

DSI::datashield.logout(conns)
```

Swapping the `privacy` argument moves between postures: `"trusted_internal"` for
ordinary federated training, `"clinical_default"` for the Secure Aggregation
profile over the `sklearn_logreg`, `sklearn_ridge`, `sklearn_sgd`,
`pytorch_logreg`, `pytorch_mlp` and secure `xgboost_histogram` templates,
`"high_sensitivity_dp"` for Opacus DP-SGD on the compatible PyTorch templates,
and `"clinical_update_noise"` for the bounded-noise XGBoost histogram route. The
epsilon curves below repeat the DP and update-noise rows across a sweep of
requested budgets.

## Evidence map

The committed evidence is split by privacy posture. The `trusted_internal`
profile exercises ordinary federated training on public datasets. The
`clinical_default` profile requires Secure Aggregation, fixed three-client
participation and suppressed per-node metrics. `high_sensitivity_dp` adds
patient-level DP-SGD for compatible PyTorch losses, and
`clinical_update_noise` applies bounded histogram noise to the secure XGBoost
histogram route.

```{r evidence-map}
evidence_map <- data.frame(
  evidence = c(
    "trusted_internal",
    "clinical_default_secagg",
    "high_sensitivity_dp",
    "xgboost_histogram_update_noise"
  ),
  datasets = c(
    paste(unique(subset(results, evidence == "trusted_internal")$dataset),
          collapse = "; "),
    paste(unique(subset(results, evidence == "clinical_default_secagg")$dataset),
          collapse = "; "),
    paste(unique(subset(results, evidence == "high_sensitivity_dp")$dataset),
          collapse = "; "),
    paste(unique(subset(results, evidence == "xgboost_histogram_update_noise")$dataset),
          collapse = "; ")
  ),
  templates = c(
    paste(unique(subset(results, evidence == "trusted_internal")$model),
          collapse = "; "),
    paste(unique(subset(results, evidence == "clinical_default_secagg")$model),
          collapse = "; "),
    paste(unique(subset(results, evidence == "high_sensitivity_dp")$model),
          collapse = "; "),
    paste(unique(subset(results, evidence == "xgboost_histogram_update_noise")$model),
          collapse = "; ")
  ),
  n_runs = c(
    nrow(subset(results, evidence == "trusted_internal")),
    nrow(subset(results, evidence == "clinical_default_secagg")),
    nrow(subset(results, evidence == "high_sensitivity_dp")),
    nrow(subset(results, evidence == "xgboost_histogram_update_noise"))
  ),
  failures = c(
    sum(subset(results, evidence == "trusted_internal")$failures),
    sum(subset(results, evidence == "clinical_default_secagg")$failures),
    sum(subset(results, evidence == "high_sensitivity_dp")$failures),
    sum(subset(results, evidence == "xgboost_histogram_update_noise")$failures)
  ),
  stringsAsFactors = FALSE
)
knitr::kable(evidence_map, row.names = FALSE)
```

## Result tables

```{r policy-table}
policy_table <- unique(
  results[, c("evidence", "privacy_profile", "privacy_mechanism",
              "secagg_required", "dp_scope")]
)
knitr::kable(policy_table, row.names = FALSE)
```

```{r trusted-table}
knitr::kable(
  subset(results, evidence == "trusted_internal"),
  digits = 4,
  row.names = FALSE
)
```

```{r secagg-table}
knitr::kable(
  subset(results, evidence == "clinical_default_secagg"),
  digits = 4,
  row.names = FALSE
)
```

```{r dp-table}
knitr::kable(
  subset(results, evidence == "high_sensitivity_dp"),
  digits = 4,
  row.names = FALSE
)
```

```{r xgboost-update-noise-table}
knitr::kable(
  subset(results, evidence == "xgboost_histogram_update_noise"),
  digits = 4,
  row.names = FALSE
)
```

The `clinical_default` runs require Secure Aggregation, suppress per-node
metrics and avoid exact per-site count release. This evidence covers logistic
regression, ridge, SGD, PyTorch MLP, PyTorch logistic regression and the
secure boosted-stump path of the XGBoost histogram template. In this secure
path, each boosting round aggregates one fixed-shape root histogram under
SecAgg+; the server derives one split and two leaf values from the protected
aggregate, and the next round asks clients to apply the previous stump before
building the next encrypted histogram. The representative `clinical_update_noise`
XGBoost row uses the `epsilon = 12` point from the committed update-noise curve:
Gaussian noise is added to the bounded histogram contribution before aggregation,
which demonstrates the stricter privacy profile and its utility cost. The
`high_sensitivity_dp` evidence uses Opacus DP-SGD with `delta = 1e-5` and
clipping norm `1` for both PyTorch MLP and PyTorch logistic regression, in
addition to the Secure Aggregation requirement.

## DP Utility Curve

```{r dp-curve-table}
knitr::kable(
  curve_results[, c("epsilon", "dataset", "model", "rounds", "central_auc",
                    "federated_auc", "delta_auc", "failures")],
  digits = 4,
  row.names = FALSE
)
```

```{r dp-curve-plot, fig.width=6.4, fig.height=3.6, fig.alt="Differential-privacy utility curves for PyTorch logistic regression and MLP on CDC Diabetes Health Indicators."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  central_lines <- unique(curve_results[, c("model", "central_auc")])
  ggplot2::ggplot(
    curve_results,
    ggplot2::aes(x = epsilon, y = federated_auc, color = model, group = model)
  ) +
    ggplot2::geom_hline(
      data = central_lines,
      ggplot2::aes(yintercept = central_auc, color = model),
      linewidth = 0.3,
      linetype = "dashed",
      inherit.aes = FALSE
    ) +
    ggplot2::geom_line(linewidth = 0.5) +
    ggplot2::geom_point(size = 2.2) +
    ggplot2::scale_x_continuous(breaks = c(2, 4, 8)) +
    ggplot2::labs(
      x = "Requested epsilon",
      y = "Held-out AUC",
      color = "Template",
      caption = "Dashed lines: centralised non-DP baselines on the same split."
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}
```

## XGBoost Histogram Update-Noise Curve

```{r xgboost-update-noise-curve-table}
knitr::kable(
  xgb_curve_results[, c("epsilon", "dataset", "model", "central_auc",
                        "federated_auc", "delta_auc", "failures")],
  digits = 4,
  row.names = FALSE
)
```

```{r xgboost-update-noise-curve-plot, fig.width=5.8, fig.height=3.3, fig.alt="Update-noise utility curve for secure histogram XGBoost on CDC Diabetes Health Indicators."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    xgb_curve_results,
    ggplot2::aes(x = epsilon, y = federated_auc)
  ) +
    ggplot2::geom_hline(
      yintercept = xgb_curve_results$central_auc[[1]],
      linewidth = 0.3,
      linetype = "dashed",
      color = "grey35"
    ) +
    ggplot2::geom_line(linewidth = 0.5, color = "#704C5E") +
    ggplot2::geom_point(size = 2.2, color = "#704C5E") +
    ggplot2::scale_x_continuous(breaks = c(8, 12, 16)) +
    ggplot2::labs(
      x = "Requested epsilon",
      y = "Held-out AUC",
      caption = "Dashed line: centralised non-noisy baseline on the same split."
    ) +
    ggplot2::theme_minimal(base_size = 11)
}
```

The XGBoost curve uses the `clinical_update_noise` profile rather than
patient-level DP-SGD. The noise is added to bounded histogram contributions
before protected aggregation, so it demonstrates the utility trade-off for a
stricter histogram-update profile. The retained curve shows the expected
pattern: the most aggressive budget has the largest utility cost, while the
moderate and relaxed budgets move back toward the SecAgg-only reference.

## Centralised versus federated AUC

```{r auc-plot, fig.width=8, fig.height=5, fig.alt="Federated minus centralised AUC deltas by dataset, model and privacy profile."}
plot_df <- results
plot_df$label <- paste(plot_df$privacy_profile, plot_df$dataset, plot_df$model, sep = " / ")
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    plot_df,
    ggplot2::aes(x = stats::reorder(label, delta_auc), y = delta_auc,
                 fill = privacy_profile)
  ) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = "grey35") +
    ggplot2::geom_col(width = 0.72) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = "Federated AUC - centralised AUC") +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}
```

All reported runs completed with status `pass` and zero recorded client
failures. The full XGBoost histogram implementation remains available for
trusted internal validation. Clinical profiles use secure boosted stumps rather
than arbitrary-depth tree construction because the latter requires variable
split and leaf-control phases that are not a stable fit for SecAgg+ workflows.
The secure route still validates the privacy-relevant computation: histogram
construction on each server, protected aggregation of those histograms, and
server-side split and leaf computation from the aggregate only.
