---
title: "Clinical method-family benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical method-family benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dsFlowerClient)
```

This vignette records the non-binary clinical method-family validation used for
the dsFlower thesis evidence. The benchmark uses the SUPPORT2 study from the
UCI Machine Learning Repository and executes five PyTorch templates under the
`clinical_default` profile, so server-side row policies and Secure Aggregation
are active.

The benchmark complements the binary clinical benchmarks: classification and
XGBoost are covered there, while this vignette covers continuous regression,
count regression, multiclass classification, multilabel classification and
time-to-event survival.

```{r load-evidence}
evidence_path <- system.file(
  "extdata", "dsflower_method_family_results.json",
  package = "dsFlowerClient"
)
evidence <- jsonlite::fromJSON(evidence_path)

cat("Dataset:", evidence$dataset_label, "\n")
cat("Privacy profile:", evidence$privacy_profile, "\n")
cat("Secure Aggregation supported:", evidence$secagg_supported, "\n")
cat("Rows:", evidence$n_total, "\n")
print(as.data.frame(evidence$site_train_n))
```

The rows are split evenly across three DataSHIELD servers. The event count shown
below is used by the survival route disclosure guard.

```{r site-events}
site_summary <- data.frame(
  server = names(evidence$site_train_n),
  train_n = as.integer(unlist(evidence$site_train_n, use.names = FALSE)),
  event_n = as.integer(unlist(evidence$site_event_n, use.names = FALSE))
)
site_summary
```

Each row in the result table compares the centralized PyTorch loss against the
loss obtained by evaluating the federated model artefact saved by dsFlower after
the SecAgg run completed. The route is accepted only when all clients complete
without failures and the federated loss remains within the declared validation
envelope for that method family.

```{r results-table}
results <- evidence$results
results[, c(
  "family", "method", "centralized_metric", "centralized_loss",
  "federated_loss", "delta_loss", "federated_n_failures",
  "validation_status"
)]
```

```{r loss-plot, fig.width=7, fig.height=4.2, fig.alt="Point-range chart comparing centralized and federated SUPPORT2 losses across method families."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- results
  plot_data$label <- sub("^pytorch_", "", plot_data$method)
  plot_data$abs_delta <- abs(plot_data$delta_loss)
  ggplot2::ggplot(plot_data, ggplot2::aes(
    x = stats::reorder(label, abs_delta),
    y = abs_delta
  )) +
    ggplot2::geom_col(width = 0.68, fill = "#4D6A7F") +
    ggplot2::coord_flip() +
    ggplot2::labs(
      x = NULL,
      y = "Absolute federated-centralized loss delta",
      title = "SUPPORT2 method-family validation under clinical_default"
    ) +
    ggplot2::theme_minimal(base_size = 12)
}
```

The canonical run can be regenerated from a local three-Opal demonstration
deployment with:

```{r rerun-command, eval=FALSE}
Sys.setenv(
  DSFLOWER_SUPPORT2_LIMIT = "3000",
  DSFLOWER_FAMILY_PRIVACY_PROFILE = "clinical_default"
)
source(system.file(
  "demos", "benchmark_method_families.R",
  package = "dsFlowerClient"
))
```

