---
title: "Clinical method-family benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical method-family benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dsFlowerClient)
```

This vignette records the non-binary clinical method-family validation used for
the dsFlower thesis evidence. The benchmark uses SUPPORT2, the Study to
Understand Prognoses and Preferences for Outcomes and Risks of Treatments, from
the UCI Machine Learning Repository and executes PyTorch templates under the
`clinical_default` profile, so server-side row policies and Secure Aggregation
are active. A second evidence file repeats the DP-compatible non-survival
families under `high_sensitivity_dp`, where SecAgg+ is combined with Opacus
DP-SGD.

The benchmark complements the binary clinical benchmarks: classification and
XGBoost are covered there, while this vignette covers continuous regression,
count regression, multiclass classification, multilabel classification and
time-to-event survival.

```{r task-map}
task_map <- data.frame(
  family = c(
    "Continuous regression",
    "Count regression",
    "Multiclass classification",
    "Multilabel classification",
    "Survival"
  ),
  target = c(
    "Severity score derived from SUPPORT2 variables",
    "Comorbidity count",
    "Diagnosis class",
    "Death, hospital death and diabetes labels",
    "Hospital time-to-event with event indicator"
  ),
  template = c(
    "pytorch_linear_regression",
    "pytorch_poisson",
    "pytorch_multiclass",
    "pytorch_multilabel",
    "pytorch_coxph"
  ),
  privacy_route = c(
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg only: Cox partial likelihood is risk-set based"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(task_map)
```

```{r load-evidence}
evidence_path <- system.file(
  "extdata", "dsflower_method_family_results.json",
  package = "dsFlowerClient"
)
evidence <- jsonlite::fromJSON(evidence_path)
dp_evidence_path <- system.file(
  "extdata", "dsflower_method_family_dp_results.json",
  package = "dsFlowerClient"
)
dp_evidence <- jsonlite::fromJSON(dp_evidence_path)

cat("Dataset:", evidence$dataset_label, "\n")
cat("Privacy profile:", evidence$privacy_profile, "\n")
cat("Secure Aggregation supported:", evidence$secagg_supported, "\n")
cat("Rows:", evidence$n_total, "\n")
print(as.data.frame(evidence$site_train_n))
```

The rows are split evenly across three DataSHIELD servers. The event count shown
below is used by the survival route disclosure guard.

```{r site-events}
site_summary <- data.frame(
  server = names(evidence$site_train_n),
  train_n = as.integer(unlist(evidence$site_train_n, use.names = FALSE)),
  event_n = as.integer(unlist(evidence$site_event_n, use.names = FALSE))
)
site_summary
```

Each row in the result table compares the centralized PyTorch loss against the
loss obtained by evaluating the federated model artefact saved by dsFlower after
the SecAgg run completed. The route is accepted only when all clients complete
without failures and the federated loss remains within the declared validation
envelope for that method family.

```{r results-table}
results <- evidence$results
results[, c(
  "family", "method", "centralized_metric", "centralized_loss",
  "federated_loss", "delta_loss", "federated_n_failures",
  "validation_status"
)]
```

```{r dp-results-table}
dp_results <- dp_evidence$results
dp_results[, c(
  "family", "method", "centralized_metric", "centralized_loss",
  "federated_loss", "delta_loss", "federated_n_failures",
  "validation_status"
)]
```

The DP-SGD table is defined over the SUPPORT2 routes whose losses decompose by
sample. CoxPH is validated under Secure Aggregation in the table above; its
partial-likelihood objective couples samples through risk sets, so its validated
privacy route is SecAgg rather than Opacus DP-SGD. The four non-survival
families shown here are the DP-SGD-compatible SUPPORT2 routes; all four complete
under the stricter profile and remain within their declared utility envelopes.
Multiclass classification uses six federated rounds in this profile because the
stricter DP-SGD optimiser required a slightly longer training budget than the
non-DP family run.

```{r compatibility-table}
compatibility <- data.frame(
  family = c(
    "Continuous regression",
    "Count regression",
    "Multiclass classification",
    "Multilabel classification",
    "Survival"
  ),
  clinical_default = c("pass", "pass", "pass", "pass", "pass"),
  high_sensitivity_dp = c("pass", "pass", "pass", "pass", "SecAgg route"),
  reason = c(
    "MSE is decomposable by row.",
    "Poisson NLL is decomposable by row.",
    "Cross entropy is decomposable by row.",
    "Binary cross entropy is decomposable by row.",
    "Cox partial likelihood is validated under SecAgg."
  ),
  stringsAsFactors = FALSE
)
knitr::kable(compatibility)
```

```{r loss-plot, fig.width=7, fig.height=4.2, fig.alt="Point-range chart comparing centralized and federated SUPPORT2 losses across method families."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- results
  plot_data$label <- sub("^pytorch_", "", plot_data$method)
  plot_data$abs_delta <- abs(plot_data$delta_loss)
  ggplot2::ggplot(plot_data, ggplot2::aes(
    x = stats::reorder(label, abs_delta),
    y = abs_delta
  )) +
    ggplot2::geom_col(width = 0.68, fill = "#4D6A7F") +
    ggplot2::coord_flip() +
    ggplot2::labs(
      x = NULL,
      y = "Absolute federated-centralized loss delta",
      title = "SUPPORT2 method-family validation under clinical_default"
    ) +
    ggplot2::theme_minimal(base_size = 12)
}
```

```{r dp-loss-plot, fig.width=7, fig.height=3.8, fig.alt="Bar chart comparing DP-SGD federated-centralized SUPPORT2 losses for DP-compatible method families."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- dp_results
  plot_data$label <- sub("^pytorch_", "", plot_data$method)
  plot_data$abs_delta <- abs(plot_data$delta_loss)
  ggplot2::ggplot(plot_data, ggplot2::aes(
    x = stats::reorder(label, abs_delta),
    y = abs_delta
  )) +
    ggplot2::geom_col(width = 0.68, fill = "#6F5B3E") +
    ggplot2::coord_flip() +
    ggplot2::labs(
      x = NULL,
      y = "Absolute federated-centralized loss delta",
      title = "SUPPORT2 DP-SGD subset under high_sensitivity_dp"
    ) +
    ggplot2::theme_minimal(base_size = 12)
}
```

The canonical run can be regenerated from a local three-Opal demonstration
deployment with:

```{r rerun-command, eval=FALSE}
Sys.setenv(
  DSFLOWER_SUPPORT2_LIMIT = "3000",
  DSFLOWER_FAMILY_PRIVACY_PROFILE = "clinical_default"
)
source(system.file(
  "demos", "benchmark_method_families.R",
  package = "dsFlowerClient"
))
```

```{r rerun-dp-command, eval=FALSE}
Sys.setenv(
  DSFLOWER_SUPPORT2_LIMIT = "3000",
  DSFLOWER_FAMILY_PRIVACY_PROFILE = "high_sensitivity_dp",
  DSFLOWER_FAMILY_MODELS =
    "pytorch_linear_regression,pytorch_poisson,pytorch_multiclass,pytorch_multilabel",
  DSFLOWER_FAMILY_MULTICLASS_ROUNDS = "6",
  DSFLOWER_FAMILY_DP_EPSILON = "8",
  DSFLOWER_FAMILY_DP_DELTA = "1e-5",
  DSFLOWER_FAMILY_DP_CLIP = "1",
  DSFLOWER_FAMILY_EVIDENCE_FILE =
    "inst/extdata/dsflower_method_family_dp_results.json"
)
source(system.file(
  "demos", "benchmark_method_families.R",
  package = "dsFlowerClient"
))
```
