---
title: "Clinical method-family benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical method-family benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

> The Cox/survival route shown here is not registered or runnable in the current
> enforced-DP runtime. Its calls are non-executed historical evidence only.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(dsFlowerClient)
```

This vignette records the non-binary clinical method-family validation used for
the dsFlower thesis evidence. The benchmark uses SUPPORT2, the Study to
Understand Prognoses and Preferences for Outcomes and Risks of Treatments, from
the UCI Machine Learning Repository and executes PyTorch templates under the
`clinical_default` profile, so server-side row policies and Secure Aggregation
are active. A second evidence file repeats the DP-compatible non-survival
families under `high_sensitivity_dp`, where SecAgg+ is combined with Opacus
DP-SGD.

The benchmark complements the binary clinical benchmarks by covering continuous
regression, count regression, multiclass classification, multilabel
classification and time-to-event survival.

```{r task-map}
task_map <- data.frame(
  family = c(
    "Continuous regression",
    "Count regression",
    "Multiclass classification",
    "Multilabel classification",
    "Survival"
  ),
  target = c(
    "Severity score derived from SUPPORT2 variables",
    "Comorbidity count",
    "Diagnosis class",
    "Death, hospital death and diabetes labels",
    "Hospital time-to-event with event indicator"
  ),
  template = c(
    "pytorch_linear_regression",
    "pytorch_poisson",
    "pytorch_multiclass",
    "pytorch_multilabel",
    "pytorch_coxph"
  ),
  privacy_route = c(
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg and DP-SGD",
    "SecAgg only: Cox partial likelihood is risk-set based"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(task_map)
```

It is laid out in two parts: first the commands you would run to reproduce the
benchmark, then the recorded results of that run loaded from the committed
evidence artifacts. No live Opal servers are required to render this page; the
command chunks are shown for reference and are not executed.

## Running This Benchmark

The SUPPORT2 cohort is partitioned across three Opal nodes that are already set
up and serving the table `dsflower_demo.support2_cohort`. Each family is one
`ds.flower.fit()` call over the server-side partitions; only aggregate training
history and model parameters are returned to the researcher.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that hold the partitioned SUPPORT2 cohort.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.support2_cohort")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.support2_cohort")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.support2_cohort")
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = "D")
```

**2. Run a method family.** Train one template through dsFlower with FedAvg over
the server-side table. The continuous-regression family is shown here; the
standardized numeric SUPPORT2 columns are the features and the derived severity
score is the target.

```{r fit, eval=FALSE}
support2_features <- c(
  "age", "num_co", "scoma", "meanbp", "hrt",
  "resp", "temp", "crea", "sod", "wblc"
)

fit <- ds.flower.fit(
  conns,
  symbol   = "D",
  target   = "severity_score",
  features = support2_features,
  model    = "pytorch_linear_regression",
  strategy = "fedavg",
  privacy  = "clinical_default",
  rounds   = 3
)

DSI::datashield.logout(conns)
```

The other four families reuse the same `ds.flower.fit()` call, swapping the
model template and target: `pytorch_poisson` on `comorbidity_count`,
`pytorch_multiclass` on `diagnosis_class`, `pytorch_multilabel` on the
`c("label_death", "label_hospital_death", "label_diabetes")` labels, and
`pytorch_coxph` on the `c("time", "event")` survival pair, each over three
federated rounds. To run the DP-SGD subset, swap the profile to
`privacy = "high_sensitivity_dp"`; the four non-survival families remain
compatible, with multiclass using six rounds for the longer DP optimiser budget,
while CoxPH stays on its SecAgg route.

## Recorded Results

The values below come from the committed evidence artifacts for the runs
described above; no live servers are contacted when the article renders.

```{r load-evidence}
evidence_path <- system.file(
  "extdata", "dsflower_method_family_results.json",
  package = "dsFlowerClient"
)
evidence <- jsonlite::fromJSON(evidence_path)
dp_evidence_path <- system.file(
  "extdata", "dsflower_method_family_dp_results.json",
  package = "dsFlowerClient"
)
dp_evidence <- jsonlite::fromJSON(dp_evidence_path)
```

The benchmark uses the `r evidence$dataset_label` (`r evidence$n_total` rows)
under the `r evidence$privacy_profile` profile, with Secure Aggregation support
`r evidence$secagg_supported`. The rows are split evenly across three DataSHIELD
servers; the per-site training and event counts below are used by the survival
route disclosure guard.

```{r site-events}
site_summary <- data.frame(
  server = names(evidence$site_train_n),
  train_n = as.integer(unlist(evidence$site_train_n, use.names = FALSE)),
  event_n = as.integer(unlist(evidence$site_event_n, use.names = FALSE))
)
knitr::kable(site_summary)
```

Each row in the result table compares the centralized PyTorch loss against the
loss obtained by evaluating the federated model artefact saved by dsFlower after
the SecAgg run completed. The route is accepted only when all clients complete
without failures and the federated loss remains within the declared validation
envelope for that method family.

```{r results-table}
results <- evidence$results
knitr::kable(results[, c(
  "family", "method", "centralized_metric", "centralized_loss",
  "federated_loss", "delta_loss", "federated_n_failures",
  "validation_status"
)], digits = 4)
```

```{r dp-results-table}
dp_results <- dp_evidence$results
knitr::kable(dp_results[, c(
  "family", "method", "centralized_metric", "centralized_loss",
  "federated_loss", "delta_loss", "federated_n_failures",
  "validation_status"
)], digits = 4)
```

The DP-SGD table is defined over the SUPPORT2 routes whose losses decompose by
sample. CoxPH is validated under Secure Aggregation in the table above; its
partial-likelihood objective couples samples through risk sets, so its validated
privacy route is SecAgg rather than Opacus DP-SGD. The four non-survival
families shown here are the DP-SGD-compatible SUPPORT2 routes; all four complete
under the stricter profile and remain within their declared utility envelopes.
Multiclass classification uses six federated rounds in this profile because the
stricter DP-SGD optimiser required a slightly longer training budget than the
non-DP family run.

```{r compatibility-table}
compatibility <- data.frame(
  family = c(
    "Continuous regression",
    "Count regression",
    "Multiclass classification",
    "Multilabel classification",
    "Survival"
  ),
  clinical_default = c("pass", "pass", "pass", "pass", "pass"),
  high_sensitivity_dp = c("pass", "pass", "pass", "pass", "SecAgg route"),
  reason = c(
    "MSE is decomposable by row.",
    "Poisson NLL is decomposable by row.",
    "Cross entropy is decomposable by row.",
    "Binary cross entropy is decomposable by row.",
    "Cox partial likelihood is validated under SecAgg."
  ),
  stringsAsFactors = FALSE
)
knitr::kable(compatibility)
```

```{r loss-plot, fig.width=7, fig.height=4.2, fig.alt="Bar chart comparing centralized and federated SUPPORT2 losses across method families."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- results
  plot_data$label <- sub("^pytorch_", "", plot_data$method)
  plot_data$abs_delta <- abs(plot_data$delta_loss)
  ggplot2::ggplot(plot_data, ggplot2::aes(
    x = stats::reorder(label, abs_delta),
    y = abs_delta
  )) +
    ggplot2::geom_col(width = 0.68, fill = "#4D6A7F") +
    ggplot2::coord_flip() +
    ggplot2::labs(
      x = NULL,
      y = "Absolute federated-centralized loss delta",
      title = "SUPPORT2 method-family validation under clinical_default"
    ) +
    ggplot2::theme_minimal(base_size = 12)
}
```

```{r dp-loss-plot, fig.width=7, fig.height=3.8, fig.alt="Bar chart comparing DP-SGD federated-centralized SUPPORT2 losses for DP-compatible method families."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot_data <- dp_results
  plot_data$label <- sub("^pytorch_", "", plot_data$method)
  plot_data$abs_delta <- abs(plot_data$delta_loss)
  ggplot2::ggplot(plot_data, ggplot2::aes(
    x = stats::reorder(label, abs_delta),
    y = abs_delta
  )) +
    ggplot2::geom_col(width = 0.68, fill = "#6F5B3E") +
    ggplot2::coord_flip() +
    ggplot2::labs(
      x = NULL,
      y = "Absolute federated-centralized loss delta",
      title = "SUPPORT2 DP-SGD subset under high_sensitivity_dp"
    ) +
    ggplot2::theme_minimal(base_size = 12)
}
```
