---
title: "Radiomics Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Radiomics Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_VIGNETTE_DEMOS"), "true")
```

This vignette demonstrates downstream federated modeling on radiomics-derived
features. By default the script uses a deterministic LUNG1-style fixture with
feature names and signal structure similar to radiomics studies. When real
radiomics features are available, point the script at a CSV/RDS file:

```sh
export DSFLOWER_LUNG1_FEATURES="/path/to/radiomics_features.csv"
export DSFLOWER_LUNG1_TARGET="outcome"
Rscript inst/demos/benchmark_lung1_radiomics.R
```

The external file must contain an `id` column or one will be generated. The
target column is renamed internally to `outcome`, and all numeric columns other
than the target are used as model features.

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "benchmark_lung1_radiomics.R", package = "dsFlowerClient")
if (!nzchar(demo_file)) demo_file <- file.path("..", "inst", "demos", "benchmark_lung1_radiomics.R")
source(demo_file)
```

Validation run on 2026-05-10:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "features", "central_auc",
             "federated_auc", "central_accuracy", "federated_accuracy"),
  value = c(420, 316, 104, 3, 10, 0.7217, 0.7205, 0.6827, 0.6635)
)
```

This vignette covers the modeling stage. Image segmentation and radiomics
extraction should happen upstream in `dsImaging`; this benchmark consumes the
derived feature table as a federated learning input.
