---
title: "Demo Results Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo Results Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This page is the landing point for the rendered dsFlowerClient demos. It links
to each transparent workflow article and records the final validation numbers
that were available at handoff time. The demos are intended to show how the
researcher connects to Opal/DataSHIELD, stages or references server-side data,
launches Flower training, and checks the resulting federated model.

## Demo Results

```{r demo-results}
demo_results <- data.frame(
  demo = c(
    "Breast Cancer Benchmark",
    "Heart Disease Benchmark",
    "MedMNIST Benchmark",
    "Radiomics Benchmark",
    "Radiomics To Flower",
    "Direct Image ResNet",
    "Method Validation Suite",
    "Vision Method Validation"
  ),
  article = c(
    "demo-breast-cancer.html",
    "demo-heart-disease.html",
    "demo-medmnist.html",
    "demo-lung1-radiomics.html",
    "radiomics-to-flower.html",
    "direct-image-resnet.html",
    "validation-overview.html",
    "validation-vision-overview.html"
  ),
  data_path = c(
    "mlbench BreastCancer table",
    "UCI Cleveland or deterministic fallback",
    "BreastMNIST pooled image features",
    "LUNG1-style radiomics features",
    "dsImaging-derived LUNG1 radiomics table",
    "dsImaging direct NIfTI image resource",
    "Synthetic tabular/sequence/survival cohort",
    "Synthetic PNG image/mask cohort"
  ),
  sites = c(3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L),
  central_metric = c(
    "AUC 0.9878",
    "AUC 0.8919",
    "AUC 0.7654",
    "AUC 0.7217",
    "not applicable",
    "loss 0.5540",
    "17/17 methods pass",
    "3/3 methods pass"
  ),
  federated_metric = c(
    "AUC 0.9908",
    "AUC 0.8926",
    "AUC 0.7985",
    "AUC 0.7205",
    "loss 0.6503 -> 0.6474",
    "loss 0.5847",
    "17/17 acceptable",
    "3/3 acceptable"
  ),
  failures = c(0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L),
  stringsAsFactors = FALSE
)
knitr::kable(demo_results)
```

## Article Links

- [API Quickstart](api-quickstart.html)
- [Breast Cancer Benchmark](demo-breast-cancer.html)
- [Heart Disease Benchmark](demo-heart-disease.html)
- [MedMNIST Benchmark](demo-medmnist.html)
- [Radiomics Benchmark](demo-lung1-radiomics.html)
- [Radiomics To Flower](radiomics-to-flower.html)
- [Direct Image ResNet](direct-image-resnet.html)
- [All Method Validation Overview](validation-overview.html)
- [Vision Method Validation Overview](validation-vision-overview.html)

## Common Execution Shape

All live demos use the same DataSHIELD shape: build Opal logins, assign a
server-side symbol, call `ds.flower.fit()` or the lower-level
`connect -> prepare -> run` flow, then inspect metrics and cleanup state.

```{r common-shape, eval=FALSE}
library(dsFlowerClient)
library(DSI)
library(DSOpal)

builder <- DSI::newDSLoginBuilder()
builder$append(
  server = "opal1",
  url = "https://localhost:8443",
  user = Sys.getenv("OPAL_USER", "administrator"),
  password = Sys.getenv("OPAL_PASSWORD", "admin123"),
  table = "dsflower_demo.demo_site1",
  driver = "OpalDriver"
)
# Repeat builder$append(...) for opal2 and opal3.

conns <- DSI::datashield.login(
  logins = builder$build(),
  assign = TRUE,
  symbol = "D"
)

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = c("x1", "x2", "x3"),
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)

ds.flower.metrics(fit)
DSI::datashield.logout(conns)
```

For full method-by-method validation numbers, use the rendered validation
overview. It reads the committed JSON result files and links to every individual
method article.

