---
title: "Demo Results Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo Results Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This page is the landing point for the rendered dsFlowerClient demos. The
current evidence is split into four groups: clinical privacy-profile runs,
SUPPORT2 method-family runs, imaging handoff/direct-image runs and catalogue
coverage.

All live demos use the same DataSHIELD execution pattern: the researcher
connects to Opal, references or stages server-side data, launches Flower
training, and inspects approved model-scale outputs.

## Current Clinical And Privacy Results

```{r clinical-privacy-results}
clinical_privacy <- data.frame(
  profile = c("trusted_internal", "clinical_default",
              "clinical_update_noise XGBoost",
              "high_sensitivity_dp", "DP-SGD epsilon curves",
              "XGBoost update-noise curve",
              "clinical_default method families",
              "high_sensitivity_dp method families"),
  article = c(rep("clinical-privacy-benchmarks.html", 6),
              "clinical-method-family-benchmarks.html",
              "clinical-method-family-benchmarks.html"),
  data_path = c(
    "Breast Cancer, Heart Disease, Pima Diabetes and CDC Diabetes",
    "Breast Cancer, Pima Diabetes and CDC Diabetes",
    "CDC Diabetes",
    "CDC Diabetes",
    "CDC Diabetes",
    "CDC Diabetes",
    "SUPPORT2",
    "SUPPORT2"
  ),
  mechanism = c(
    "Controlled federated execution",
    "Secure Aggregation plus stricter metric/count policy",
    "Secure Aggregation plus histogram update noise",
    "Secure Aggregation plus Opacus DP-SGD",
    "Opacus DP-SGD at epsilon 2, 4 and 8 for two PyTorch templates",
    "Histogram update noise at epsilon 8, 12 and 16",
    "Secure Aggregation plus family-specific loss validation",
    "Secure Aggregation plus Opacus DP-SGD for the four compatible SUPPORT2 families"
  ),
  result = c(
    "6/6 runs pass, 0 client failures",
    "8/8 runs pass, 0 client failures",
    "1/1 run passes, 0 client failures",
    "2/2 runs pass, 0 client failures",
    "6/6 curve points pass, 0 client failures",
    "3/3 curve points pass, 0 client failures",
    "5/5 runs pass, 0 client failures",
    "4/4 DP-SGD-compatible runs pass, 0 client failures"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(clinical_privacy)
```

## Imaging Handoff Results

```{r imaging-handoff-results}
imaging_results <- data.frame(
  demo = c(
    "LUNG1 dsImaging Radiomics Evidence",
    "LUNG1 Direct dsImaging Image Evidence",
    "PathMNIST Direct Image ResNet Benchmark"
  ),
  article = c(
    "lung1-radiomics-to-flower.html",
    "lung1-direct-image-dsimaging-resnet.html",
    "pathmnist-direct-image-resnet.html"
  ),
  data_path = c(
    "dsImaging-derived LUNG1 radiomics table",
    "dsImaging-resolved LUNG1 NIfTI assets",
    "MedMNIST PathMNIST server-side PNG files"
  ),
  sites = c(3L, 3L, 3L),
  result = c(
    "422 rows; loss 0.6503 -> 0.6474; 0 client failures",
    "9 images; central loss 0.5540; federated loss 0.5847; 0 client failures",
    "1,500 images; central accuracy 0.9947; federated accuracy 0.9860; SecAgg branch pass"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(imaging_results)
```

## Catalogue Coverage

```{r template-validation-coverage}
validation_results <- data.frame(
  suite = c("Non-vision method validation", "Vision method validation"),
  article = c("validation-overview.html", "validation-vision-overview.html"),
  data_path = c(
    "Controlled tabular, sequence and survival cohort",
    "Controlled image and mask cohort"
  ),
  sites = c(3L, 3L),
  result = c("17/17 methods pass", "3/3 methods pass"),
  role = c("template catalogue coverage", "vision path coverage"),
  stringsAsFactors = FALSE
)
knitr::kable(validation_results)
```

## Article Links

- [API Quickstart](api-quickstart.html)
- [Clinical Privacy Benchmarks](clinical-privacy-benchmarks.html)
- [Clinical Method-Family Benchmarks](clinical-method-family-benchmarks.html)
- [PathMNIST Direct Image ResNet Benchmark](pathmnist-direct-image-resnet.html)
- [LUNG1 Direct dsImaging Image Evidence](lung1-direct-image-dsimaging-resnet.html)
- [LUNG1 dsImaging Radiomics Evidence](lung1-radiomics-to-flower.html)
- [Validation Evidence Overview](validation-overview.html)
- [Vision Method Validation Overview](validation-vision-overview.html)

## Common Execution Shape

All live demos use the same DataSHIELD shape: build Opal logins, assign a
server-side symbol, call `ds.flower.fit()` or the lower-level
`connect -> prepare -> run` flow, then inspect metrics and cleanup state.

```{r common-shape, eval=FALSE}
library(dsFlowerClient)
library(DSI)
library(DSOpal)

builder <- DSI::newDSLoginBuilder()
builder$append(
  server = "opal1",
  url = "https://localhost:8443",
  user = Sys.getenv("OPAL_USER", "administrator"),
  password = Sys.getenv("OPAL_PASSWORD", "admin123"),
  table = "dsflower_demo.demo_site1",
  driver = "OpalDriver"
)
# Repeat builder$append(...) for opal2 and opal3.

conns <- DSI::datashield.login(
  logins = builder$build(),
  assign = TRUE,
  symbol = "D"
)

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = c("x1", "x2", "x3"),
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)

ds.flower.metrics(fit)
DSI::datashield.logout(conns)
```

For full method-by-method validation numbers, use the rendered validation
overview. It reads the committed JSON result files and links to every individual
method article.
