---
title: "Breast Cancer Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Breast Cancer Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette demonstrates a federated logistic-regression benchmark on the
Wisconsin Breast Cancer dataset from `mlbench`. The script creates a local
central baseline, partitions the training rows across the configured Opal
servers, runs Flower through DataSHIELD, and writes a JSON/RDS audit record.

The default Opal configuration is declared from R so the same vignette can be
executed interactively or rendered in a live demo environment:

```{r opal-config, eval=FALSE}
opal_urls <- c(
  "https://localhost:8443",
  "https://localhost:8444",
  "https://localhost:8445"
)
opal_user <- Sys.getenv("OPAL_USER", "administrator")
opal_password <- Sys.getenv("OPAL_PASSWORD", "admin123")
```

## What The Script Does

The benchmark script is intentionally transparent. It loads the public dataset,
creates a binary outcome, partitions the training rows across the configured
sites, uploads one Opal table per site, logs into DataSHIELD, and launches a
federated logistic regression with `ds.flower.fit()`.

```{r workflow-shape, eval=FALSE}
data("BreastCancer", package = "mlbench")
raw <- BreastCancer[stats::complete.cases(BreastCancer), , drop = FALSE]

feature_cols <- setdiff(names(raw), c("Id", "Class"))
features <- paste0("bc_", seq_along(feature_cols))

dataset <- data.frame(
  id = paste0("bcw_", seq_len(nrow(raw))),
  setNames(lapply(raw[feature_cols], function(x) as.numeric(as.character(x))), features),
  outcome = as.integer(raw$Class == "malignant"),
  check.names = FALSE
)

run_dsflower_benchmark(
  data = dataset,
  features = features,
  demo_id = "breast_cancer_wisconsin",
  dataset_label = "Breast Cancer Wisconsin diagnostic benchmark",
  data_mode = "mlbench::BreastCancer",
  default_rounds = 2L
)
```

The helper expands to the same explicit DataSHIELD flow used in production
demos:

```{r opal-training-shape, eval=FALSE}
builder <- DSI::newDSLoginBuilder()
builder$append(
  server = "opal1",
  url = "https://localhost:8443",
  user = Sys.getenv("OPAL_USER", "administrator"),
  password = Sys.getenv("OPAL_PASSWORD", "admin123"),
  table = "dsflower_demo.demo_breast_cancer_wisconsin_site1",
  driver = "OpalDriver"
)
# opal2 and opal3 are appended in the same way.

conns <- DSI::datashield.login(
  logins = builder$build(),
  assign = TRUE,
  symbol = "D"
)

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = features,
  model = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)
```

Validation run on 2026-05-12:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "central_auc", "federated_auc",
             "central_accuracy", "federated_accuracy"),
  value = c(683, 513, 170, 3, 0.9878, 0.9908, 0.9471, 0.9647)
)
```

The expected output directory is `dsflower_output/demo_benchmarks/`, with
`summary.json` and `benchmark.rds` for each run. The Flower run must finish with
zero client failures and no active SuperNodes left behind.
