---
title: "Heart Disease Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Heart Disease Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_VIGNETTE_DEMOS"), "true")
```

This vignette demonstrates a clinical-tabular benchmark using the processed
Cleveland Heart Disease dataset from the UCI Machine Learning Repository. The
demo downloads the dataset, removes rows with missing values, builds a central
baseline, and trains the same feature set through DataSHIELD/Flower across the
configured Opal servers.

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"
Rscript inst/demos/benchmark_heart_disease.R
```

If the UCI URL is temporarily unavailable, the script falls back to a
deterministic heart-like fixture. The `data_mode` field in `summary.json`
records which path was used.

## What The Script Does

The demo first tries to download the processed Cleveland table, converts the
multi-valued disease field to a binary outcome, removes incomplete rows, and
uses all clinical covariates as features. If the download is unavailable, it
uses a deterministic fixture with the same column shape so that the DataSHIELD
and Flower path remains reproducible.

```{r data-shape, eval=FALSE}
dataset <- load_uci_cleveland()
features <- setdiff(names(dataset), c("id", "outcome"))

run_dsflower_benchmark(
  data = dataset,
  features = features,
  demo_id = "uci_heart_disease",
  dataset_label = "UCI Heart Disease federated benchmark",
  data_mode = "UCI Heart Disease processed Cleveland",
  default_rounds = 2L
)
```

The shared benchmark helper then performs the operational steps explicitly:

```{r operational-shape, eval=FALSE}
# 1. Stratified train/test split and feature standardization.
# 2. Stratified partition of train rows across opal1, opal2, opal3.
# 3. opalr::opal.table_save(...) uploads one table per site.
# 4. DSI::datashield.login(..., assign = TRUE, symbol = "D").
# 5. ds.flower.fit(...) launches the federated training run.

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = features,
  model = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)

post_caps <- DSI::datashield.aggregate(
  conns,
  as.symbol("flowerGetCapabilitiesDS")
)
```

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "benchmark_heart_disease.R", package = "dsFlowerClient")
if (!nzchar(demo_file)) demo_file <- file.path("..", "inst", "demos", "benchmark_heart_disease.R")
source(demo_file)
```

Validation run on 2026-05-10:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "central_auc", "federated_auc",
             "central_accuracy", "federated_accuracy"),
  value = c(297, 223, 74, 3, 0.8919, 0.8926, 0.8378, 0.8514)
)
```

The run passes only when Flower reports zero client failures and the post-run
capability check shows no active SuperNodes.
