---
title: "Breast Cancer Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Breast Cancer Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_VIGNETTE_DEMOS"), "true")
```

This vignette demonstrates a federated logistic-regression benchmark on the
Wisconsin Breast Cancer dataset from `mlbench`. The script creates a local
central baseline, partitions the training rows across the configured Opal
servers, runs Flower through DataSHIELD, and writes a JSON/RDS audit record.

The default Opal configuration is:

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"
Rscript inst/demos/benchmark_breast_cancer.R
```

The demo can also be launched while rendering this vignette:

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "benchmark_breast_cancer.R", package = "dsFlowerClient")
if (!nzchar(demo_file)) demo_file <- file.path("..", "inst", "demos", "benchmark_breast_cancer.R")
source(demo_file)
```

Validation run on 2026-05-10:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "central_auc", "federated_auc",
             "central_accuracy", "federated_accuracy"),
  value = c(683, 513, 170, 3, 0.9878, 0.9908, 0.9471, 0.9647)
)
```

The expected output directory is `dsflower_output/demo_benchmarks/`, with
`summary.json` and `benchmark.rds` for each run. The Flower run must finish with
zero client failures and no active SuperNodes left behind.
