---
title: "MedMNIST Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MedMNIST Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_VIGNETTE_DEMOS"), "true")
```

This vignette demonstrates that image-origin data can be turned into a compact
federated learning task without sending images to the client. The demo uses
BreastMNIST from MedMNIST, average-pools each 28x28 image to a 7x7 feature map,
and trains a binary logistic model through the same DataSHIELD/Flower path used
by the tabular demos.

The script bootstraps the Python client environment with `medmnist` when needed.

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"
Rscript inst/demos/benchmark_medmnist.R
```

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "benchmark_medmnist.R", package = "dsFlowerClient")
if (!nzchar(demo_file)) demo_file <- file.path("..", "inst", "demos", "benchmark_medmnist.R")
source(demo_file)
```

Validation run on 2026-05-10:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "features", "central_auc",
             "federated_auc", "central_accuracy", "federated_accuracy"),
  value = c(510, 383, 127, 3, 49, 0.7654, 0.7985, 0.6929, 0.7165)
)
```

This is not intended to replace a deep-learning imaging benchmark. It is a
small, reproducible proof that image-derived feature matrices can enter the
federated training path cleanly.
