---
title: "MedMNIST Federated Benchmark"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MedMNIST Federated Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_VIGNETTE_DEMOS"), "true")
```

This vignette demonstrates that image-origin data can be turned into a compact
federated learning task without sending images to the client. The demo uses
BreastMNIST from MedMNIST, average-pools each 28x28 image to a 7x7 feature map,
and trains a binary logistic model through the same DataSHIELD/Flower path used
by the tabular demos.

The script bootstraps the Python client environment with `medmnist` when needed.

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"
Rscript inst/demos/benchmark_medmnist.R
```

## What The Script Does

This is an image-origin benchmark, but the federated model receives a compact
feature table rather than raw pixels. The script downloads BreastMNIST through
Python, balances a small subset, average-pools every `28 x 28` image to a
`7 x 7` feature map, and writes 49 numeric columns named `px_00` to `px_48`.

```{r feature-shape, eval=FALSE}
dataset <- load_breastmnist()
features <- grep("^px_", names(dataset), value = TRUE)

run_dsflower_benchmark(
  data = dataset,
  features = features,
  demo_id = "medmnist_breastmnist",
  dataset_label = "MedMNIST BreastMNIST tabularized pixel benchmark",
  data_mode = "medmnist::BreastMNIST pooled 7x7 pixels",
  default_rounds = 2L
)
```

After tabularization, the DataSHIELD path is identical to the clinical-tabular
demos: each site receives a table, the client logs into all Opals, and Flower
trains a federated logistic model without sending the pooled image-derived table
back to the researcher.

```{r training-shape, eval=FALSE}
conns <- DSI::datashield.login(
  logins = builder$build(),
  assign = TRUE,
  symbol = "D"
)

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = features,
  model = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)
```

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "benchmark_medmnist.R", package = "dsFlowerClient")
if (!nzchar(demo_file)) demo_file <- file.path("..", "inst", "demos", "benchmark_medmnist.R")
source(demo_file)
```

Validation run on 2026-05-10:

```{r validation-table, echo=FALSE}
data.frame(
  metric = c("rows", "train", "test", "sites", "features", "central_auc",
             "federated_auc", "central_accuracy", "federated_accuracy"),
  value = c(510, 383, 127, 3, 49, 0.7654, 0.7985, 0.6929, 0.7165)
)
```

This is not intended to replace a deep-learning imaging benchmark. It is a
small, reproducible proof that image-derived feature matrices can enter the
federated training path cleanly.
