---
title: "Demo Results Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo Results Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This page is the landing point for the rendered dsFlowerClient demos. The main
evidence is split into three groups: public benchmark runs with centralised and
federated test metrics, LUNG1 imaging handoff runs that show composition with
dsImaging, and the template validation suite that exercises every authorised
method. Together they show the same DataSHIELD execution pattern: the
researcher connects to Opal, references or stages server-side data, launches
Flower training, and inspects approved model-scale outputs.

## Public Benchmark Results

```{r public-benchmark-results}
public_benchmarks <- data.frame(
  demo = c(
    "Breast Cancer Benchmark",
    "Heart Disease Benchmark",
    "MedMNIST Benchmark"
  ),
  article = c(
    "demo-breast-cancer.html",
    "demo-heart-disease.html",
    "demo-medmnist.html"
  ),
  data_path = c(
    "UCI Breast Cancer Wisconsin (Original)",
    "UCI Heart Disease processed Cleveland",
    "MedMNIST BreastMNIST pooled image features"
  ),
  sites = c(3L, 3L, 3L),
  central_metric = c(
    "AUC 0.9878",
    "AUC 0.8919",
    "AUC 0.7654"
  ),
  federated_metric = c(
    "AUC 0.9908",
    "AUC 0.8926",
    "AUC 0.7985"
  ),
  failures = c(0L, 0L, 0L),
  stringsAsFactors = FALSE
)
knitr::kable(public_benchmarks)
```

## Imaging Handoff Results

```{r imaging-handoff-results}
imaging_results <- data.frame(
  demo = c(
    "LUNG1 dsImaging Radiomics Evidence",
    "LUNG1 Direct dsImaging Image Evidence"
  ),
  article = c(
    "lung1-radiomics-to-flower.html",
    "lung1-direct-image-dsimaging-resnet.html"
  ),
  data_path = c(
    "dsImaging-derived LUNG1 radiomics table",
    "dsImaging-resolved LUNG1 NIfTI assets"
  ),
  sites = c(3L, 3L),
  result = c(
    "422 rows; loss 0.6503 -> 0.6474; 0 client failures",
    "9 images; central loss 0.5540; federated loss 0.5847; 0 client failures"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(imaging_results)
```

## Template Validation Coverage

```{r template-validation-coverage}
validation_results <- data.frame(
  suite = c("Non-vision method validation", "Vision method validation"),
  article = c("validation-overview.html", "validation-vision-overview.html"),
  data_path = c(
    "Controlled tabular, sequence and survival cohort",
    "Controlled image and mask cohort"
  ),
  sites = c(3L, 3L),
  result = c("17/17 methods pass", "3/3 methods pass"),
  stringsAsFactors = FALSE
)
knitr::kable(validation_results)
```

## Article Links

- [API Quickstart](api-quickstart.html)
- [Breast Cancer Benchmark](demo-breast-cancer.html)
- [Heart Disease Benchmark](demo-heart-disease.html)
- [MedMNIST Benchmark](demo-medmnist.html)
- [LUNG1 Direct dsImaging Image Evidence](lung1-direct-image-dsimaging-resnet.html)
- [LUNG1 dsImaging Radiomics Evidence](lung1-radiomics-to-flower.html)
- [All Method Validation Overview](validation-overview.html)
- [Vision Method Validation Overview](validation-vision-overview.html)

The repository also keeps self-contained smoke demos for development
(`demo-lung1-radiomics.html`, `radiomics-to-flower.html` and
`direct-image-resnet.html`). They are useful for checking the mechanics without
external data, but the thesis evidence above uses public datasets or dsImaging
assets.

## Common Execution Shape

All live demos use the same DataSHIELD shape: build Opal logins, assign a
server-side symbol, call `ds.flower.fit()` or the lower-level
`connect -> prepare -> run` flow, then inspect metrics and cleanup state.

```{r common-shape, eval=FALSE}
library(dsFlowerClient)
library(DSI)
library(DSOpal)

builder <- DSI::newDSLoginBuilder()
builder$append(
  server = "opal1",
  url = "https://localhost:8443",
  user = Sys.getenv("OPAL_USER", "administrator"),
  password = Sys.getenv("OPAL_PASSWORD", "admin123"),
  table = "dsflower_demo.demo_site1",
  driver = "OpalDriver"
)
# Repeat builder$append(...) for opal2 and opal3.

conns <- DSI::datashield.login(
  logins = builder$build(),
  assign = TRUE,
  symbol = "D"
)

fit <- ds.flower.fit(
  conns,
  symbol = "D",
  target = "outcome",
  features = c("x1", "x2", "x3"),
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)

ds.flower.metrics(fit)
DSI::datashield.logout(conns)
```

For full method-by-method validation numbers, use the rendered validation
overview. It reads the committed JSON result files and links to every individual
method article.
