---
title: "Direct Image ResNet Smoke Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Direct Image ResNet Smoke Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_IMAGE_VIGNETTE"), "true")
```

This vignette covers a small direct-image workflow. Unlike the radiomics and
MedMNIST-tabular demos, this one trains a PyTorch vision template from image
files stored on the server side. The Opal table contains only sample metadata:
`id`, `relative_path`, and `label`.

The demo is intentionally small and uses the `sandbox_open` profile by default.
Clinical vision profiles require substantially larger cohorts and Secure
Aggregation-capable Flower runtimes.

## Layout

Each Rock sees the same configured image root:

```text
/tmp/dsflower_direct_image_smoke/
  site1/
    img_001.png
    img_002.png
  site2/
    img_001.png
  site3/
    img_001.png
```

Each Opal receives a site-specific metadata table:

```text
id,relative_path,label
site1_img_001,site1/img_001.png,0
site1_img_002,site1/img_002.png,1
```

`dsFlower` stages the metadata, writes an image manifest, and the
`pytorch_resnet18` ClientApp loads the PNG files inside the Rock.

## Run

For the local three-Opal Docker setup:

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"

Rscript inst/demos/direct_image_resnet_smoke.R
```

The script generates synthetic PNGs, copies each site folder into the matching
Rock container, sets `dsflower.image_data_root`, uploads metadata tables, and
runs one federated ResNet-18 round.

If the images are already available inside each Rock, skip the Docker copy:

```sh
export DSFLOWER_IMAGE_DEMO_SKIP_DOCKER_COPY=true
export DSFLOWER_IMAGE_DEMO_SERVER_ROOT="/path/visible/in/each/rock"
Rscript inst/demos/direct_image_resnet_smoke.R
```

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "direct_image_resnet_smoke.R",
                         package = "dsFlowerClient")
if (!nzchar(demo_file)) {
  demo_file <- file.path("..", "inst", "demos",
                         "direct_image_resnet_smoke.R")
}
source(demo_file)
```

## API Shape

The core training spec is:

```{r api-shape, eval=FALSE}
recipe <- ds.flower.recipe(
  model = ds.flower.model.pytorch_resnet18(
    n_classes = 2L,
    batch_size = 4L,
    local_epochs = 1L
  ),
  strategy = ds.flower.strategy.fedavg(),
  privacy = ds.flower.privacy.sandbox_open(),
  target = "label",
  num_rounds = 1L
)
```

The current direct-image table path uses the low-level prepare call so the
server receives `data_type = "image"` and stages an image manifest:

```{r prepare-shape, eval=FALSE}
ds.flower.nodes.prepare(
  conns,
  symbol = "flower",
  target_column = "label",
  run_config = c(recipe$model$params, list(data_type = "image")),
  privacy = recipe$privacy,
  template_name = recipe$model$template
)
```

The higher-level `ds.flower.fit(resource = ..., model = "resnet18")` path is
intended for `dsImaging` resources whose descriptors already declare
`source_kind = "image_bundle"`.

## dsImaging Resource Handoff

When images are already published as a `dsImaging` resource, use the direct
resource demo instead of copying files into each Rock manually:

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export DSFLOWER_IMAGING_RESOURCE="dsdemo.lung1_study"
export DSFLOWER_IMAGING_TARGET="os_2yr_alive"
Rscript inst/demos/dsimaging_direct_image_resnet.R
```

That path calls `ds.flower.connect(resource = ...)`, so the server initializes
`dsImaging`, carries the imaging descriptor into `dsFlower`, stages the
metadata, resolves image paths from `sample_manifests`, and trains directly
from the image assets in the imaging store or bucket.
