---
title: "Direct Image ResNet Validation Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Direct Image ResNet Validation Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_demo <- identical(Sys.getenv("DSFLOWER_RUN_IMAGE_VIGNETTE"), "true")
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

This vignette covers a small direct-image validation workflow. Unlike the
radiomics and MedMNIST-tabular demos, this one trains a PyTorch vision template
from image files stored on the server side. The Opal table contains only sample
metadata; image pixels remain in the Rock-side imaging store or bucket.

The demo is intentionally small and uses the `sandbox_open` profile by default.
Clinical vision profiles require substantially larger cohorts and Secure
Aggregation-capable Flower runtimes.

## Layout

Each Rock sees the same configured image root:

```text
/tmp/dsflower_direct_image_smoke/
  site1/
    img_001.png
    img_002.png
  site2/
    img_001.png
  site3/
    img_001.png
```

Each Opal receives a site-specific metadata table:

```text
id,relative_path,label
site1_img_001,site1/img_001.png,0
site1_img_002,site1/img_002.png,1
```

`dsFlower` stages the metadata, writes an image manifest, and the
`pytorch_resnet18` ClientApp loads the PNG files inside the Rock.

## Run

For the local three-Opal Docker setup:

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"

Rscript inst/demos/direct_image_resnet_smoke.R
```

The script generates synthetic PNGs, copies each site folder into the matching
Rock container, sets `dsflower.image_data_root`, uploads metadata tables, and
runs one federated ResNet-18 round.

If the images are already available inside each Rock, skip the Docker copy:

```sh
export DSFLOWER_IMAGE_DEMO_SKIP_DOCKER_COPY=true
export DSFLOWER_IMAGE_DEMO_SERVER_ROOT="/path/visible/in/each/rock"
Rscript inst/demos/direct_image_resnet_smoke.R
```

```{r run-demo, eval=run_demo}
demo_file <- system.file("demos", "direct_image_resnet_smoke.R",
                         package = "dsFlowerClient")
if (!nzchar(demo_file)) {
  demo_file <- file.path("..", "inst", "demos",
                         "direct_image_resnet_smoke.R")
}
source(demo_file)
```

## API Shape

The core training spec is:

```{r api-shape, eval=FALSE}
recipe <- ds.flower.recipe(
  model = ds.flower.model.pytorch_resnet18(
    n_classes = 2L,
    batch_size = 4L,
    local_epochs = 1L
  ),
  strategy = ds.flower.strategy.fedavg(),
  privacy = ds.flower.privacy.sandbox_open(),
  target = "label",
  num_rounds = 1L
)
```

The current direct-image table path uses the low-level prepare call so the
server receives `data_type = "image"` and stages an image manifest:

```{r prepare-shape, eval=FALSE}
ds.flower.nodes.prepare(
  conns,
  symbol = "flower",
  target_column = "label",
  run_config = c(recipe$model$params, list(data_type = "image")),
  privacy = recipe$privacy,
  template_name = recipe$model$template
)
```

The higher-level `ds.flower.fit(resource = ..., model = "resnet18")` path is
intended for `dsImaging` resources whose descriptors already declare
`source_kind = "image_bundle"`.

## dsImaging Resource Handoff

When images are already published as a `dsImaging` resource, use the direct
resource demo instead of copying files into each Rock manually:

```sh
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export DSFLOWER_IMAGING_RESOURCE="dsdemo.lung1_study"
export DSFLOWER_IMAGING_TARGET="os_2yr_alive"
Rscript inst/demos/dsimaging_direct_image_resnet.R
```

That path calls `ds.flower.connect(resource = ...)`, so the server initializes
`dsImaging`, carries the imaging descriptor into `dsFlower`, stages the
metadata, resolves image paths from `sample_manifests`, and trains directly
from the image assets in the imaging store or bucket.

Set `DSFLOWER_IMAGING_LOCAL_WORKDIR` to a local copy of the same demo study to
also run a centralized PyTorch baseline over the same NIfTI images:

```sh
export DSFLOWER_IMAGING_LOCAL_WORKDIR="/tmp/dsimaging_lung1_study"
export DSFLOWER_IMAGING_REQUIRE_LOCAL_BASELINE=true
Rscript inst/demos/dsimaging_direct_image_resnet.R
```

The script writes:

```text
summary.json
run.rds
imaging_metadata.rds
imaging_assets.rds
post_capabilities.rds
local_baseline.json
local_baseline.log
local_vs_federated.json
```

## Local vs Federated Validation

The validation run below was executed on 2026-05-11 against three local
Opal/Rock nodes using the `dsdemo.lung1_study` imaging resource. Each node held
three LUNG1 cases; the centralized baseline used the same nine NIfTI images from
the local dsImaging work directory. This is a path validation, not a clinical
performance claim: the cohort is deliberately tiny and the evaluation is on the
same training cohort.

```{r validation-table}
validation <- data.frame(
  run = c("centralized local", "federated dsImaging"),
  engine = c("PyTorch ResNet18", "DataSHIELD + Flower ResNet18"),
  samples = c(9L, 9L),
  sites_or_clients = c(1L, 3L),
  rounds_or_epochs = c(1L, 1L),
  loss = c(0.5540125, 0.5847445),
  accuracy = c(0.7777778, NA_real_),
  failures = c(NA_integer_, 0L)
)
knitr::kable(validation, digits = 4)
```

```{r validation-loss-plot, fig.width=7, fig.height=4, fig.alt="Bar chart comparing centralized and federated ResNet18 loss on the same direct imaging demo cohort."}
plot_data <- data.frame(
  run = factor(c("centralized local", "federated dsImaging"),
               levels = c("centralized local", "federated dsImaging")),
  loss = c(0.5540125, 0.5847445)
)
if (has_ggplot2) {
  ggplot2::ggplot(plot_data, ggplot2::aes(x = run, y = loss, fill = run)) +
    ggplot2::geom_col(width = 0.62, show.legend = FALSE) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.4f", loss)),
                       vjust = -0.35, size = 3.5) +
    ggplot2::scale_fill_manual(values = c("#4C78A8", "#F58518")) +
    ggplot2::coord_cartesian(ylim = c(0, 0.7)) +
    ggplot2::labs(x = NULL, y = "Loss",
                  title = "Centralized vs federated direct-image run") +
    ggplot2::theme_minimal(base_size = 11)
} else {
  barplot(plot_data$loss, names.arg = as.character(plot_data$run),
          ylim = c(0, 0.7), ylab = "Loss",
          main = "Centralized vs federated direct-image run",
          col = c("#4C78A8", "#F58518"))
}
```

The federated run completed one round with three clients, zero client failures,
and no active SuperNodes left after cleanup. The resulting comparison file was:

```text
dsflower_output/direct_dsimaging_images/20260511_140204/local_vs_federated.json
```

The key output was:

```json
{
  "federated": {
    "rounds": 1,
    "clients": 3,
    "loss": 0.5847444699870216,
    "n_failures": 0
  },
  "centralized": {
    "epochs": 1,
    "n_samples": 9,
    "class_counts": {"0": 7, "1": 2},
    "device": "cpu",
    "loss": 0.5540124575297037,
    "accuracy": 0.7777777777777778
  },
  "delta": {
    "loss_federated_minus_centralized": 0.03073201245731793
  }
}
```
