---
title: "Composing Experiments with Recipes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Composing Experiments with Recipes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
quiet <- function(expr) invisible(capture.output(suppressMessages(expr)))
```

## The recipe system

A recipe in dsFlowerClient is a composition of four building blocks:

- **Task**: classification or regression
- **Model**: the learning algorithm and its hyperparameters
- **Strategy**: how the server aggregates updates from the nodes
- **Privacy**: whether to apply differential privacy

Each block is an independent S3 object. You can mix and match them freely.
The recipe itself is just a specification. Nothing runs until you call
`ds.flower.run.start()`.

```{r load}
library(dsFlowerClient)
```

## Available models

### Logistic Regression

A solid default for binary classification. Converges fast in federated
settings because the model is small (one weight per feature).

```{r logreg}
model_lr <- ds.flower.model.sklearn_logreg(
  penalty  = "l2",
  C        = 1.0,
  max_iter = 200L
)
model_lr
```

### SGD Classifier

Supports several loss functions and learning rate schedules. The
`"log_loss"` loss function gives logistic regression with SGD optimization.

```{r sgd}
model_sgd <- ds.flower.model.sklearn_sgd(
  loss        = "log_loss",
  alpha       = 0.0001,
  lr_schedule = "optimal"
)
model_sgd
```

### Ridge Classifier

L2-regularized linear classifier. Good for high-dimensional data.

```{r ridge}
model_ridge <- ds.flower.model.sklearn_ridge(alpha = 1.0)
model_ridge
```

### PyTorch MLP

For problems where a linear model is not expressive enough:

```{r mlp}
model_mlp <- ds.flower.model.pytorch_mlp(
  hidden_layers = c(128L, 64L),
  learning_rate = 0.001,
  batch_size    = 32L,
  local_epochs  = 2L
)
model_mlp
```

`local_epochs` controls how many passes over the local data each node
makes per round. More epochs can speed up convergence but risk local model
drift (especially with non-IID data).

## Aggregation strategies

### FedAvg

The standard algorithm. Each node trains, sends updated weights, and the
server computes a weighted average proportional to dataset size.

```{r fedavg}
strat_avg <- ds.flower.strategy.fedavg(
  fraction_fit          = 1.0,
  fraction_evaluate     = 1.0,
  min_fit_clients       = 2L,
  min_available_clients = 2L
)
strat_avg
```

### FedProx

Adds a proximal term to keep local models closer to the global model.
Helps with heterogeneous (non-IID) data.

```{r fedprox}
strat_prox <- ds.flower.strategy.fedprox(
  proximal_mu           = 0.1,
  min_fit_clients       = 2L,
  min_available_clients = 2L
)
strat_prox
```

## Privacy settings

### Research mode

No privacy enhancements. Suitable for trusted environments.

```{r privacy-research}
priv <- ds.flower.privacy.research()
priv
```

### Differential privacy

Clips gradient norms and adds calibrated noise. Smaller epsilon = stronger
privacy but noisier model.

```{r privacy-dp}
priv_dp <- ds.flower.privacy.dp(
  epsilon       = 1.0,
  delta         = 1e-5,
  clipping_norm = 1.0
)
priv_dp
```

## Putting it together

```{r recipe}
recipe_cls <- ds.flower.recipe(
  task            = ds.flower.task.classification(),
  model           = model_lr,
  strategy        = strat_avg,
  privacy         = ds.flower.privacy.research(),
  num_rounds      = 10L,
  target_column   = "diagnosis",
  feature_columns = c("age", "bmi", "glucose", "insulin")
)
recipe_cls
```

Regression works the same way:

```{r regression}
recipe_reg <- ds.flower.recipe(
  task            = ds.flower.task.regression(),
  model           = ds.flower.model.sklearn_sgd(loss = "squared_error"),
  strategy        = strat_avg,
  num_rounds      = 10L,
  target_column   = "price",
  feature_columns = c("sqft", "bedrooms", "bathrooms")
)
recipe_reg
```

## Live demo: training different models across 3 sites

Let's run real federated training with different model configurations
against three Opal servers.

```{r connect-recipes, cache = FALSE}
library(DSI)
library(DSOpal)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "site_a", url = "https://localhost:8443",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")
builder$append(server = "site_b", url = "https://localhost:8444",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")
builder$append(server = "site_c", url = "https://localhost:8445",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")

conns <- DSI::datashield.login(logins = builder$build(), assign = FALSE)
cat("Connected to:", paste(names(conns), collapse = ", "), "\n")
```

### Experiment 1: Logistic Regression with FedAvg

```{r exp1-logreg}
quiet(ds.flower.nodes.init(conns, resource = "dsflower_test.flower_node"))
quiet(ds.flower.nodes.prepare(conns, target_column = "target",
                               feature_columns = c("f1", "f2", "f3", "f4", "f5")))
quiet(ds.flower.superlink.start())
Sys.sleep(2)
quiet(ds.flower.nodes.ensure(conns, symbol = "flower"))
Sys.sleep(10)

recipe_exp1 <- ds.flower.recipe(
  task     = ds.flower.task.classification(),
  model    = ds.flower.model.sklearn_logreg(C = 1.0, max_iter = 200L),
  strategy = ds.flower.strategy.fedavg(
    min_fit_clients = 3L, min_available_clients = 3L),
  num_rounds      = 5L,
  target_column   = "target",
  feature_columns = c("f1", "f2", "f3", "f4", "f5")
)

cat("=== Logistic Regression + FedAvg (5 rounds) ===\n\n")
run_exp1 <- ds.flower.run.start(recipe_exp1, verbose = TRUE)
cat(sprintf("\nExit status: %s\n", run_exp1$status))

quiet(ds.flower.nodes.cleanup(conns, symbol = "flower"))
quiet(ds.flower.superlink.stop())
```

### Experiment 2: SGD Classifier with FedAvg

```{r exp2-sgd}
quiet(ds.flower.nodes.init(conns, resource = "dsflower_test.flower_node"))
quiet(ds.flower.nodes.prepare(conns, target_column = "target",
                               feature_columns = c("f1", "f2", "f3", "f4", "f5")))
quiet(ds.flower.superlink.start())
Sys.sleep(2)
quiet(ds.flower.nodes.ensure(conns, symbol = "flower"))
Sys.sleep(10)

recipe_exp2 <- ds.flower.recipe(
  task     = ds.flower.task.classification(),
  model    = ds.flower.model.sklearn_sgd(loss = "log_loss", alpha = 0.0001),
  strategy = ds.flower.strategy.fedavg(
    min_fit_clients = 3L, min_available_clients = 3L),
  num_rounds      = 5L,
  target_column   = "target",
  feature_columns = c("f1", "f2", "f3", "f4", "f5")
)

cat("=== SGD Classifier + FedAvg (5 rounds) ===\n\n")
run_exp2 <- ds.flower.run.start(recipe_exp2, verbose = TRUE)
cat(sprintf("\nExit status: %s\n", run_exp2$status))

quiet(ds.flower.nodes.cleanup(conns, symbol = "flower"))
quiet(ds.flower.superlink.stop())
```

### Experiment 3: Ridge Classifier with FedAvg

```{r exp3-ridge}
quiet(ds.flower.nodes.init(conns, resource = "dsflower_test.flower_node"))
quiet(ds.flower.nodes.prepare(conns, target_column = "target",
                               feature_columns = c("f1", "f2", "f3", "f4", "f5")))
quiet(ds.flower.superlink.start())
Sys.sleep(2)
quiet(ds.flower.nodes.ensure(conns, symbol = "flower"))
Sys.sleep(10)

recipe_exp3 <- ds.flower.recipe(
  task     = ds.flower.task.classification(),
  model    = ds.flower.model.sklearn_ridge(alpha = 1.0),
  strategy = ds.flower.strategy.fedavg(
    min_fit_clients = 3L, min_available_clients = 3L),
  num_rounds      = 5L,
  target_column   = "target",
  feature_columns = c("f1", "f2", "f3", "f4", "f5")
)

cat("=== Ridge Classifier + FedAvg (5 rounds) ===\n\n")
run_exp3 <- ds.flower.run.start(recipe_exp3, verbose = TRUE)
cat(sprintf("\nExit status: %s\n", run_exp3$status))

quiet(ds.flower.nodes.cleanup(conns, symbol = "flower"))
quiet(ds.flower.superlink.stop())
```

## Under the hood: from recipe to Flower App

When you call `ds.flower.run.start(recipe)`, the recipe is converted into
a Flower App:

1. `.build_flower_app()` copies a Python template from
   `inst/flower_templates/<model_template>/` containing `client_app.py`
   (local training) and `server_app.py` (aggregation strategy).

2. `.write_pyproject_toml()` serializes all recipe parameters into TOML:
   ```
   [tool.flwr.app.config]
   num-server-rounds = 5
   task-type = "classification"
   C = 1.0
   strategy = "FedAvg"
   ```

3. `flwr run` installs the app, starts a ServerApp, and orchestrates
   training through the SuperLink.

## Server-side constraints

| Option | Default | Effect |
|:-------|:--------|:-------|
| `nfilter.subset` | 3 | Minimum rows per node |
| `dsflower.max_rounds` | 500 | Maximum training rounds |
| `dsflower.allow_custom_config` | FALSE | Custom run configs? |
| `dsflower.allow_supernode_spawn` | TRUE | Can SuperNodes start? |
| `dsflower.max_concurrent_runs` | 5 | Simultaneous runs |

```{r final-cleanup-recipes, include = FALSE, cache = FALSE}
DSI::datashield.logout(conns)
```
