---
title: "Federated Learning with Medical Images"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Federated Learning with Medical Images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Connect to DataSHIELD

```{r}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

conns <- DSI::datashield.login(
  logins = data.frame(
    server   = "hospital_a",
    url      = "https://opal.hospital-a.org",
    user     = "researcher",
    password = "****",
    driver   = "OpalDriver",
    profile  = "default",
    stringsAsFactors = FALSE
  ),
  assign = FALSE
)
```

## Assign imaging resource

The admin has already configured the imaging resource in Opal pointing
to the S3/MinIO collection. The researcher just references it by name.

```{r}
DSI::datashield.assign.resource(conns,
  symbol   = "img_res",
  resource = list(hospital_a = "IMAGING.brain_mri_collection"))
```

## Initialize imaging and federated learning handles

```{r}
# dsImaging resolves the S3 resource -> downloads manifest
DSI::datashield.assign.expr(conns, "img",
  expr = quote(imagingInitDS("img_res")))

# dsFlower creates a federated learning handle from the imaging data
DSI::datashield.assign.expr(conns, "flower",
  expr = quote(flowerInitDS("img")))
```

## Prepare training

Images are downloaded from S3 to the server's local staging area.
The researcher never sees the images -- only the server handles them.

```{r}
privacy <- ds.flower.privacy.clinical_default()

ds.flower.nodes.prepare(conns, "flower",
  target_column = "label",
  privacy       = privacy,
  template_name = "pytorch_resnet18")
```

## Start SuperLink and connect SuperNodes

```{r}
# SuperLink on the researcher's machine (TLS auto-configured)
ds.flower.superlink.start()

# SuperNodes on each hospital server connect to the SuperLink
ds.flower.nodes.ensure(conns, "flower",
  template_name = "pytorch_resnet18")
```

## Define recipe and train

```{r}
recipe <- ds.flower.recipe(
  task     = ds.flower.task.classification(),
  model    = ds.flower.model.pytorch_resnet18(n_classes = 2),
  strategy = ds.flower.strategy.fedavg(),
  privacy  = privacy,
  num_rounds      = 10,
  target_column   = "label",
  feature_columns = character(0)
)

run <- ds.flower.run.start(recipe, conns)
print(run)
```

## Use the model

```{r}
# List saved models
ds.flower.models()

# Load and predict (uses Python via the client venv)
ds.flower.predict(run$output_dir, new_images_df, type = "prob")

# Save to a specific location
ds.flower.save_model(run, "my_model.rds")
```

## Cleanup

```{r}
ds.flower.superlink.stop()
ds.flower.nodes.cleanup(conns, "flower")
DSI::datashield.logout(conns)
```

## With segmentation masks (U-Net)

If masks were generated by dsRadiomics, pass the generation ID:

```{r}
ds.flower.nodes.prepare(conns, "flower",
  target_column = "mask_path",
  privacy       = privacy,
  template_name = "pytorch_unet2d",
  config        = list(segmentation_generation_id = generation_id))

recipe <- ds.flower.recipe(
  task     = ds.flower.task.segmentation(),
  model    = ds.flower.model.pytorch_unet2d(n_classes = 1),
  strategy = ds.flower.strategy.fedbn(),
  privacy  = privacy,
  num_rounds      = 20,
  target_column   = "mask_path",
  feature_columns = character(0)
)
```

## Detached mode (long training)

```{r}
# Start SuperLink as daemon (survives R session exit)
ds.flower.superlink.start(detached = TRUE)

# ... start training, close laptop ...

# Next day: reconnect
ds.flower.superlink.attach()
ds.flower.superlink.status()
```
