---
title: "LUNG1 dsImaging Radiomics To dsFlower Evidence"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{LUNG1 dsImaging Radiomics To dsFlower Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  file.path("inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  system.file("extdata", "dsflower_lung1_radiomics_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
evidence <- jsonlite::fromJSON(artifact_path, simplifyVector = FALSE)
```

This article records the real LUNG1 radiomics handoff run used as thesis
evidence. The rendered article uses the recorded output of the live May 2026
workflow so that package checks do not require active Opal servers. The live
workflow is implemented in `inst/demos/lung1_radiomics_to_flower.R`.

The validation path is:

1. `dsImaging` resolves the LUNG1 imaging resource and radiomics assets.
2. `ds.imaging.radiomics.load_features()` loads each site's derived feature
   table into a server-side DataSHIELD symbol.
3. `dsFlowerClient` trains `sklearn_logreg` with FedAvg over the server-side
   tables.
4. Only aggregate training history and model parameters are returned.

## Tutorial: Running The dsImaging To dsFlower Handoff

The commands below are the live workflow from
`inst/demos/lung1_radiomics_to_flower.R`. They show how the dsImaging calls
feed server-side radiomic feature tables straight into a federated dsFlower
fit, without images, masks or row-level features ever leaving the nodes. This
article first walks through these commands, then loads the recorded results
of the run below. The chunks are shown for reference and are not executed when
the article renders.

**1. Configure the nodes and the imaging resource.** Point at the three
Opal/Rock nodes and the LUNG1 study resource registered in each.

```{r tutorial-config, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsImagingClient)
library(dsFlowerClient)

urls <- trimws(strsplit(Sys.getenv(
  "DSFLOWER_OPAL_URLS",
  "https://localhost:8443,https://localhost:8444,https://localhost:8445"
), ",", fixed = TRUE)[[1]])
users <- rep(Sys.getenv("OPAL_USER", "administrator"), length(urls))
passwords <- rep(Sys.getenv("OPAL_PASSWORD", "admin123"), length(urls))

logins <- data.frame(
  server = paste0("opal", seq_along(urls)),
  url = urls, user = users, password = passwords,
  driver = "OpalDriver",
  options = "list(ssl_verifyhost=0L, ssl_verifypeer=0L)",
  stringsAsFactors = FALSE
)
conns <- DSI::datashield.login(logins = logins, assign = FALSE)

resource <- paste0(Sys.getenv("LUNG1_OPAL_PROJECT", "dsdemo"), ".",
                   Sys.getenv("LUNG1_OPAL_RESOURCE", "lung1_full_study"))
```

**2. Initialise dsImaging and load radiomic features server-side.**
`ds.imaging.init()` binds the imaging resource; then, per site,
`ds.imaging.radiomics.load_features()` materialises the derived feature table
into the server-side symbol `rad`. The images and masks never move.

```{r tutorial-dsimaging, eval=FALSE}
ds.imaging.init(conns, resource, symbol = "img")

radiomics_result <- readRDS(
  file.path(Sys.getenv("LUNG1_WORKDIR", "/tmp/dsimaging_lung1_full"),
            "datashield_radiomics_result.rds")
)
for (srv in names(radiomics_result)) {
  item <- radiomics_result[[srv]]
  ds.imaging.radiomics.load_features(
    conns[srv],
    dataset_id = item$dataset_id,
    asset_id = item$asset_id,
    symbol = "rad",
    include_metadata = TRUE,
    syntactic_names = TRUE
  )
}
```

**3. Train the federated model on the derived features.** Hand the
server-side `rad` table to `ds.flower.fit()` with logistic regression and
FedAvg. Only aggregate training history and model parameters come back.

```{r tutorial-fit, eval=FALSE}
features <- c(
  "original_firstorder_Energy",
  "original_shape_Compactness1",
  "original_glrlm_RunLengthNonUniformity",
  "wavelet.HLH_glrlm_RunLengthNonUniformity",
  "age",
  "gender_male"
)

fit <- ds.flower.fit(
  conns,
  symbol = "rad",
  target = "os_2yr_alive",
  features = features,
  model = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy = "fedavg",
  privacy = Sys.getenv("DSFLOWER_RADFLOWER_PRIVACY", "trusted_internal"),
  rounds = as.integer(Sys.getenv("DSFLOWER_RADFLOWER_ROUNDS", "2"))
)
DSI::datashield.logout(conns)
```

## Recorded Results

```{r recorded-run}
cat("Source script: inst/demos/lung1_radiomics_to_flower.R\n")
cat("Recorded run_id:", evidence$run_id, "\n")
```

## Server-Side Cohort

```{r cohort}
site_dim_names <- grep("opal", names(evidence$dimensions), value = TRUE)
site_dims <- do.call(rbind, lapply(site_dim_names, function(name) {
  dims <- unlist(evidence$dimensions[[name]])
  data.frame(
    site = sub("dimensions of rad in ", "", name, fixed = TRUE),
    rows = dims[[1]],
    columns = dims[[2]]
  )
}))
combined_dims <- unlist(evidence$dimensions[["dimensions of rad in combined studies"]])
knitr::kable(site_dims)

data.frame(
  resource = evidence$resource,
  target = evidence$target,
  combined_rows = combined_dims[[1]],
  combined_columns = combined_dims[[2]],
  stringsAsFactors = FALSE
)
```

## Features

```{r features}
feature_roles <- c(rep("radiomic", 4), "metadata", "metadata")
feature_table <- data.frame(
  feature = unlist(evidence$features),
  role = feature_roles,
  stringsAsFactors = FALSE
)
knitr::kable(feature_table)
```

## Federated Training

```{r training}
training <- data.frame(
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy_profile = evidence$privacy,
  rounds = evidence$rounds,
  max_iter = evidence$max_iter,
  stringsAsFactors = FALSE
)
knitr::kable(training)
```

```{r federated-output}
history <- do.call(rbind, lapply(evidence$history, function(x) {
  data.frame(
    round = x$round,
    loss = x$loss,
    n_clients = x$n_clients,
    n_failures = x$n_failures
  )
}))
knitr::kable(history, digits = 4)

cat("[dsFlower] run_id:", evidence$run_id, "\n")
cat("[dsFlower] model_id:", evidence$model_id, "\n")
cat("[dsFlower] output_dir:", evidence$flower_output_dir, "\n")
cat("[acceptance] final loss:", sprintf("%.4f", tail(history$loss, 1L)), "\n")
cat("[acceptance] total client failures:",
    sum(as.integer(history$n_failures)), "\n")
cat("[acceptance] PASS:",
    evidence$status == 0 && sum(as.integer(history$n_failures)) == 0, "\n")
```

```{r loss-plot, fig.width=7, fig.height=4, fig.alt="Line chart showing two rounds of federated logistic-regression loss on the LUNG1 radiomics-derived cohort."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(history, ggplot2::aes(x = round, y = loss)) +
    ggplot2::geom_line(color = "#2C6F7D", linewidth = 0.8) +
    ggplot2::geom_point(color = "#2C6F7D", size = 2.8) +
    ggplot2::scale_x_continuous(breaks = history$round) +
    ggplot2::coord_cartesian(ylim = c(min(history$loss) - 0.01,
                                      max(history$loss) + 0.01)) +
    ggplot2::labs(x = "Federated round", y = "Loss",
                  title = "LUNG1 radiomics-derived federated loss") +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot(history$round, history$loss, type = "b",
       xlab = "Federated round", ylab = "Loss",
       main = "LUNG1 radiomics-derived federated loss")
}
```

## Evidence Scope

```{r evidence-scope}
scope <- data.frame(
  item = c("Validated claim", "Explicit non-claim", "Global model shape"),
  value = c(
    "End-to-end functional validation that dsFlower can train on dsImaging-derived radiomic feature tables without moving images, masks or row-level feature data to the researcher.",
    "This two-round logistic-regression run is not a clinical performance claim.",
    "1 x 6 coefficients plus intercept"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(scope)
```
