---
title: "LUNG1 dsImaging Radiomics To dsFlower Evidence"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{LUNG1 dsImaging Radiomics To dsFlower Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  file.path("inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  system.file("extdata", "dsflower_lung1_radiomics_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
evidence <- jsonlite::fromJSON(artifact_path, simplifyVector = FALSE)
```

This article records the real LUNG1 radiomics handoff run used as thesis
evidence. The rendered article uses the recorded output of the live May 2026
workflow so that package checks do not require active Opal servers. The live
workflow is implemented in `inst/demos/lung1_radiomics_to_flower.R`.

The validation path is:

1. `dsImaging` resolves the LUNG1 imaging resource and radiomics assets.
2. `ds.imaging.radiomics.load_features()` loads each site's derived feature
   table into a server-side DataSHIELD symbol.
3. `dsFlowerClient` trains `sklearn_logreg` with FedAvg over the server-side
   tables.
4. Only aggregate training history and model parameters are returned.

## Running The dsImaging To dsFlower Handoff

The steps below walk through the dsImaging to dsFlower handoff: connect to the
nodes, load the server-side radiomic feature tables with dsImaging, and train a
federated model with dsFlower. The LUNG1 study is already registered as an
imaging resource on three Opal/Rock nodes, each holding its own slice of
patients; images, masks and row-level features never leave those nodes. The
chunks are shown for reference and are not executed when the article renders;
the recorded results of the real run follow.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that serve the LUNG1 imaging resource.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsImagingClient)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********")
conns <- DSI::datashield.login(builder$build(), assign = FALSE)

resource <- "dsdemo.lung1_full_study"
```

**2. Load radiomic features server-side with dsImaging.** `ds.imaging.init()`
binds the imaging resource; `ds.imaging.radiomics.load_features()` then
materialises the derived feature table into the server-side symbol `rad`. The
`dataset_id` and `asset_id` identify the radiomic asset produced by the
dsImaging feature-extraction step on each node; `include_metadata = TRUE` adds
the clinical columns and `syntactic_names = TRUE` makes the pyradiomics names
R-safe. The images and masks never move.

```{r load-features, eval=FALSE}
ds.imaging.init(conns, resource, symbol = "img")

ds.imaging.radiomics.load_features(
  conns,
  dataset_id       = "lung1_full_study",
  asset_id         = "pyradiomics_features",
  symbol           = "rad",
  include_metadata = TRUE,
  syntactic_names  = TRUE
)
```

**3. Train the federated model on the derived features.** Hand the
server-side `rad` table to `ds.flower.fit()` with logistic regression and
FedAvg. Only aggregate training history and model parameters come back.

```{r fit, eval=FALSE}
features <- c(
  "original_firstorder_Energy",
  "original_shape_Compactness1",
  "original_glrlm_RunLengthNonUniformity",
  "wavelet.HLH_glrlm_RunLengthNonUniformity",
  "age",
  "gender_male"
)

fit <- ds.flower.fit(
  conns,
  symbol       = "rad",
  target       = "os_2yr_alive",
  features     = features,
  model        = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy     = "fedavg",
  privacy      = "trusted_internal",
  rounds       = 2
)

DSI::datashield.logout(conns)
```

## Recorded Results

These results are loaded from the committed evidence artifact for run
`r evidence$run_id`; no live Opal servers are contacted when the article
renders.

## Server-Side Cohort

```{r cohort}
site_dim_names <- grep("opal", names(evidence$dimensions), value = TRUE)
site_dims <- do.call(rbind, lapply(site_dim_names, function(name) {
  dims <- unlist(evidence$dimensions[[name]])
  data.frame(
    site = sub("dimensions of rad in ", "", name, fixed = TRUE),
    rows = dims[[1]],
    columns = dims[[2]]
  )
}))
combined_dims <- unlist(evidence$dimensions[["dimensions of rad in combined studies"]])
knitr::kable(site_dims)
```

Pooled across the three nodes, the `r evidence$resource` cohort holds
`r combined_dims[[1]]` rows and `r combined_dims[[2]]` columns, with
`r evidence$target` as the federated target.

## Features

```{r features}
feature_roles <- c(rep("radiomic", 4), "metadata", "metadata")
feature_table <- data.frame(
  feature = unlist(evidence$features),
  role = feature_roles,
  stringsAsFactors = FALSE
)
knitr::kable(feature_table)
```

## Federated Training

```{r training}
training <- data.frame(
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy_profile = evidence$privacy,
  rounds = evidence$rounds,
  max_iter = evidence$max_iter,
  stringsAsFactors = FALSE
)
knitr::kable(training)
```

```{r federated-output}
history <- do.call(rbind, lapply(evidence$history, function(x) {
  data.frame(
    round = x$round,
    loss = x$loss,
    n_clients = x$n_clients,
    n_failures = x$n_failures
  )
}))
final_loss <- round(tail(history$loss, 1L), 4)
total_failures <- sum(as.integer(history$n_failures))
run_status <- if (evidence$status == 0 && total_failures == 0) "passed" else "failed"
knitr::kable(history, digits = 4)
```

The run finished with a final federated loss of `r final_loss` and
`r total_failures` total client failures across `r nrow(history)` round(s);
the acceptance check **`r run_status`**.

```{r loss-plot, fig.width=7, fig.height=4, fig.alt="Line chart showing two rounds of federated logistic-regression loss on the LUNG1 radiomics-derived cohort."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(history, ggplot2::aes(x = round, y = loss)) +
    ggplot2::geom_line(color = "#2C6F7D", linewidth = 0.8) +
    ggplot2::geom_point(color = "#2C6F7D", size = 2.8) +
    ggplot2::scale_x_continuous(breaks = history$round) +
    ggplot2::coord_cartesian(ylim = c(min(history$loss) - 0.01,
                                      max(history$loss) + 0.01)) +
    ggplot2::labs(x = "Federated round", y = "Loss",
                  title = "LUNG1 radiomics-derived federated loss") +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot(history$round, history$loss, type = "b",
       xlab = "Federated round", ylab = "Loss",
       main = "LUNG1 radiomics-derived federated loss")
}
```

## Evidence Scope

**Validated claim.** End-to-end functional validation that dsFlower can train
on dsImaging-derived radiomic feature tables without moving images, masks or
row-level feature data to the researcher.

**Explicit non-claim.** This two-round logistic-regression run is not a
clinical performance claim.

**Global model shape.** 1 x 6 coefficients plus intercept.
