---
title: "LUNG1 Direct dsImaging Image ResNet Evidence"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{LUNG1 Direct dsImaging Image ResNet Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_lung1_direct_image_results.json"),
  file.path("inst", "extdata", "dsflower_lung1_direct_image_results.json"),
  system.file("extdata", "dsflower_lung1_direct_image_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
evidence <- jsonlite::fromJSON(artifact_path, simplifyVector = FALSE)
```

This article records the LUNG1 direct-image run used as thesis evidence. The
run trained `pytorch_resnet18` directly from NIfTI assets resolved by
`dsImaging`, without first converting the images into radiomic features. It is a
technical path validation rather than a clinical imaging result. The rendered
article uses the recorded output of the live May 2026 workflow so that package
checks do not require active Opal servers. The live workflow is implemented in
`inst/demos/dsimaging_direct_image_resnet.R`.

The validation path is:

1. `dsImaging` resolves the LUNG1 imaging resource server-side with
   `ds.imaging.init()`.
2. `dsFlowerClient` wraps that server-side imaging handle with
   `ds.flower.connect()` and trains the `pytorch_resnet18` template directly
   from the local NIfTI files, never moving image pixels.
3. A centralized PyTorch baseline on the pooled demo cohort provides a
   path-level comparison.
4. Only aggregate training history and model parameters are returned.

## Running This Validation

The steps below walk through the dsImaging to dsFlower handoff: connect to the
nodes, resolve the server-side image resource with dsImaging, and train a
federated vision model with dsFlower. The LUNG1 study is already registered as
an imaging resource on three Opal/Rock nodes, each holding its own slice of
patients; the NIfTI pixels never leave those nodes. The chunks are shown for
reference and are not executed when the article renders; the recorded results of
the real run follow.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that serve the LUNG1 imaging resource.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsImagingClient)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********")
conns <- DSI::datashield.login(builder$build(), assign = FALSE)

resource <- "dsdemo.imaging_demo"
```

**2. Resolve the image resource with dsImaging.** `ds.imaging.init()` binds the
imaging resource into the server-side symbol `img`; the NIfTI pixels stay on
each Rock.

```{r init-image, eval=FALSE}
ds.imaging.init(conns, resource, symbol = "img")
```

**3. Train the federated model with dsFlower.** Hand the server-side `img`
handle to `ds.flower.connect()`, build a recipe around the `pytorch_resnet18`
template, run FedAvg, then disconnect and log out. Only aggregate training
history and model parameters come back.

```{r fit, eval=FALSE}
flower <- ds.flower.connect(conns, symbol = "img")

recipe <- ds.flower.recipe(
  model      = ds.flower.model.pytorch_resnet18(
    n_classes = 2L, learning_rate = 0.0005,
    batch_size = 1L, local_epochs = 1L
  ),
  strategy   = ds.flower.strategy.fedavg(),
  privacy    = ds.flower.privacy.sandbox_open(),
  target     = "os_2yr_alive",
  num_rounds = 1
)

run <- ds.flower.run(flower, recipe, verbose = TRUE)

ds.flower.disconnect(flower)
DSI::datashield.logout(conns)
```

## Recorded Results

These results are loaded from the committed evidence artifact for run
`r evidence$run_id`; no live Opal servers are contacted when the article
renders.

## Server-Side Image Resource

```{r cohort}
site_counts <- data.frame(
  site = names(evidence$metadata_n),
  samples = as.integer(unlist(evidence$metadata_n)),
  row.names = NULL
)
class_counts <- data.frame(
  class = names(evidence$local_baseline$class_counts),
  samples = as.integer(unlist(evidence$local_baseline$class_counts)),
  row.names = NULL
)

knitr::kable(site_counts)
knitr::kable(class_counts)
```

The `r evidence$resource` resource exposes NIfTI images across the three nodes,
`r sum(site_counts$samples)` samples in total, with `r evidence$target` as the
federated target.

## Training Configuration

```{r training}
training <- data.frame(
  model = evidence$model,
  strategy = "fedavg",
  privacy_profile = evidence$privacy,
  rounds = evidence$rounds,
  local_epochs = evidence$local_baseline$epochs,
  batch_size = evidence$local_baseline$batch_size,
  learning_rate = evidence$local_baseline$learning_rate,
  stringsAsFactors = FALSE
)
knitr::kable(training)
```

## Centralized vs Federated Result

```{r comparison}
comparison <- data.frame(
  run = c("centralized local", "federated dsFlower"),
  engine = c("PyTorch ResNet18", "DataSHIELD + Flower ResNet18"),
  samples = c(evidence$local_baseline$n_samples, sum(site_counts$samples)),
  sites_or_clients = c(1L, evidence$local_vs_federated$federated$clients),
  rounds_or_epochs = c(evidence$local_baseline$epochs,
                       evidence$local_vs_federated$federated$rounds),
  loss = c(evidence$local_baseline$eval$loss,
           evidence$local_vs_federated$federated$loss),
  accuracy = c(evidence$local_baseline$eval$accuracy, NA_real_),
  failures = c(NA_integer_, evidence$local_vs_federated$federated$n_failures)
)
knitr::kable(comparison, digits = 4)
```

The centralized PyTorch baseline reached an evaluation loss of
`r round(evidence$local_baseline$eval$loss, 4)`; the federated dsFlower run
reached `r round(evidence$local_vs_federated$federated$loss, 4)` over
`r evidence$local_vs_federated$federated$rounds` round(s) (delta
`r round(evidence$local_vs_federated$delta$loss_federated_minus_centralized, 4)`)
with `r evidence$local_vs_federated$federated$n_failures` client failures.

```{r loss-plot, fig.width=7, fig.height=4, fig.alt="Bar chart comparing centralized and federated ResNet18 loss on nine LUNG1 NIfTI images."}
plot_data <- data.frame(
  run = factor(comparison$run, levels = comparison$run),
  loss = comparison$loss
)
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(plot_data, ggplot2::aes(x = run, y = loss, fill = run)) +
    ggplot2::geom_col(width = 0.62, show.legend = FALSE) +
    ggplot2::scale_fill_manual(values = c("#4C78A8", "#F58518")) +
    ggplot2::labs(x = NULL, y = "Loss",
                  title = "LUNG1 direct-image ResNet path validation") +
    ggplot2::theme_minimal(base_size = 11)
} else {
  barplot(plot_data$loss, names.arg = as.character(plot_data$run),
          ylim = c(0, max(plot_data$loss) + 0.15), ylab = "Loss",
          main = "LUNG1 direct-image ResNet path validation",
          col = c("#4C78A8", "#F58518"))
}
```

## Cleanup And Evidence Scope

```{r cleanup}
cleanup <- data.frame(
  site = names(evidence$active_supernodes_after),
  active_supernodes = as.integer(unlist(evidence$active_supernodes_after)),
  row.names = NULL
)
knitr::kable(cleanup)
```

After the run, every node reported zero active SuperNodes, so the cleanup step
**`r if (all(cleanup$active_supernodes == 0L)) "passed" else "failed"`**.

**Validated claim.** End-to-end functional validation that dsFlower can consume
dsImaging-resolved server-side NIfTI image assets and train a vision template
without transferring image pixels to the researcher.

**Explicit non-claim.** This nine-image run is a technical path validation, not
a clinical imaging benchmark.
