---
title: "LUNG1 dsImaging Radiomics To dsFlower Evidence"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{LUNG1 dsImaging Radiomics To dsFlower Evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  file.path("inst", "extdata", "dsflower_lung1_radiomics_results.json"),
  system.file("extdata", "dsflower_lung1_radiomics_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
evidence <- jsonlite::fromJSON(artifact_path, simplifyVector = FALSE)
```

This article records the real LUNG1 radiomics handoff run used as thesis
evidence. The rendered article uses the recorded output of the live May 2026
workflow so that package checks do not require active Opal servers. The live
workflow is implemented in `inst/demos/lung1_radiomics_to_flower.R`.

The validation path is:

1. `dsImaging` resolves the LUNG1 imaging resource and radiomics assets.
2. `ds.imaging.radiomics.load_features()` loads each site's derived feature
   table into a server-side DataSHIELD symbol.
3. `dsFlowerClient` trains `sklearn_logreg` with FedAvg over the server-side
   tables.
4. Only aggregate training history and model parameters are returned.

```{r recorded-run}
cat("Source script: inst/demos/lung1_radiomics_to_flower.R\n")
cat("Recorded run_id:", evidence$run_id, "\n")
```

## Reproduction Command

```{r reproduction-command, eval=FALSE}
Sys.setenv(
  DSFLOWER_OPAL_URLS = "https://localhost:8443,https://localhost:8444,https://localhost:8445",
  OPAL_USER = "administrator",
  OPAL_PASSWORD = "admin123",
  LUNG1_WORKDIR = "/tmp/dsimaging_lung1_full",
  LUNG1_OPAL_PROJECT = "dsdemo",
  LUNG1_OPAL_RESOURCE = "lung1_full_study",
  DSFLOWER_RADFLOWER_ROUNDS = "2",
  DSFLOWER_RADFLOWER_PRIVACY = "trusted_internal"
)
source(system.file("demos", "lung1_radiomics_to_flower.R",
                   package = "dsFlowerClient"))
```

## Server-Side Cohort

```{r cohort}
site_dim_names <- grep("opal", names(evidence$dimensions), value = TRUE)
site_dims <- do.call(rbind, lapply(site_dim_names, function(name) {
  dims <- unlist(evidence$dimensions[[name]])
  data.frame(
    site = sub("dimensions of rad in ", "", name, fixed = TRUE),
    rows = dims[[1]],
    columns = dims[[2]]
  )
}))
combined_dims <- unlist(evidence$dimensions[["dimensions of rad in combined studies"]])
knitr::kable(site_dims)

data.frame(
  resource = evidence$resource,
  target = evidence$target,
  combined_rows = combined_dims[[1]],
  combined_columns = combined_dims[[2]],
  stringsAsFactors = FALSE
)
```

## Features

```{r features}
feature_roles <- c(rep("radiomic", 4), "metadata", "metadata")
feature_table <- data.frame(
  feature = unlist(evidence$features),
  role = feature_roles,
  stringsAsFactors = FALSE
)
knitr::kable(feature_table)
```

## Federated Training

```{r training}
training <- data.frame(
  model = "sklearn_logreg",
  strategy = "fedavg",
  privacy_profile = evidence$privacy,
  rounds = evidence$rounds,
  max_iter = evidence$max_iter,
  stringsAsFactors = FALSE
)
knitr::kable(training)
```

```{r federated-output}
history <- do.call(rbind, lapply(evidence$history, function(x) {
  data.frame(
    round = x$round,
    loss = x$loss,
    n_clients = x$n_clients,
    n_failures = x$n_failures
  )
}))
knitr::kable(history, digits = 4)

cat("[dsFlower] run_id:", evidence$run_id, "\n")
cat("[dsFlower] model_id:", evidence$model_id, "\n")
cat("[dsFlower] output_dir:", evidence$flower_output_dir, "\n")
cat("[acceptance] final loss:", sprintf("%.4f", tail(history$loss, 1L)), "\n")
cat("[acceptance] total client failures:",
    sum(as.integer(history$n_failures)), "\n")
cat("[acceptance] PASS:",
    evidence$status == 0 && sum(as.integer(history$n_failures)) == 0, "\n")
```

```{r loss-plot, fig.width=7, fig.height=4, fig.alt="Line chart showing two rounds of federated logistic-regression loss on the LUNG1 radiomics-derived cohort."}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(history, ggplot2::aes(x = round, y = loss)) +
    ggplot2::geom_line(color = "#2C6F7D", linewidth = 0.8) +
    ggplot2::geom_point(color = "#2C6F7D", size = 2.8) +
    ggplot2::scale_x_continuous(breaks = history$round) +
    ggplot2::coord_cartesian(ylim = c(min(history$loss) - 0.01,
                                      max(history$loss) + 0.01)) +
    ggplot2::labs(x = "Federated round", y = "Loss",
                  title = "LUNG1 radiomics-derived federated loss") +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot(history$round, history$loss, type = "b",
       xlab = "Federated round", ylab = "Loss",
       main = "LUNG1 radiomics-derived federated loss")
}
```

## Evidence Scope

```{r evidence-scope}
scope <- data.frame(
  item = c("Validated claim", "Explicit non-claim", "Global model shape"),
  value = c(
    "End-to-end functional validation that dsFlower can train on dsImaging-derived radiomic feature tables without moving images, masks or row-level feature data to the researcher.",
    "This two-round logistic-regression run is not a clinical performance claim.",
    "1 x 6 coefficients plus intercept"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(scope)
```
