---
title: "PathMNIST Direct Image ResNet Benchmark"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{PathMNIST Direct Image ResNet Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_pathmnist_direct_image_results.json"),
  file.path("inst", "extdata", "dsflower_pathmnist_direct_image_results.json"),
  system.file("extdata", "dsflower_pathmnist_direct_image_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
if (is.na(artifact_path)) {
  stop("PathMNIST direct-image evidence JSON is not available.")
}
evidence <- jsonlite::fromJSON(artifact_path)
```

This vignette records a direct-image dsFlower benchmark on PathMNIST, a
MedMNIST v2 histopathology image dataset. The run uses a balanced binary subset
of adipose tissue versus colorectal adenocarcinoma epithelium, split evenly
across three Opal/Rock servers. The Opal table contains metadata and
relative image paths; image files are read inside each Rock by the
`pytorch_resnet18` template.

The benchmark has two branches. The `trusted_internal` branch compares a
federated ResNet-18 checkpoint with a centralized PyTorch baseline over the same
1,500 images. The `consortium_internal` branch repeats the federated execution
under the profile that requires Secure Aggregation, suppresses per-node metrics
and still produces a global model checkpoint.

It is laid out in two parts: first the commands you would run to reproduce the
benchmark, then the recorded results of that run loaded from the committed
evidence artifact. No live Opal servers are required to render this page; the
command chunks are shown for reference and are not executed.

## Running This Benchmark

Each node holds an image metadata table (`label` plus the relative image paths)
while the PNG files stay on the matching Rock server; only aggregate training
history and model parameters return to the researcher.

**1. Connect to the DataSHIELD nodes.** Open a single session across the three
Opal nodes that serve the PathMNIST metadata table.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "node1", url = "https://opal-node-1.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.pathmnist_cohort")
builder$append(server = "node2", url = "https://opal-node-2.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.pathmnist_cohort")
builder$append(server = "node3", url = "https://opal-node-3.example.org",
               user = "researcher", password = "********",
               table = "dsflower_demo.pathmnist_cohort")
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = "D")
```

**2. Run the federated fit.** Build the ResNet-18 model and recipe, stage the
nodes, run FedAvg over the server-side image tables, then clean up and log out.

```{r fit, eval=FALSE}
model <- ds.flower.model.pytorch_resnet18(
  n_classes = 2L, learning_rate = 0.001, batch_size = 64L,
  local_epochs = 1L, image_size = 32L
)

recipe <- ds.flower.recipe(
  model      = model,
  strategy   = ds.flower.strategy.fedavg(),
  privacy    = ds.flower.privacy.trusted_internal(),
  target     = "label",
  num_rounds = 2
)

ds.flower.nodes.init(conns, data = "D", symbol = "flower")
ds.flower.nodes.prepare(
  conns, symbol = "flower",
  target_column = "label",
  run_config = c(recipe$model$params, recipe$strategy$params,
                 list(data_type = "image")),
  privacy = recipe$privacy, template_name = recipe$model$template
)
ds.flower.nodes.ensure(conns, symbol = "flower", template_name = recipe$model$template)
fit <- ds.flower.run.start(recipe, conns = conns, verbose = TRUE)
ds.flower.nodes.cleanup(conns, symbol = "flower")

DSI::datashield.logout(conns)
```

The `consortium_internal` branch repeats the same run with
`privacy = ds.flower.privacy.consortium_internal()`, which enforces Secure
Aggregation and suppresses the per-node metrics.

## Recorded Results

The values below come from the committed evidence artifact for the run described
above; no live servers are contacted when the article renders.

## Dataset And Split

```{r dataset}
class_table <- data.frame(
  source_class = names(evidence$dataset$classes),
  binary_label = as.integer(unlist(evidence$dataset$binary_mapping)),
  class_name = unname(unlist(evidence$dataset$classes)),
  row.names = NULL
)
site_table <- evidence$site_counts

knitr::kable(class_table)
knitr::kable(site_table)
```

The benchmark draws `r evidence$n_total` `r evidence$dataset$dataset` images
from `r evidence$dataset$source`, split into `r evidence$n_per_site` rows on
each of `r evidence$n_sites` nodes, resized to
`r evidence$image_size`x`r evidence$image_size`, and trains `r evidence$model`
for `r evidence$rounds` federated rounds.

## Centralized Baseline

```{r centralized}
centralized <- data.frame(
  engine = "centralized PyTorch",
  samples = evidence$centralized$n_samples,
  epochs = evidence$centralized$epochs,
  loss = evidence$centralized$eval$loss,
  accuracy = evidence$centralized$eval$accuracy,
  stringsAsFactors = FALSE
)
knitr::kable(centralized, digits = 4)
```

The centralized PyTorch baseline trained on `r evidence$centralized$n_samples`
samples for `r evidence$centralized$epochs` epoch(s), reaching an evaluation
loss of `r round(evidence$centralized$eval$loss, 4)` and accuracy
`r round(evidence$centralized$eval$accuracy, 4)`.

## Federated Trusted Branch

```{r trusted}
trusted_profile <- evidence$profile_results$trusted_internal
trusted_history <- trusted_profile$history
comparison <- evidence$trusted_internal_comparison

knitr::kable(trusted_history, digits = 4)

trusted_table <- data.frame(
  run = c("centralized", "federated checkpoint"),
  loss = c(comparison$centralized_loss,
           comparison$federated_checkpoint_loss),
  accuracy = c(comparison$centralized_accuracy,
               comparison$federated_checkpoint_accuracy),
  failures = c(NA_integer_, comparison$federated_failures),
  stringsAsFactors = FALSE
)
knitr::kable(trusted_table, digits = 4)
```

The federated checkpoint's final history loss was
`r round(comparison$federated_history_loss, 4)`; against the centralized
baseline the checkpoint loss delta was `r round(comparison$delta_loss, 4)` and
the accuracy delta `r round(comparison$delta_accuracy, 4)`, leaving this branch
**`r comparison$validation_status`**.

```{r trusted-plot, fig.width=7, fig.height=3.6, fig.alt="Bar chart comparing centralized and federated PathMNIST ResNet loss and accuracy."}
plot_df <- data.frame(
  run = rep(c("centralized", "federated"), 2),
  metric = rep(c("loss", "accuracy"), each = 2),
  value = c(comparison$centralized_loss,
            comparison$federated_checkpoint_loss,
            comparison$centralized_accuracy,
            comparison$federated_checkpoint_accuracy),
  stringsAsFactors = FALSE
)
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(plot_df, ggplot2::aes(x = run, y = value, fill = run)) +
    ggplot2::geom_col(width = 0.62, show.legend = FALSE) +
    ggplot2::facet_wrap(~ metric, scales = "free_y") +
    ggplot2::scale_fill_manual(values = c("#4C78A8", "#F58518")) +
    ggplot2::labs(x = NULL, y = NULL) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  old <- par(mfrow = c(1, 2))
  on.exit(par(old), add = TRUE)
  barplot(plot_df$value[plot_df$metric == "loss"],
          names.arg = c("central", "fed"), ylab = "loss",
          col = c("#4C78A8", "#F58518"))
  barplot(plot_df$value[plot_df$metric == "accuracy"],
          names.arg = c("central", "fed"), ylab = "accuracy",
          col = c("#4C78A8", "#F58518"))
}
```

## Secure Aggregation Branch

```{r secagg}
secagg <- evidence$consortium_internal_secagg
```

The `consortium_internal` branch ran with Secure Aggregation required, metric
release set to `r secagg$metric_release`, `r secagg$federated_failures`
federated client failures, and a persisted global checkpoint
(`r secagg$saved_model_exists`), leaving the branch
**`r secagg$validation_status`**.

The SecAgg branch is not interpreted through its loss value because
`consortium_internal` suppresses released per-node metrics. The validation
criteria for this branch are run completion, zero federated client failures,
activation of the Secure Aggregation requirement and production of the global
model artefact.

## Evidence Scope

**Direct-image learning path.** `r evidence$n_total` PathMNIST PNG images remain
file-backed on the server side while Opal tables expose only metadata and
relative paths.

**Centralized-versus-federated utility.** The federated ResNet-18 checkpoint
reaches `r round(comparison$federated_checkpoint_accuracy, 4)` accuracy versus
`r round(comparison$centralized_accuracy, 4)` for the centralized PyTorch
baseline on the same image table.

**SecAgg profile execution.** The consortium profile completes with zero client
failures, metric suppression and a persisted global checkpoint.
