---
title: "PathMNIST Direct Image ResNet Benchmark"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{PathMNIST Direct Image ResNet Benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
artifact_paths <- c(
  file.path("..", "inst", "extdata", "dsflower_pathmnist_direct_image_results.json"),
  file.path("inst", "extdata", "dsflower_pathmnist_direct_image_results.json"),
  system.file("extdata", "dsflower_pathmnist_direct_image_results.json",
              package = "dsFlowerClient")
)
artifact_path <- artifact_paths[nzchar(artifact_paths) & file.exists(artifact_paths)][1]
if (is.na(artifact_path)) {
  stop("PathMNIST direct-image evidence JSON is not available.")
}
evidence <- jsonlite::fromJSON(artifact_path)
```

This vignette records a direct-image dsFlower benchmark on PathMNIST, a
MedMNIST v2 histopathology image dataset. The run uses a balanced binary subset
of adipose tissue versus colorectal adenocarcinoma epithelium, split evenly
across three Opal/Rock servers. The Opal table contains metadata and
relative image paths; image files are read inside each Rock by the
`pytorch_resnet18` template.

The benchmark has two branches. The `trusted_internal` branch compares a
federated ResNet-18 checkpoint with a centralized PyTorch baseline over the same
1,500 images. The `consortium_internal` branch repeats the federated execution
under the profile that requires Secure Aggregation, suppresses per-node metrics
and still produces a global model checkpoint.

## Reproduction Command

```{r reproduction-command, eval=FALSE}
Sys.setenv(
  DSFLOWER_OPAL_URLS = "https://localhost:8443,https://localhost:8444,https://localhost:8445",
  OPAL_USER = "administrator",
  OPAL_PASSWORD = "admin123",
  DSFLOWER_PATHMNIST_N_PER_SITE = "500",
  DSFLOWER_PATHMNIST_PRIVACY_PROFILES = "trusted_internal,consortium_internal"
)
source(system.file("demos", "benchmark_pathmnist_direct_image.R",
                   package = "dsFlowerClient"))
```

## Dataset And Split

```{r dataset}
class_table <- data.frame(
  source_class = names(evidence$dataset$classes),
  binary_label = as.integer(unlist(evidence$dataset$binary_mapping)),
  class_name = unname(unlist(evidence$dataset$classes)),
  row.names = NULL
)
site_table <- evidence$site_counts

summary_table <- data.frame(
  field = c("dataset", "source", "sites", "rows per site",
            "total rows", "image size", "model", "rounds"),
  value = c(evidence$dataset$dataset,
            evidence$dataset$source,
            evidence$n_sites,
            evidence$n_per_site,
            evidence$n_total,
            paste0(evidence$image_size, " x ", evidence$image_size),
            evidence$model,
            evidence$rounds),
  stringsAsFactors = FALSE
)

knitr::kable(class_table)
knitr::kable(site_table)
knitr::kable(summary_table)
```

## Centralized Baseline

```{r centralized}
centralized <- data.frame(
  engine = "centralized PyTorch",
  samples = evidence$centralized$n_samples,
  epochs = evidence$centralized$epochs,
  loss = evidence$centralized$eval$loss,
  accuracy = evidence$centralized$eval$accuracy,
  stringsAsFactors = FALSE
)
knitr::kable(centralized, digits = 4)

cat("[centralized] epoch losses:",
    paste(sprintf("%.6f", evidence$centralized$train_loss_by_epoch),
          collapse = " -> "), "\n")
cat("[centralized] eval loss:",
    sprintf("%.6f", evidence$centralized$eval$loss), "\n")
cat("[centralized] eval accuracy:",
    sprintf("%.4f", evidence$centralized$eval$accuracy), "\n")
```

## Federated Trusted Branch

```{r trusted}
trusted_profile <- evidence$profile_results$trusted_internal
trusted_history <- trusted_profile$history
comparison <- evidence$trusted_internal_comparison

knitr::kable(trusted_history, digits = 4)

trusted_table <- data.frame(
  run = c("centralized", "federated checkpoint"),
  loss = c(comparison$centralized_loss,
           comparison$federated_checkpoint_loss),
  accuracy = c(comparison$centralized_accuracy,
               comparison$federated_checkpoint_accuracy),
  failures = c(NA_integer_, comparison$federated_failures),
  stringsAsFactors = FALSE
)
knitr::kable(trusted_table, digits = 4)

cat("[federated] final history loss:",
    sprintf("%.6f", comparison$federated_history_loss), "\n")
cat("[delta] checkpoint loss - centralized loss:",
    sprintf("%.6f", comparison$delta_loss), "\n")
cat("[delta] checkpoint accuracy - centralized accuracy:",
    sprintf("%.4f", comparison$delta_accuracy), "\n")
cat("[validation] status:", comparison$validation_status, "\n")
```

```{r trusted-plot, fig.width=7, fig.height=3.6, fig.alt="Bar chart comparing centralized and federated PathMNIST ResNet loss and accuracy."}
plot_df <- data.frame(
  run = rep(c("centralized", "federated"), 2),
  metric = rep(c("loss", "accuracy"), each = 2),
  value = c(comparison$centralized_loss,
            comparison$federated_checkpoint_loss,
            comparison$centralized_accuracy,
            comparison$federated_checkpoint_accuracy),
  stringsAsFactors = FALSE
)
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(plot_df, ggplot2::aes(x = run, y = value, fill = run)) +
    ggplot2::geom_col(width = 0.62, show.legend = FALSE) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.4f", value)),
                       vjust = -0.35, size = 3.3) +
    ggplot2::facet_wrap(~ metric, scales = "free_y") +
    ggplot2::scale_fill_manual(values = c("#4C78A8", "#F58518")) +
    ggplot2::labs(x = NULL, y = NULL) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  old <- par(mfrow = c(1, 2))
  on.exit(par(old), add = TRUE)
  barplot(plot_df$value[plot_df$metric == "loss"],
          names.arg = c("central", "fed"), ylab = "loss",
          col = c("#4C78A8", "#F58518"))
  barplot(plot_df$value[plot_df$metric == "accuracy"],
          names.arg = c("central", "fed"), ylab = "accuracy",
          col = c("#4C78A8", "#F58518"))
}
```

## Secure Aggregation Branch

```{r secagg}
secagg <- evidence$consortium_internal_secagg
secagg_table <- data.frame(
  field = c("privacy profile", "secure aggregation required",
            "metric release", "federated failures",
            "global checkpoint produced", "validation status"),
  value = c("consortium_internal",
            as.character(secagg$secure_aggregation_required),
            secagg$metric_release,
            secagg$federated_failures,
            as.character(secagg$saved_model_exists),
            secagg$validation_status),
  stringsAsFactors = FALSE
)
knitr::kable(secagg_table)
```

The SecAgg branch is not interpreted through its loss value because
`consortium_internal` suppresses released per-node metrics. The validation
criteria for this branch are run completion, zero federated client failures,
activation of the Secure Aggregation requirement and production of the global
model artefact.

## Evidence Scope

```{r evidence-scope}
scope <- data.frame(
  claim = c("Direct-image learning path",
            "Centralized-versus-federated utility",
            "SecAgg profile execution"),
  evidence = c(
    "1,500 PathMNIST PNG images remain file-backed on the server side while Opal tables expose only metadata and relative paths.",
    "The federated ResNet-18 checkpoint reaches 0.9860 accuracy versus 0.9947 for the centralized PyTorch baseline on the same image table.",
    "The consortium profile completes with zero client failures, metric suppression and a persisted global checkpoint."
  ),
  stringsAsFactors = FALSE
)
knitr::kable(scope)
```
