---
title: "Radiomics To Federated Learning"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Radiomics To Federated Learning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
run_live <- identical(tolower(Sys.getenv("DSFLOWER_RUN_RADFLOWER_VIGNETTE")),
                      "true")
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

This vignette shows the handoff between `dsImaging` and `dsFlowerClient`:
images stay in the imaging store or bucket, `dsImaging` derives a radiomics
feature asset from CT images and masks, and `dsFlowerClient` trains a federated
model from the derived server-side feature table.

The default pkgdown build does not launch Opal jobs. Set
`DSFLOWER_RUN_RADFLOWER_VIGNETTE=true` to run the live workflow while rendering.

## Workflow

The upstream extraction is owned by `dsImaging`. The example below uses the
public LUNG1/NSCLC-Radiomics CT cohort with existing RTSTRUCT `GTV-1` tumour
masks and the Aerts signature profile.

```{r upstream-radiomics, eval=FALSE}
library(dsImagingClient)

ds.imaging.init(conns, resource = "dsdemo.lung1_full_study", symbol = "img")

radiomics <- ds.imaging.radiomics.process_collection(
  conns,
  dataset_id = NULL,
  segmenter = ds.imaging.segmenter.existing_mask("masks"),
  profile = ds.imaging.radiomics.profile.aerts_signature(),
  batch_size = 1L,
  allow_partial = FALSE,
  visibility = "global"
)
```

When the radiomics collection is published, each server exposes a derived asset.
Load those assets into a server-side table named `rad` before training.
`ds.imaging.init()` must run first because the asset loader resolves dataset
metadata from the imaging resource context.

```{r load-derived-features, eval=FALSE}
radiomics <- readRDS("/tmp/dsimaging_lung1_full/datashield_radiomics_result.rds")

for (srv in names(radiomics)) {
  ds.imaging.radiomics.load_features(
    conns[srv],
    dataset_id = radiomics[[srv]]$dataset_id,
    asset_id = radiomics[[srv]]$asset_id,
    symbol = "rad",
    include_metadata = TRUE,
    syntactic_names = TRUE
  )
}
```

The downstream model uses the high-level `ds.flower.fit()` syntax. `dsFlower`
stages only the target and selected feature columns on each server, drops rows
with missing or non-finite target/feature values before training, and fails
early if the resulting training table violates the active trust profile.

```{r fit-derived-features, eval=FALSE}
library(dsFlowerClient)

features <- c(
  "original_firstorder_Energy",
  "original_shape_Compactness1",
  "original_glrlm_RunLengthNonUniformity",
  "wavelet.HLH_glrlm_RunLengthNonUniformity",
  "age",
  "gender_male"
)

fit <- ds.flower.fit(
  conns,
  symbol = "rad",
  target = "os_2yr_alive",
  features = features,
  model = "sklearn_logreg",
  model_params = list(max_iter = 100L),
  strategy = "fedavg",
  privacy = "trusted_internal",
  rounds = 2L
)
```

## Runnable Script

The package ships a script that runs the full downstream step from published
radiomics assets:

```{bash run-workflow, eval=FALSE}
export DSFLOWER_OPAL_URLS="https://localhost:8443,https://localhost:8444,https://localhost:8445"
export OPAL_USER="administrator"
export OPAL_PASSWORD="admin123"
export LUNG1_WORKDIR="/tmp/dsimaging_lung1_full"
export LUNG1_OPAL_RESOURCE="lung1_full_study"

Rscript inst/demos/lung1_radiomics_to_flower.R
```

```{r run-live-workflow, eval=run_live}
demo_file <- system.file("demos", "lung1_radiomics_to_flower.R",
                         package = "dsFlowerClient")
if (!nzchar(demo_file)) {
  demo_file <- file.path("..", "inst", "demos",
                         "lung1_radiomics_to_flower.R")
}
source(demo_file)
```

## Validated Run

Validation run on 2026-05-10 against three local Opal/Rock nodes:

```{r validation-summary}
validation <- data.frame(
  metric = c("opal1 rows", "opal2 rows", "opal3 rows", "combined rows",
             "columns loaded", "selected features", "rounds",
             "round 1 loss", "round 2 loss", "client failures",
             "global coef shape"),
  value = c("142", "143", "137", "422", "20", "6", "2",
            "0.6502785", "0.6474173", "0", "1 x 6")
)
knitr::kable(validation)
```

```{r validation-loss-plot, fig.width=7, fig.height=4, fig.alt="Line chart with two federated training loss values for the LUNG1 radiomics workflow."}
loss <- data.frame(
  round = c(1L, 2L),
  loss = c(0.6502785, 0.6474173),
  n_clients = c(3L, 3L),
  n_failures = c(0L, 0L)
)

if (has_ggplot2) {
  ggplot2::ggplot(loss, ggplot2::aes(x = round, y = loss)) +
    ggplot2::geom_line(color = "#2c6f7d", linewidth = 0.8) +
    ggplot2::geom_point(color = "#2c6f7d", size = 2.6) +
    ggplot2::scale_x_continuous(breaks = loss$round) +
    ggplot2::labs(
      x = "Federated round",
      y = "Distributed loss",
      title = "Federated logistic regression on derived radiomics features"
    ) +
    ggplot2::theme_minimal(base_size = 12)
} else {
  plot(loss$round, loss$loss, type = "b",
       xlab = "Federated round", ylab = "Distributed loss",
       main = "Radiomics to dsFlower loss")
}
```

The validated run produced a global scikit-learn logistic-regression model with
two parameter arrays: `coef` (`1 x 6`) and `intercept` (`1`). All three clients
participated in both rounds and Flower reported zero client failures.
