---
title: "Secure Connections: TLS Auto-Certificates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Secure Connections: TLS Auto-Certificates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
quiet <- function(expr) invisible(capture.output(suppressMessages(expr)))
```

## Why TLS matters

The Flower gRPC channel between the SuperLink (your machine) and the
SuperNodes (Opal/Rock servers) carries model parameters, gradients, and
evaluation metrics. Without encryption, anyone on the network path could:

- **Eavesdrop** on gradient updates (gradient inversion attacks can leak
  training data)
- **Redirect** a SuperNode to a rogue SuperLink that sends back a poisoned
  model
- **Tamper** with gradient updates in transit, degrading model quality

The DataSHIELD/Opal control channel is already HTTPS, but the Flower gRPC
training channel is separate. By default it runs unencrypted. Setting
`insecure = FALSE` enables TLS.

## How certificate generation works

When `insecure = FALSE`, dsFlowerClient auto-generates ephemeral TLS
certificates using the system `openssl` CLI. Let's see what it produces:

```{r cert-gen}
library(dsFlowerClient)

cert_dir <- tempfile("dsflower_certs_")
certs <- dsFlowerClient:::.generate_tls_certs(cert_dir)

cat("Generated files:\n")
for (f in list.files(cert_dir)) {
  sz <- file.size(file.path(cert_dir, f))
  cat(sprintf("  %-15s  %d bytes\n", f, sz))
}
```

Seven files. Let's look at each important one.

### The CA certificate (root of trust)

```{r ca-cert}
ca_lines <- readLines(certs$ca_cert_path)
cat(paste(ca_lines, collapse = "\n"), "\n")
```

This CA certificate gets distributed to every Opal. The CA private key
(`ca.key`) never leaves your machine:

```{r ca-key-perms}
cat("CA key permissions:", format(file.info(certs$ca_key_path)$mode), "\n")
cat("(0600 = owner read/write only)\n")
```

### The SAN configuration

Subject Alternative Names determine which hostnames/IPs the server
certificate is valid for:

```{r san-config}
cat(readLines(file.path(cert_dir, "san.cnf")), sep = "\n")
```

| SAN | Purpose |
|:----|:--------|
| `DNS:localhost` | SuperNode on the same machine |
| `DNS:host.docker.internal` | SuperNode inside Docker |
| `IP:127.0.0.1` | Loopback connections |
| `IP:<LAN IP>` | SuperNode on the local network |

### The server certificate

```{r server-cert}
srv_lines <- readLines(certs$srv_cert_path)
cat(paste(srv_lines, collapse = "\n"), "\n")
```

This is what the SuperLink presents during the TLS handshake. SuperNodes
verify it against the CA certificate.

```{r cert-cleanup, include = FALSE}
unlink(cert_dir, recursive = TRUE)
```

### The generation sequence

All seven steps take roughly 5 milliseconds total (EC P-256 is fast):

```
1. Sys.which("openssl")          → find the binary
2. ecparam -name prime256v1      → probe EC support
3. Write san.cnf                 → SAN configuration
4. ecparam -genkey + req -x509   → CA key + self-signed cert (1-day expiry)
5. ecparam -genkey + req + x509  → server key + CSR + sign with CA
6. Sys.chmod("ca.key", "0600")   → restrict permissions
7. readLines("ca.pem")           → load PEM for distribution
```

We use `-extfile san.cnf` instead of `-addext` because `-addext` is not
supported by LibreSSL (the default on macOS).

```{r openssl-version}
cat("OpenSSL version:", system2("openssl", "version", stdout = TRUE), "\n")
```

## Live demo: insecure vs TLS

Let's connect to three real Opal servers and compare both modes.

```{r connect-tls, cache = FALSE}
library(DSI)
library(DSOpal)

builder <- DSI::newDSLoginBuilder()
builder$append(server = "site_a", url = "https://localhost:8443",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")
builder$append(server = "site_b", url = "https://localhost:8444",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")
builder$append(server = "site_c", url = "https://localhost:8445",
               user = "administrator", password = "admin123",
               driver = "OpalDriver",
               options = "list(ssl_verifyhost=0, ssl_verifypeer=0)")

conns <- DSI::datashield.login(logins = builder$build(), assign = FALSE)
cat("Connected to:", paste(names(conns), collapse = ", "), "\n")
```

### Insecure mode

```{r insecure-run}
quiet(ds.flower.nodes.init(conns, resource = "dsflower_test.flower_node"))
quiet(ds.flower.nodes.prepare(conns, target_column = "target",
                               feature_columns = c("f1", "f2", "f3", "f4", "f5")))

quiet(ds.flower.superlink.start(insecure = TRUE))
Sys.sleep(2)
status_insecure <- ds.flower.superlink.status()
cat(sprintf("Insecure mode:\n"))
cat(sprintf("  Running:       %s\n", status_insecure$running))
cat(sprintf("  Insecure:      %s\n", status_insecure$insecure))
cat(sprintf("  TLS cert:      %s\n",
    ifelse(is.null(status_insecure$ca_cert_pem), "none", "present")))
cat(sprintf("  Federation ID: %s\n", status_insecure$federation_id))

quiet(ds.flower.nodes.ensure(conns, symbol = "flower"))
Sys.sleep(10)

recipe <- ds.flower.recipe(
  task     = ds.flower.task.classification(),
  model    = ds.flower.model.sklearn_logreg(C = 1.0),
  strategy = ds.flower.strategy.fedavg(
    min_fit_clients = 3L, min_available_clients = 3L),
  num_rounds      = 3L,
  target_column   = "target",
  feature_columns = c("f1", "f2", "f3", "f4", "f5")
)

cat("\nTraining (insecure, 3 sites, 3 rounds)...\n\n")
run_insecure <- ds.flower.run.start(recipe, verbose = TRUE)
cat(sprintf("\nExit status: %s\n", run_insecure$status))

quiet(ds.flower.nodes.cleanup(conns, symbol = "flower"))
quiet(ds.flower.superlink.stop())
```

### TLS mode

Now the same thing with encryption. Only one line changes:
`insecure = FALSE`.

```{r tls-run}
quiet(ds.flower.nodes.init(conns, resource = "dsflower_test.flower_node"))
quiet(ds.flower.nodes.prepare(conns, target_column = "target",
                               feature_columns = c("f1", "f2", "f3", "f4", "f5")))

quiet(ds.flower.superlink.start(insecure = FALSE))
Sys.sleep(2)
status_tls <- ds.flower.superlink.status()
cat(sprintf("TLS mode:\n"))
cat(sprintf("  Running:       %s\n", status_tls$running))
cat(sprintf("  Insecure:      %s\n", status_tls$insecure))
cat(sprintf("  TLS cert:      %s\n",
    ifelse(is.null(status_tls$ca_cert_pem), "none",
           paste0(substr(status_tls$ca_cert_pem, 1, 27), "... (",
                  nchar(status_tls$ca_cert_pem), " chars)"))))
cat(sprintf("  Federation ID: %s\n", status_tls$federation_id))

quiet(ds.flower.nodes.ensure(conns, symbol = "flower"))
Sys.sleep(10)

cat("\nTraining (TLS encrypted, 3 sites, 3 rounds)...\n\n")
run_tls <- ds.flower.run.start(recipe, verbose = TRUE)
cat(sprintf("\nExit status: %s\n", run_tls$status))

quiet(ds.flower.nodes.cleanup(conns, symbol = "flower"))
quiet(ds.flower.superlink.stop())
```

The loss values are comparable. TLS is transparent to the training
protocol: the only difference is that all gRPC traffic was encrypted.

## How the CA cert reaches the Opals

The CA certificate travels through the DataSHIELD HTTPS channel (which is
already secure). Here is the encoding chain:

```
Your machine                        Opal/Rock
-----------                         ---------

1. ca_cert_pem (string, ~500 bytes)
   |
2. .ds_encode(list(pem = ca_cert_pem))
   → JSON → Base64 → URL-safe → "B64:..." prefix
   |
3. flowerEnsureSuperNodeDS(         ──[HTTPS]──>  4. .ds_arg() decodes
      symbol, address,                                → Base64 → JSON
      fed_id, "B64:eyJw...")                          → list(pem = "-----BEGIN...")
                                                  |
                                               5. writeLines(pem, "ca.pem")
                                                  |
                                               6. flower-supernode
                                                    --root-certificates ca.pem
```

## Certificate lifecycle

| Event | What happens |
|:------|:-------------|
| `superlink.start(insecure=FALSE)` | Certs generated in temp dir (~5ms) |
| `nodes.ensure()` | CA public cert sent to each Opal via HTTPS |
| SuperNode connects | gRPC TLS handshake verifies server cert |
| Training runs | All gRPC traffic encrypted |
| `superlink.stop()` | Entire temp dir deleted (certs included) |
| 24 hours | Certs expire (safety net if stop was not called) |

## Known limitations

- **No `extra_sans` in the public API**: `.generate_tls_certs()` accepts
  extra SANs but `superlink.start()` does not expose it yet.
- **No certificate renewal**: certs expire after 24 hours. Very long
  training runs may need manually generated certificates.
- **Opal method registration**: reinstalling dsFlower on Rock requires
  re-registering methods through Opal's admin UI.

```{r final-cleanup-tls, include = FALSE, cache = FALSE}
DSI::datashield.logout(conns)
```
