---
title: "Subject-level survival and segmentation evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Subject-level survival and segmentation evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
evidence_dir <- file.path("..", "inst", "extdata", "campaign", "survival")
if (!dir.exists(evidence_dir)) {
  evidence_dir <- system.file("extdata", "campaign", "survival", package="dsFlowerClient")
}
evidence_files <- list.files(evidence_dir, pattern="\\.json$", full.names=TRUE)
evidence <- lapply(evidence_files, jsonlite::read_json, simplifyVector=FALSE)
cells <- Filter(function(x) identical(x$record_type, "cell") ||
  (!is.null(x$variant) && !is.null(x$status)), evidence)
executed <- Filter(function(x) identical(x$status, "executed"), cells)
failed <- Filter(function(x) identical(x$status, "failed"), cells)
```

`pytorch_aft` and `pytorch_discrete_hazard` train baseline survival predictors
under the existing node-enforced DP mechanism. The custodian must configure
patient privacy and a stable subject identifier. Epsilon, delta, clipping,
privacy-unit selection and the accounting population are not analyst controls.
The original 28 contracts retain their API and registry records. Binary
segmentation is included below with its public-initialisation evidence and
separately labelled random-initialisation boundary comparisons.

## Two subject likelihoods

The AFT contract uses a scalar location predictor `mu(x)` and a fixed public
Weibull shape or log-normal sigma. `dispersion` is one of `0.5`, `1`, `2`.
The location is bounded to [-10,10], and incompatible public time/dispersion
ranges fail before private data is accessed. Weibull uses its full event or
censor likelihood; log-normal censor likelihood uses stable `log_ndtr`.
Reported event density includes the original-time Jacobian.

```{r aft, eval=FALSE}
model <- ds.flower.model.pytorch_aft(
  horizon = 1825, time_scale = 365,
  distribution = "lognormal", dispersion = 1,
  hidden_layers = integer(0))
fit <- ds.flower.fit(
  conns, symbol = "D", target = c("time", "event"),
  features = c("age", "baseline_covariate"), model = model,
  feature_bounds = list(lower = c(0, 0), upper = c(100, 1)),
  rounds = 10)
```

Time is measured in days from the declared baseline. Event coding is exactly
1=event, 0=right censoring. Finite times above the public horizon are
administratively censored at the horizon; an event exactly at the horizon
remains an event. Times below the public minimum resolution, nonfinite times,
invalid event codes and duplicate subject records become safe placeholders
with `valid=0`. They are never removed from the privacy census or denominator.
No arithmetic averaging of survival outcomes occurs.

The hazard contract maps a covariate vector to K logits. K is at most 64.
Intervals are `(left, right]`: an event contributes through its event interval;
a censor contributes only completely observed intervals. Censoring inside an
interval contributes no observed non-event for that incomplete interval.

```{r hazard, eval=FALSE}
model <- ds.flower.model.pytorch_discrete_hazard(
  edges = c(0, 7, 14, 21, 30, 45, 60, 90, 120,
            180, 270, 365, 540, 730, 1095, 1460, 1825))
```

For subject i, the loss is `valid_i * sum(mask_ij * BCE(logit_ij,event_ij))/K`.
The batch loss is the mean over every sampled subject, including invalid
subjects' zero contributions. A tensor item contains the whole period vector;
Opacus sums its period effects and clips that subject's complete parameter
gradient once. Periods are never flattened into independent training examples.

## Censuses and privacy accounting

The source manifest retains M source rows and N subjects. A separate local,
server-owned artifact holds N derived subject vectors. Both censuses and the
exact source-to-subject transformation are checked before training. Invalid
subjects do not reduce N. The accountant still uses `q=1/ceiling(N/batch_size)`
and the complete pinned round/epoch horizon; K never multiplies N.

Replace-one conversion, secure Poisson draws, secure Gaussian noise, empty
noisy steps and fixed client weights remain unchanged. Distribution, dispersion,
interval edges, validity and effective transformed tensors are bound into
sticky semantics. Repeating an equivalent release does not generate a new
noise draw on the same deterministic software/device profile. Separate models,
splits or experiments on overlapping real subjects compose.

Cox partial likelihood remains excluded because one subject's risk set can
change another subject's gradient. Calling its event terms individual losses
does not make them subject-separable.

## Prediction and channel-B evaluation

Released artifacts retain the public model specification and survival settings.
Local prediction reconstructs the same head and transformation:

```{r prediction, eval=FALSE}
risk <- ds.flower.predict(fit, public_test_features, type="risk")
curves <- ds.flower.predict(fit, public_test_features, type="survival",
                            times=c(30, 180, 365))
median <- ds.flower.predict(fit, public_test_features, type="median")
```

AFT risk is `-mu`; hazard risk is negative restricted mean survival over the
public grid using left-endpoint survival. Hazard curves may cross. Curves are
constant between hazard interval ends; a hazard median beyond the horizon is
unavailable. AFT medians follow the selected public distribution.

Concordance and held-out NLL are evaluated only on public or independently
authorized analyst-local data. The campaign excludes observed-time ties and
awards half credit to exact risk ties. Its likelihood score uses original-time
AFT density or hazard NLL divided by K. These are different likelihood scales;
compare each contract with its own exact twin.

`ds.flower.validate`, private holdout, cross-validation and survival training
inside HPO reject at public preflight. No private training losses, concordance,
event counts or validity summaries are returned. This work does not implement
competing risks, delayed entry, interval censoring or time-varying covariates.

## Public evidence and preregistration

The protocol is distributed at
`tools/campaign/survival/PROTOCOL_F_SURVIVAL.md`. It fixes subject splits before
any scoring, a baseline-only covariate subset, public bounds, three seeds,
three isolated DSLite sites, epsilon 1/4/8, delta 1e-5 and clip 1. SUPPORT2 is the
primary cohort; LUNG1 is the small secondary cohort. Source hashes, split hashes,
versions, exact commits, runner hashes and mechanism details accompany executed
JSONs under `inst/extdata/campaign/survival/`.

[SUPPORT2 (UCI 880)](https://archive-beta.ics.uci.edu/dataset/880/support2)
lists CC BY 4.0. [The source description](https://hbiostat.org/data/repo/supportdesc)
identifies day-3 assessments; this campaign excludes them and outcome-derived
fields. [TCIA LUNG1 clinical data](https://www.cancerimagingarchive.net/collection/nsclc-radiomics/)
is CC BY-NC 3.0 and measures survival from treatment start. The clinical file
has noncommercial terms; it is not a commercial-use benchmark.

The preregistered primary floor is mean SUPPORT2 C-index at epsilon 8 at least
0.60 and at least null +0.05, assessed separately for each model. Failure is
reported as failure. All four historical utility-envelope checks are empirical
diagnostics, not privacy proofs. Here shortfall is `G=central-federated`, the
reverse sign of the older article's delta. Confidence intervals describe
variation over matched public split replicates; overlapping splits and only
three replicates limit their interpretation.

Log-normal AFT is the introductory example because the reviewer's exploratory
pooled probes found useful ranking and better held-out NLL than their
covariate-free null. Those probes are not release evidence. Their Weibull arms
(including nonprivate training) and hazard DP arms had worse NLL than their
nulls under the fixed schedules. Concordance measures ranking, not probability
calibration; inspect held-out NLL alongside concordance in the executed matrix.
The current summary retains v1 AFT results and uses the selected **h06**
discrete-hazard schedule from the preregistered v2 confirmation: 10 rounds,
4 local epochs, batch size 64, learning rate 0.05, and 10 intervals. The
selection record, preregistration, driver completion and all 30 confirmation
cells are packaged under `survival/hazard-v2/`. The original v1 cells and
`survival/v1/summary.json` remain boundary evidence. Schedule development and
confirmation are distinguished in those records. The v2 confirmation reuses
the public outer holdouts, so it is not independent external validation.
LUNG1 v2 ran on pod CUDA whereas its v1 evidence used CPU; those schedules
and device profiles must not be treated as a controlled schedule-only contrast.
Failed utility floors remain reported, and reviewer promotion is separate
from execution.
For the executed matrix, lower NLL is better only relative to the matching null
and likelihood family: a C-index floor or PASS does not establish calibrated
probabilities. The rendered table therefore retains every exact twin's NLL and
labels whether federated-DP NLL is worse than its matched null.

```{r evidence-census, echo=FALSE}
cat(length(executed), "executed cells and", length(failed),
    "failed attempts are archived.\n")
summary_records <- Filter(function(x) identical(x$record_type, "summary"), evidence)
if (length(summary_records)) {
  groups <- Filter(function(x) identical(x$status, "executed"), summary_records[[1]]$groups)
  mean_sd <- function(x) sprintf("%.3f ± %.3f", x$mean, x$sd)
  rows <- list()
  for (group in groups) for (epsilon in names(group$envelopes$summaries)) {
    x <- group$envelopes$summaries[[epsilon]]
    floor <- group$envelopes$utility_floor
    rows[[length(rows)+1L]] <- data.frame(
      cohort=group$dataset, subset=group$subset, model=group$variant,
      epsilon=epsilon,
      C_federated_DP=mean_sd(x$federated), C_pooled_DP=mean_sd(x$central_dp),
      C_pooled_nonprivate=mean_sd(x$central), C_null=mean_sd(x$null),
      NLL_federated_DP=mean_sd(x$heldout_nll$federated_dp),
      NLL_pooled_DP=mean_sd(x$heldout_nll$central_dp),
      NLL_pooled_nonprivate=mean_sd(x$heldout_nll$central),
      NLL_null=mean_sd(x$heldout_nll$null),
      federated_NLL_vs_null=if (x$heldout_nll$federated_dp$mean > x$heldout_nll$null$mean)
        "worse" else "not worse",
      shortfall=x$gap$mean,
      floor=if (is.null(floor) || epsilon!="8") "not designated" else
        if (isTRUE(floor$pass)) "PASS" else "FAIL")
  }
  if (length(rows)) knitr::kable(do.call(rbind,rows),digits=3,
                                 caption="Public held-out scores, mean ± SD over matched seeds")
}
if (length(failed)) {
  knitr::kable(do.call(rbind,lapply(failed,function(x) data.frame(
    cohort=x$dataset$dataset,model=x$variant,status=x$status,
    reason=x$error))))
}
```

Only archived executed cells support a measured-utility statement. A failed
attempt has no substituted score. Promotion remains a reviewer decision after
all mathematical, gradient, accounting, authority, staging, sticky and API
compatibility gates pass.

## Binary segmentation: current public initialisation and retained boundaries

The contract uses one selected image per subject, a frozen ImageNet ResNet-18
encoder, canonical 128 by 128 masks, and a subject-separable BCE/Dice loss.
Invalid subjects remain in the privacy census with zero loss validity.
Private validation, holdout, cross-validation and private-score HPO are
unsupported. The [binary segmentation guide](binary-segmentation.html) describes
the complete schema, prediction mapping and fail-closed boundaries.

Current results come from **protocol v5**, whose SHA256 begins `28a713f8`.
The selected public-initialisation route uses a narrow 9,521-parameter decoder,
60 public pretraining epochs on BUSI, 20 federated rounds and nominal batch 16.
Batch 64 is a sensitivity arm. BUSI is the public dataset described by
Al-Dhabyani et al. (2020), obtained through its Kaggle mirror; the packaged
`public-pretraining/audit.json`, per-seed records and
`public-pretraining-manifest.json` retain dataset, checkpoint and audit digests.
The campaign's audited benchmark binding loads public decoder tensors before
private training. These results do not describe the default random decoder
initialisation or a general analyst checkpoint-upload API.

The original packaged batch files were exact v3 copies. `RECONCILE_V5.md` and
`audit-results.json` record the stale-file finding and fresh artifact-inference
checks. Current `batch16/` and `batch64/` records now contain the actual v5
results; `retained/v3/` and `retained/v4/` retain random-initialisation boundary evidence. The
comparison below labels each protocol explicitly and never pools their seeds.
V4 and v5 selected the same narrow decoder and fine-tuning schedule; v5 adds
the disclosed BUSI pretraining stage. V3 uses a different decoder and schedule.
All comparisons retain the public-development and reused-test limitations.

```{r segmentation-release-evidence, echo=FALSE, error=FALSE}
seg_dir <- file.path(dirname(evidence_dir), "segmentation")
protocol_hash <- digest::digest(file=file.path(seg_dir, "protocol-v5.md"),
                                algo="sha256")
stopifnot(startsWith(protocol_hash, "28a713f8"))
seg_rows <- list()
add_seg <- function(path, version, cohort, subset, epsilons, batch, digits=3L) {
  record <- jsonlite::read_json(file.path(seg_dir, path), simplifyVector=FALSE)
  stopifnot(identical(record$status, "executed"))
  if (identical(version, "v5 public-init")) {
    stopifnot(identical(record$protocol_sha256, protocol_hash))
  }
  for (epsilon in epsilons) {
    cells <- Filter(function(x) identical(x$variant, subset) && x$epsilon == epsilon,
                    c(record$replicates, record$additional_replicates))
    stopifnot(length(cells) == 3L,
              length(unique(vapply(cells, function(x) x$seed, numeric(1)))) == 3L)
    values <- vapply(cells, function(x) x$federated_dp$foreground_positive$dice,
                     numeric(1))
    stopifnot(all(is.finite(values)))
    seg_rows[[length(seg_rows)+1L]] <<- data.frame(
      protocol=version, cohort=cohort, subset=subset, batch=batch,
      epsilon=epsilon, seeds=length(values),
      foreground_Dice=sprintf(paste0("%.", digits, "f"), mean(values)),
      SD=sprintf("%.3f", stats::sd(values)))
  }
}
add_seg("retained/v4/batch16/busbra-evidence.json", "v4 random-init boundary",
        "BUS-BRA", "full", c(1,4,8), 16, digits=2L)
add_seg("batch16/busbra-evidence.json", "v5 public-init", "BUS-BRA", "full", c(1,4,8), 16)
add_seg("batch16/busbra-evidence.json", "v5 public-init", "BUS-BRA", "heterogeneous", 8, 16)
add_seg("batch16/busbra-evidence.json", "v5 public-init", "BUS-BRA", "small192", 8, 16)
add_seg("batch16/breast-evidence.json", "v5 public-init", "BrEaST", "full", 8, 16)
add_seg("batch64/busbra-evidence.json", "v5 public-init", "BUS-BRA", "full", 8, 64)
knitr::kable(do.call(rbind, seg_rows), row.names=FALSE,
             caption="Federated-DP foreground-positive subject Dice: three-seed means")
```

These are public held-out metrics at threshold 0.5, with foreground-positive
and empty-mask strata kept separate in the full records. The selected v5
route clears its recorded utility floor, while the random-initialisation
boundaries remain visible. Execution and utility floors do not establish
calibration, clinical validity or reviewer acceptance. The segmentation
contract remains `vetted=FALSE`.
