---
title: "Subject-level survival contracts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Subject-level survival contracts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
evidence_dir <- file.path("..", "inst", "extdata", "campaign", "survival")
if (!dir.exists(evidence_dir)) {
  evidence_dir <- system.file("extdata", "campaign", "survival", package="dsFlowerClient")
}
evidence_files <- list.files(evidence_dir, pattern="\\.json$", full.names=TRUE)
evidence <- lapply(evidence_files, jsonlite::read_json, simplifyVector=FALSE)
cells <- Filter(function(x) identical(x$record_type, "cell") ||
  (!is.null(x$variant) && !is.null(x$status)), evidence)
executed <- Filter(function(x) identical(x$status, "executed"), cells)
failed <- Filter(function(x) identical(x$status, "failed"), cells)
```

`pytorch_aft` and `pytorch_discrete_hazard` train baseline survival predictors
under the existing node-enforced DP mechanism. The custodian must configure
patient privacy and a stable subject identifier. Epsilon, delta, clipping,
privacy-unit selection and the accounting population are not analyst controls.
The original 28 contracts retain their API and registry records. Segmentation
is developed separately; this article reserves a section for that work.

## Two subject likelihoods

The AFT contract uses a scalar location predictor `mu(x)` and a fixed public
Weibull shape or log-normal sigma. `dispersion` is one of `0.5`, `1`, `2`.
The location is bounded to [-10,10], and incompatible public time/dispersion
ranges fail before private data is accessed. Weibull uses its full event or
censor likelihood; log-normal censor likelihood uses stable `log_ndtr`.
Reported event density includes the original-time Jacobian.

```{r aft, eval=FALSE}
model <- ds.flower.model.pytorch_aft(
  horizon = 1825, time_scale = 365,
  distribution = "weibull", dispersion = 1,
  hidden_layers = integer(0))
fit <- ds.flower.fit(
  conns, symbol = "D", target = c("time", "event"),
  features = c("age", "baseline_covariate"), model = model,
  feature_bounds = list(lower = c(0, 0), upper = c(100, 1)),
  rounds = 10)
```

Time is measured in days from the declared baseline. Event coding is exactly
1=event, 0=right censoring. Finite times above the public horizon are
administratively censored at the horizon; an event exactly at the horizon
remains an event. Times below the public minimum resolution, nonfinite times,
invalid event codes and duplicate subject records become safe placeholders
with `valid=0`. They are never removed from the privacy census or denominator.
No arithmetic averaging of survival outcomes occurs.

The hazard contract maps a covariate vector to K logits. K is at most 64.
Intervals are `(left, right]`: an event contributes through its event interval;
a censor contributes only completely observed intervals. Censoring inside an
interval contributes no observed non-event for that incomplete interval.

```{r hazard, eval=FALSE}
model <- ds.flower.model.pytorch_discrete_hazard(
  edges = c(0, 7, 14, 21, 30, 45, 60, 90, 120,
            180, 270, 365, 540, 730, 1095, 1460, 1825))
```

For subject i, the loss is `valid_i * sum(mask_ij * BCE(logit_ij,event_ij))/K`.
The batch loss is the mean over every sampled subject, including invalid
subjects' zero contributions. A tensor item contains the whole period vector;
Opacus sums its period effects and clips that subject's complete parameter
gradient once. Periods are never flattened into independent training examples.

## Censuses and privacy accounting

The source manifest retains M source rows and N subjects. A separate local,
server-owned artifact holds N derived subject vectors. Both censuses and the
exact source-to-subject transformation are checked before training. Invalid
subjects do not reduce N. The accountant still uses `q=1/ceiling(N/batch_size)`
and the complete pinned round/epoch horizon; K never multiplies N.

Replace-one conversion, secure Poisson draws, secure Gaussian noise, empty
noisy steps and fixed client weights remain unchanged. Distribution, dispersion,
interval edges, validity and effective transformed tensors are bound into
sticky semantics. Repeating an equivalent release does not generate a new
noise draw on the same deterministic software/device profile. Separate models,
splits or experiments on overlapping real subjects compose.

Cox partial likelihood remains excluded because one subject's risk set can
change another subject's gradient. Calling its event terms individual losses
does not make them subject-separable.

## Prediction and channel-B evaluation

Released artifacts retain the public model specification and survival settings.
Local prediction reconstructs the same head and transformation:

```{r prediction, eval=FALSE}
risk <- ds.flower.predict(fit, public_test_features, type="risk")
curves <- ds.flower.predict(fit, public_test_features, type="survival",
                            times=c(30, 180, 365))
median <- ds.flower.predict(fit, public_test_features, type="median")
```

AFT risk is `-mu`; hazard risk is negative restricted mean survival over the
public grid using left-endpoint survival. Hazard curves may cross. Curves are
constant between hazard interval ends; a hazard median beyond the horizon is
unavailable. AFT medians follow the selected public distribution.

Concordance and held-out NLL are evaluated only on public or independently
authorized analyst-local data. The campaign excludes observed-time ties and
awards half credit to exact risk ties. Its likelihood score uses original-time
AFT density or hazard NLL divided by K. These are different likelihood scales;
compare each contract with its own exact twin.

`ds.flower.validate`, private holdout, cross-validation and survival training
inside HPO reject at public preflight. No private training losses, concordance,
event counts or validity summaries are returned. This work does not implement
competing risks, delayed entry, interval censoring or time-varying covariates.

## Public evidence and preregistration

The protocol is distributed at
`tools/campaign/survival/PROTOCOL_F_SURVIVAL.md`. It fixes subject splits before
any scoring, a baseline-only covariate subset, public bounds, three seeds,
three isolated DSLite sites, epsilon 1/4/8, delta 1e-5 and clip 1. SUPPORT2 is the
primary cohort; LUNG1 is the small secondary cohort. Source hashes, split hashes,
versions, exact commits, runner hashes and mechanism details accompany executed
JSONs under `inst/extdata/campaign/survival/`.

[SUPPORT2 (UCI 880)](https://archive-beta.ics.uci.edu/dataset/880/support2)
lists CC BY 4.0. [The source description](https://hbiostat.org/data/repo/supportdesc)
identifies day-3 assessments; this campaign excludes them and outcome-derived
fields. [TCIA LUNG1 clinical data](https://www.cancerimagingarchive.net/collection/nsclc-radiomics/)
is CC BY-NC 3.0 and measures survival from treatment start. The clinical file
has noncommercial terms; it is not a commercial-use benchmark.

The preregistered primary floor is mean SUPPORT2 C-index at epsilon 8 at least
0.60 and at least null +0.05, assessed separately for each model. Failure is
reported as failure. All four historical utility-envelope checks are empirical
diagnostics, not privacy proofs. Here shortfall is `G=central-federated`, the
reverse sign of the older article's delta. Confidence intervals describe
variation over matched public split replicates; overlapping splits and only
three replicates limit their interpretation.

```{r evidence-census, echo=FALSE}
cat(length(executed), "executed cells and", length(failed),
    "failed attempts are archived.\n")
summary_records <- Filter(function(x) identical(x$record_type, "summary"), evidence)
if (length(summary_records)) {
  groups <- Filter(function(x) identical(x$status, "executed"), summary_records[[1]]$groups)
  rows <- list()
  for (group in groups) for (epsilon in names(group$envelopes$summaries)) {
    x <- group$envelopes$summaries[[epsilon]]
    floor <- group$envelopes$utility_floor
    rows[[length(rows)+1L]] <- data.frame(
      cohort=group$dataset, subset=group$subset, model=group$variant,
      epsilon=epsilon, C_index=x$federated$mean,
      CI95=paste(round(unlist(x$federated$ci95),3),collapse=" to "),
      heldout_NLL=x$heldout_nll$federated_dp$mean,
      NLL_CI95=paste(round(unlist(x$heldout_nll$federated_dp$ci95),3),collapse=" to "),
      shortfall=x$gap$mean,
      floor=if (is.null(floor) || epsilon!="8") "not designated" else
        if (isTRUE(floor$pass)) "PASS" else "FAILED")
  }
  if (length(rows)) knitr::kable(do.call(rbind,rows),digits=3)
}
if (length(failed)) {
  knitr::kable(do.call(rbind,lapply(failed,function(x) data.frame(
    cohort=x$dataset$dataset,model=x$variant,status=x$status,
    reason=x$error))))
}
```

Only archived executed cells support a measured-utility statement. A failed
attempt has no substituted score. Promotion remains a reviewer decision after
all mathematical, gradient, accounting, authority, staging, sticky and API
compatibility gates pass.

## Segmentation — reserved for the parallel implementation

**This section is intentionally reserved.** This survival increment does not
register or implement a segmentation contract and makes no segmentation
utility claim. The parallel implementation will supply its contract, subject
selection, mask semantics, evidence and limitations here.
