---
title: "All Method Validation Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{All Method Validation Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
method_result_path <- method_paths[nzchar(method_paths) & file.exists(method_paths)][1]
method_validation <- jsonlite::fromJSON(method_result_path)
method_results <- method_validation$results
vision_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
vision_result_path <- vision_paths[nzchar(vision_paths) & file.exists(vision_paths)][1]
vision_validation <- if (!is.na(vision_result_path)) jsonlite::fromJSON(vision_result_path) else NULL
vision_results <- if (!is.null(vision_validation)) vision_validation$results else NULL
family_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_family_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_family_results.json'),
  system.file('extdata', 'dsflower_method_family_results.json', package = 'dsFlowerClient')
)
family_result_path <- family_paths[nzchar(family_paths) & file.exists(family_paths)][1]
family_validation <- if (!is.na(family_result_path)) jsonlite::fromJSON(family_result_path) else NULL
family_results <- if (!is.null(family_validation)) family_validation$results else NULL
column_or <- function(frame, name, default) {
  if (name %in% names(frame)) frame[[name]] else rep(default, nrow(frame))
}
method_display <- data.frame(
  suite = 'tabular_sequence_survival',
  method = method_results$method,
  task = method_results$task,
  centralized_metric = method_results$centralized_metric,
  centralized_loss = method_results$centralized_loss,
  federated_status = method_results$federated_status,
  federated_loss = method_results$federated_loss,
  delta_loss = method_results$delta_loss,
  acceptable_loss = method_results$acceptable_loss,
  validation_status = method_results$validation_status,
  stringsAsFactors = FALSE
)
vision_display <- if (!is.null(vision_results)) data.frame(
  suite = 'vision',
  method = vision_results$method,
  task = vision_results$task,
  centralized_metric = vision_results$centralized_metric,
  centralized_loss = vision_results$centralized_loss,
  federated_status = vision_results$federated_status,
  federated_loss = vision_results$federated_loss,
  delta_loss = vision_results$delta_loss,
  acceptable_loss = column_or(vision_results, 'acceptable_loss',
                              vision_results$validation_status == 'pass'),
  validation_status = vision_results$validation_status,
  stringsAsFactors = FALSE
) else method_display[0, ]
family_display <- if (!is.null(family_results)) data.frame(
  suite = 'clinical_support2_secagg',
  method = family_results$method,
  task = family_results$task,
  centralized_metric = family_results$centralized_metric,
  centralized_loss = family_results$centralized_loss,
  federated_status = family_results$federated_status,
  federated_loss = family_results$federated_loss,
  delta_loss = family_results$delta_loss,
  acceptable_loss = family_results$acceptable_loss,
  validation_status = family_results$validation_status,
  stringsAsFactors = FALSE
) else method_display[0, ]
results <- rbind(method_display, vision_display, family_display)
```

This macro-vignette summarizes the live dsFlower validation suite across
tabular, sequence, survival, XGBoost, image-classification, and segmentation
templates. Each method has its own vignette. The core method-validation suite
uses deterministic synthetic fixtures for broad template coverage; the clinical
benchmarks add real public datasets for the thesis-facing evidence.

The suite is deliberately a functional and numerical sanity check, not a
privacy claim and not, by itself, a clinical benchmark. It uses `sandbox_open`
so that per-node diagnostics can be observed during development. Production
studies should use the stricter DataSHIELD privacy profiles. The SUPPORT2
method-family benchmark and the clinical privacy benchmark use
`clinical_default` or `high_sensitivity_dp`, so Secure Aggregation and the
profile-level disclosure guards are active.

```{r}
summary_table <- data.frame(
  suite = c('tabular_sequence_survival', 'vision', 'clinical_support2_secagg'),
  generated_at = c(
    method_validation$generated_at,
    if (!is.null(vision_validation)) vision_validation$generated_at else NA,
    if (!is.null(family_validation)) family_validation$generated_at else NA
  ),
  privacy_profile = c(
    method_validation$privacy_profile,
    if (!is.null(vision_validation)) vision_validation$privacy_profile else NA,
    if (!is.null(family_validation)) family_validation$privacy_profile else NA
  ),
  n_sites = c(
    method_validation$n_sites,
    if (!is.null(vision_validation)) vision_validation$n_sites else NA,
    if (!is.null(family_validation)) family_validation$n_sites else NA
  ),
  n_total = c(
    method_validation$n_total,
    if (!is.null(vision_validation)) vision_validation$n_total else NA,
    if (!is.null(family_validation)) family_validation$n_total else NA
  ),
  secagg_supported = c(
    method_validation$secagg_supported,
    NA,
    if (!is.null(family_validation)) family_validation$secagg_supported else NA
  ),
  stringsAsFactors = FALSE
)
knitr::kable(summary_table)
```

```{r}
knitr::kable(results)
```

```{r, fig.width=8, fig.height=5, fig.alt="Horizontal bar chart of federated minus centralized loss by method."}
plot_df <- results[is.finite(results$delta_loss), , drop = FALSE]
if (nrow(plot_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  plot_df$label <- paste(plot_df$suite, plot_df$method, sep = ': ')
  ggplot2::ggplot(plot_df, ggplot2::aes(x = stats::reorder(label, delta_loss), y = delta_loss, fill = task)) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = 'grey40') +
    ggplot2::geom_col(width = 0.7) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = 'Federated loss - centralized loss') +
    ggplot2::theme_minimal(base_size = 11)
}
```

Per-method vignettes:

- [clinical method-family benchmarks](clinical-method-family-benchmarks.html)

- [sklearn_logreg](validation-sklearn-logreg.html)
- [sklearn_ridge](validation-sklearn-ridge.html)
- [sklearn_sgd](validation-sklearn-sgd.html)
- [sklearn_svm](validation-sklearn-svm.html)
- [sklearn_elastic_net](validation-sklearn-elastic-net.html)
- [pytorch_logreg](validation-pytorch-logreg.html)
- [pytorch_mlp](validation-pytorch-mlp.html)
- [pytorch_linear_regression](validation-pytorch-linear-regression.html)
- [pytorch_multiclass](validation-pytorch-multiclass.html)
- [pytorch_multilabel](validation-pytorch-multilabel.html)
- [pytorch_poisson](validation-pytorch-poisson.html)
- [pytorch_coxph](validation-pytorch-coxph.html)
- [pytorch_lognormal_aft](validation-pytorch-lognormal-aft.html)
- [pytorch_cause_specific_cox](validation-pytorch-cause-specific-cox.html)
- [pytorch_lstm](validation-pytorch-lstm.html)
- [pytorch_tcn](validation-pytorch-tcn.html)
- [xgboost](validation-xgboost.html)

Vision method vignettes:

- [pytorch_resnet18](validation-vision-pytorch-resnet18.html)
- [pytorch_densenet121](validation-vision-pytorch-densenet121.html)
- [pytorch_unet2d](validation-vision-pytorch-unet2d.html)

The vision-specific overview remains available at
[Vision Method Validation Overview](validation-vision-overview.html).

The full suite is produced by repeating the inline Opal/DataSHIELD/Flower
workflow shown in each per-method vignette for every method listed above.
The persisted JSON artifacts are the audit trail consumed by this
pkgdown site.
