---
title: "Validation: sklearn_ridge"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation: sklearn_ridge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "sklearn_ridge"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records the catalogue coverage check for this authorised
template. The federated run used three Opal/Rock nodes and the same pooled
synthetic validation cohort was used for the local baseline. The purpose is
method-path validation: data staging, template verification, Flower
execution, result persistence, and cleanup. It complements the biomedical
reports by checking the full template surface under deterministic
conditions.

It is laid out in two parts: first the commands you would run to
reproduce the validation, then the recorded results of that run loaded
from the committed evidence artifact. No live Opal servers are required to
render this page; the command chunks are shown for reference and are not
executed.

## Running This Validation

To check this template we compare a **centralized** baseline against the
**federated** dsFlower run. The validation cohort is one synthetic dataset:
the centralized baseline is fit on the pooled copy, and the very same rows
are partitioned across three local Opal nodes that are already set up and
serving the table `dsflower_demo.validation_cohort`. The federated fit never
sees the pooled data, only the three node-resident partitions.

**1. Connect to the DataSHIELD nodes.** Open a single session across the
three Opal nodes that hold the partitioned validation cohort.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = 'node1', url = 'https://opal-node-1.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
builder$append(server = 'node2', url = 'https://opal-node-2.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
builder$append(server = 'node3', url = 'https://opal-node-3.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = 'D')
```

**2. Run the federated fit.** Train this template through dsFlower with
FedAvg over the server-side table, then log out. Only aggregate training
history and model parameters are returned to the researcher.

```{r federated-fit, eval=FALSE}
fit <- ds.flower.fit(
  conns,
  symbol   = 'D',
  target   = 'outcome',
  features = paste0('x', 1:12),
  model    = 'sklearn_ridge',
  strategy = 'fedavg',
  privacy  = 'sandbox_open',
  rounds   = 1
)

DSI::datashield.logout(conns)
```

## Recorded Results

The values below come from the committed evidence artifact for the run
described above; no live servers are contacted when the article renders.

This run trained the `r row$method` template on the `r row$task` task
(target `r row$target`, `r row$n_features` synthetic features) for
`r row$rounds` federated round(s) under the `r validation$privacy_profile`
profile, across `r validation$n_sites` Opal/Rock nodes holding the same
pooled cohort used for the centralized baseline.

The centralized baseline reached a `r row$centralized_metric` of
`r fmt(row$centralized_loss)`; the federated run reached
`r fmt(row$federated_loss)` with `r row$federated_n_failures` client
failures, leaving this template **`r row$validation_status`**. The
acceptance comparison is shown below, followed by the centralized vs.
federated loss.

```{r acceptance-output}
max_allowed_loss <- row$centralized_loss * row$acceptance_max_loss_ratio +
  row$acceptance_max_loss_margin
acceptance_output <- data.frame(
  centralized_loss = row$centralized_loss,
  federated_loss = row$federated_loss,
  max_allowed_loss = max_allowed_loss,
  federated_n_failures = row$federated_n_failures,
  acceptable_loss = row$acceptable_loss,
  validation_status = row$validation_status
)
knitr::kable(acceptance_output, digits = 4)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for sklearn_ridge."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```
