---
title: "Validation: sklearn_logreg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation: sklearn_logreg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "sklearn_logreg"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records the catalogue coverage check for this authorised
template. The federated run used three Opal/Rock nodes and the same pooled
synthetic validation cohort was used for the local baseline. The purpose is
method-path validation: data staging, template verification, Flower
execution, result persistence, and cleanup. It complements the biomedical
reports by checking the full template surface under deterministic
conditions.

It is laid out as a short tutorial: first the commands you would run to
reproduce the validation, then the recorded results of that run loaded
from the committed evidence artifact. No live Opal servers are required to
render this page; the command chunks are shown for reference and are not
executed.

## Tutorial: How To Run This Validation

To check this template we compare a **centralized** baseline against the
**federated** dsFlower run. The validation cohort is one synthetic dataset:
the centralized baseline is fit on the pooled copy, and the very same rows
are partitioned across three local Opal nodes that are already set up and
serving the table `dsflower_demo.validation_cohort`. The federated fit never
sees the pooled data, only the three node-resident partitions.

**1. Connect to the DataSHIELD nodes.** Open a single session across the
three Opal nodes that hold the partitioned validation cohort.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = 'node1', url = 'https://opal-node-1.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
builder$append(server = 'node2', url = 'https://opal-node-2.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
builder$append(server = 'node3', url = 'https://opal-node-3.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.validation_cohort')
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = 'D')
```

**2. Run the federated fit.** Train this template through dsFlower with
FedAvg over the server-side table, then log out. Only aggregate training
history and model parameters are returned to the researcher.

```{r federated-fit, eval=FALSE}
fit <- ds.flower.fit(
  conns,
  symbol   = 'D',
  target   = 'outcome',
  features = paste0('x', 1:12),
  model    = 'sklearn_logreg',
  strategy = 'fedavg',
  privacy  = 'sandbox_open',
  rounds   = 1
)

DSI::datashield.logout(conns)
```

## Recorded Results

The remaining chunks are evaluated when pkgdown renders the article. They
load the committed evidence artifact for the run above and report the
centralized vs. federated comparison, the audit trail, and the acceptance
decision.

```{r}
overview <- data.frame(
  Field = c(
    'method', 'task', 'target', 'n_features', 'rounds',
    'centralized_metric', 'centralized_loss', 'centralized_accuracy',
    'federated_status', 'federated_loss', 'federated_n_failures',
    'delta_loss', 'acceptance_max_loss_ratio',
    'acceptance_max_loss_margin', 'acceptable_loss',
    'validation_status'
  ),
  Value = c(
    row$method, row$task, row$target, row$n_features, row$rounds,
    row$centralized_metric, fmt(row$centralized_loss), fmt(row$centralized_accuracy),
    row$federated_status, fmt(row$federated_loss), fmt(row$federated_n_failures),
    fmt(row$delta_loss), fmt(row$acceptance_max_loss_ratio),
    fmt(row$acceptance_max_loss_margin), row$acceptable_loss,
    row$validation_status
  )
)
knitr::kable(overview)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for sklearn_logreg."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```

### Step-By-Step Run Output

The following chunks print the audit trail for the committed catalogue
coverage run, reconstructed from the persisted evidence artifact.

```{r step-inputs}
target_cols <- trimws(strsplit(row$target, ',', fixed = TRUE)[[1]])
site_table_paths <- validation$table_paths
input_trace <- data.frame(
  step = c(
    '1. Build validation cohort',
    '2. Split across sites',
    '3. Select task target',
    '4. Select model features',
    '5. Materialize Opal tables'
  ),
  observed_output = c(
    paste(validation$n_total, 'synthetic rows generated for this suite'),
    paste(validation$n_sites, 'sites x', validation$n_per_site, 'rows per site'),
    paste(target_cols, collapse = ', '),
    paste(row$n_features, 'features:', paste0('x1..x', row$n_features)),
    paste(site_table_paths, collapse = '\n')
  ),
  stringsAsFactors = FALSE
)
knitr::kable(input_trace)
```

```{r centralized-output}
centralized_output <- data.frame(
  step = c(
    '1. Load pooled validation cohort',
    '2. Train centralized baseline',
    '3. Evaluate centralized baseline'
  ),
  observed_output = c(
    paste(validation$n_total, 'pooled rows with', row$n_features, 'features'),
    paste('status =', row$centralized_status, '| metric =', row$centralized_metric),
    paste('loss =', fmt(row$centralized_loss),
          '| accuracy =', fmt(row$centralized_accuracy))
  ),
  stringsAsFactors = FALSE
)
knitr::kable(centralized_output)
```

```{r centralized-console}
cat(
  sprintf('[centralized] method: %s\n', row$method),
  sprintf('[centralized] task: %s | target: %s | features: %s\n',
          row$task, row$target, paste0('x1..x', row$n_features)),
  sprintf('[centralized] status: %s\n', row$centralized_status),
  sprintf('[centralized] %s: %.4f\n', row$centralized_metric, row$centralized_loss),
  sprintf('[centralized] accuracy: %s\n', fmt(row$centralized_accuracy)),
  sep = ''
)
```

```{r dsflower-output}
max_allowed_loss <- row$centralized_loss * row$acceptance_max_loss_ratio +
  row$acceptance_max_loss_margin
dsflower_output <- data.frame(
  step = c(
    '1. Login through DataSHIELD',
    '2. Prepare server-side dsFlower nodes',
    '3. Start Flower SuperLink/SuperNodes',
    '4. Run federated training',
    '5. Collect metrics and cleanup'
  ),
  observed_output = c(
    paste(validation$n_sites, 'Opal/Rock nodes connected'),
    paste('template =', row$method, '| privacy =', validation$privacy_profile),
    paste('federated_run_id =', row$federated_run_id),
    paste('rounds =', row$rounds, '| federated_loss =', fmt(row$federated_loss),
          '| client_failures =', fmt(row$federated_n_failures)),
    paste('federated_status =', row$federated_status,
          '| validation_status =', row$validation_status)
  ),
  stringsAsFactors = FALSE
)
knitr::kable(dsflower_output)
```

```{r dsflower-console}
cat(
  sprintf('[dsFlower] DataSHIELD login: %s nodes\n', validation$n_sites),
  sprintf('[dsFlower] table paths: %s\n', paste(site_table_paths, collapse = ' | ')),
  sprintf('[dsFlower] model: %s | strategy: fedavg | rounds: %s\n',
          row$method, row$rounds),
  sprintf('[dsFlower] run id: %s\n', row$federated_run_id),
  sprintf('[Flower] aggregate_fit/evaluate: clients=%s failures=%s\n',
          validation$n_sites, fmt(row$federated_n_failures)),
  sprintf('[Flower] final federated loss: %.4f\n', row$federated_loss),
  sprintf('[cleanup] status: %s\n', row$federated_status),
  sep = ''
)
```

```{r acceptance-output}
acceptance_output <- data.frame(
  centralized_loss = row$centralized_loss,
  federated_loss = row$federated_loss,
  max_allowed_loss = max_allowed_loss,
  federated_n_failures = row$federated_n_failures,
  acceptable_loss = row$acceptable_loss,
  validation_status = row$validation_status
)
knitr::kable(acceptance_output, digits = 4)
```

```{r}
if (nzchar(row$error)) cat(row$error)
```
