---
title: "Method Validation Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Method Validation Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
```

This macro-vignette summarizes the live dsFlower method-validation run.
The experiment created one synthetic cohort, split it across three Opal/Rock
servers, trained each supported federated template, and compared the final
federated loss with a centralized baseline trained on the pooled cohort.

The suite is deliberately a functional and numerical sanity check, not a
privacy claim and not a clinical benchmark. It uses `sandbox_open` so that
per-node diagnostics can be observed during development. Production studies
should use the stricter DataSHIELD privacy profiles.

```{r}
summary_table <- data.frame(
  Field = c('generated_at', 'privacy_profile', 'n_sites', 'n_per_site', 'n_total', 'secagg_supported'),
  Value = c(validation$generated_at, validation$privacy_profile, validation$n_sites,
            validation$n_per_site, validation$n_total, validation$secagg_supported)
)
knitr::kable(summary_table)
```

```{r}
display <- results[, c(
  'method', 'task', 'centralized_metric', 'centralized_loss',
  'federated_status', 'federated_loss', 'delta_loss', 'validation_status'
)]
knitr::kable(display)
```

```{r, fig.width=8, fig.height=5}
plot_df <- results[is.finite(results$delta_loss), , drop = FALSE]
if (nrow(plot_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(plot_df, ggplot2::aes(x = stats::reorder(method, delta_loss), y = delta_loss, fill = task)) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = 'grey40') +
    ggplot2::geom_col(width = 0.7) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = 'Federated loss - centralized loss') +
    ggplot2::theme_minimal(base_size = 11)
}
```

The image-classification validation with images sourced from dsImaging is
documented separately in `direct-image-resnet`. XGBoost is included here
as a security guardrail check: centralized XGBoost trains locally, but the
federated template is blocked unless server-side Secure Aggregation is
available.

To reproduce the complete suite against configured Opal servers:

```sh
Rscript inst/demos/validate_methods.R
```
