---
title: "Vision Validation: pytorch_densenet121"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vision Validation: pytorch_densenet121}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "pytorch_densenet121"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records the catalogue coverage check for a dsFlower vision
template. Synthetic PNG images were split across three Opal/Rock nodes,
trained federatively through DataSHIELD, and compared with a centralized
PyTorch baseline trained on the pooled synthetic cohort. The purpose is
method-path validation: image staging, template execution, result
persistence, and cleanup. The current biomedical direct-image benchmark is
the PathMNIST vignette, and the dsImaging handoff evidence is reported in
the LUNG1 vignettes.

It is laid out in two parts: first the commands you would run to reproduce
the validation, then the recorded results of that run loaded from the
committed evidence artifact. No live Opal servers are required to render
this page; the command chunks are shown for reference and are not executed.

## Running This Validation

The federated run trains this vision template through dsFlower and compares
it with a centralized PyTorch baseline. Each node holds an image metadata
table (`id`, `relative_path`, `mask_path`, `label`) while the PNG files stay
on the matching Rock server; only aggregate training history and model
parameters return to the researcher.

**1. Connect to the DataSHIELD nodes.** Open a single session across the
three Opal nodes that serve the image metadata table.

```{r connect, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)

builder <- DSI::newDSLoginBuilder()
builder$append(server = 'node1', url = 'https://opal-node-1.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.vision_cohort')
builder$append(server = 'node2', url = 'https://opal-node-2.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.vision_cohort')
builder$append(server = 'node3', url = 'https://opal-node-3.example.org',
               user = 'researcher', password = '********',
               table = 'dsflower_demo.vision_cohort')
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = 'D')
```

**2. Run the federated fit.** Build the model and recipe, stage the nodes,
run FedAvg over the server-side image tables, then clean up and log out.
Only aggregate training history and model parameters are returned to the
researcher.

```{r federated-fit, eval=FALSE}
model <- ds.flower.model.pytorch_densenet121(
  n_classes = 2L, learning_rate = 0.001,
  batch_size = 2L, local_epochs = 1L, image_size = 64L
)

recipe <- ds.flower.recipe(
  model      = model,
  strategy   = ds.flower.strategy.fedavg(),
  privacy    = ds.flower.privacy.sandbox_open(),
  target     = 'label',
  num_rounds = 1
)

ds.flower.nodes.init(conns, data = 'D', symbol = 'flower')
ds.flower.nodes.prepare(
  conns, symbol = 'flower',
  target_column = 'label',
  run_config = c(recipe$model$params, recipe$strategy$params,
                 list(data_type = 'image')),
  privacy = recipe$privacy, template_name = recipe$model$template
)
ds.flower.nodes.ensure(conns, symbol = 'flower', template_name = recipe$model$template)
fit <- ds.flower.run.start(recipe, conns = conns, verbose = TRUE)
ds.flower.nodes.cleanup(conns, symbol = 'flower')

DSI::datashield.logout(conns)
```

## Recorded Results

The values below come from the committed evidence artifact for the run
described above; no live servers are contacted when the article renders.

This run trained the `r row$method` template on the `r row$task` task
(target `r row$target`) for `r row$rounds` federated round(s) under the
`r validation$privacy_profile` profile, across `r validation$n_sites`
Opal/Rock nodes that staged `r validation$n_total` synthetic
`r validation$image_size`x`r validation$image_size` PNG samples.

The centralized baseline reached a `r row$centralized_metric` of
`r fmt(row$centralized_loss)`; the federated run reached
`r fmt(row$federated_loss)` (delta `r fmt(row$delta_loss)`) with
`r row$federated_n_failures` client failures, leaving this template
**`r row$validation_status`**. The comparison is shown below, followed by
the centralized vs. federated loss.

```{r comparison-output}
comparison <- data.frame(
  centralized_loss = row$centralized_loss,
  centralized_accuracy = row$centralized_accuracy,
  federated_loss = row$federated_loss,
  delta_loss = row$delta_loss,
  federated_n_failures = row$federated_n_failures,
  validation_status = row$validation_status
)
knitr::kable(comparison, digits = 4)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for pytorch_densenet121."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```
