---
title: "Validation: pytorch_poisson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation: pytorch_poisson}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "pytorch_poisson"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records a live validation of the method against a centralized baseline.
The federated run used three Opal/Rock nodes and the same pooled synthetic
validation cohort was used for the local baseline. The purpose is method-path
validation: data staging, template verification, Flower execution, result
persistence, and cleanup. It is not a clinical performance benchmark.

```{r}
overview <- data.frame(
  Field = c(
    'method', 'task', 'target', 'n_features', 'rounds',
    'centralized_metric', 'centralized_loss', 'centralized_accuracy',
    'federated_status', 'federated_loss', 'federated_n_failures',
    'delta_loss', 'acceptance_max_loss_ratio',
    'acceptance_max_loss_margin', 'acceptable_loss',
    'validation_status'
  ),
  Value = c(
    row$method, row$task, row$target, row$n_features, row$rounds,
    row$centralized_metric, fmt(row$centralized_loss), fmt(row$centralized_accuracy),
    row$federated_status, fmt(row$federated_loss), fmt(row$federated_n_failures),
    fmt(row$delta_loss), fmt(row$acceptance_max_loss_ratio),
    fmt(row$acceptance_max_loss_margin), row$acceptable_loss,
    row$validation_status
  )
)
knitr::kable(overview)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for pytorch_poisson."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```

## Inline Execution Path

The validation row above was generated by running this method through the
same inline DataSHIELD/Flower flow shown below. The vignette does not
delegate execution to an external launcher; the rendered result is loaded
from the versioned JSON artifact written by the live run.

```{r opal-config, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)
library(opalr)

opal_urls <- trimws(strsplit(Sys.getenv(
  'DSFLOWER_OPAL_URLS',
  'https://localhost:8443,https://localhost:8444,https://localhost:8445'
), ',', fixed = TRUE)[[1]])
opal_users <- rep(Sys.getenv('OPAL_USER', 'administrator'), length(opal_urls))
opal_passwords <- rep(Sys.getenv('OPAL_PASSWORD', 'admin123'), length(opal_urls))
project <- Sys.getenv('DSFLOWER_DEMO_PROJECT', 'dsflower_demo')
```

```{r synthetic-cohort, eval=FALSE}
set.seed(4242)
n_sites <- length(opal_urls)
n_per_site <- validation$n_per_site
n <- n_sites * n_per_site
p <- row$n_features
x <- scale(matrix(stats::rnorm(n * p), nrow = n, ncol = p))
colnames(x) <- paste0('x', seq_len(p))

cohort <- data.frame(id = sprintf('mv_%04d', seq_len(n)), x)
cohort$outcome <- as.integer(0.9 * x[, 1] - 0.7 * x[, 2] +
                             0.45 * x[, 3] + stats::rnorm(n, sd = 0.4) > 0)
cohort$y <- 1.4 * x[, 1] - 1.1 * x[, 4] + 0.5 * x[, 7] +
            stats::rnorm(n, sd = 0.45)
cohort$class3 <- as.integer(cut(x[, 2] - 0.5 * x[, 5] + 0.35 * x[, 8],
                               breaks = stats::quantile(
                                 x[, 2] - 0.5 * x[, 5] + 0.35 * x[, 8],
                                 probs = c(0, 1 / 3, 2 / 3, 1)
                               ), include.lowest = TRUE, labels = FALSE)) - 1L
cohort$label_a <- as.integer(x[, 1] + x[, 6] + stats::rnorm(n, sd = 0.5) > 0)
cohort$label_b <- as.integer(-x[, 2] + x[, 7] + stats::rnorm(n, sd = 0.5) > 0)
cohort$label_c <- as.integer(x[, 3] - x[, 8] + stats::rnorm(n, sd = 0.5) > 0)
cohort$count <- stats::rpois(n, lambda = exp(pmin(pmax(0.2 + 0.35 * x[, 1] -
                                                    0.25 * x[, 9], -2), 2)))
cohort$time <- exp(2.0 - 0.25 * x[, 1] + 0.18 * x[, 5] +
                   stats::rnorm(n, sd = 0.35))
cohort$event <- as.integer(stats::runif(n) > 0.35)
cohort$event_type <- ifelse(cohort$event == 0L, 0L,
                            ifelse(0.6 * x[, 10] - 0.45 * x[, 11] > 0, 1L, 2L))
cohort$site <- rep(seq_len(n_sites), length.out = n)
```

```{r upload-and-connect, eval=FALSE}
table_paths <- character(n_sites)
for (i in seq_len(n_sites)) {
  opal <- opalr::opal.login(
    username = opal_users[[i]], password = opal_passwords[[i]],
    url = opal_urls[[i]],
    opts = list(ssl_verifyhost = 0L, ssl_verifypeer = 0L)
  )
  table_name <- paste0('method_validation_', method_id, '_site', i)
  site_data <- cohort[cohort$site == i, setdiff(names(cohort), 'site'), drop = FALSE]
  opalr::opal.table_save(opal, site_data, project = project, table = table_name,
                         id.name = 'id', policy = 'generate',
                         overwrite = TRUE, force = TRUE)
  table_paths[[i]] <- paste(project, table_name, sep = '.')
  opalr::opal.logout(opal)
}

builder <- DSI::newDSLoginBuilder()
for (i in seq_len(n_sites)) {
  builder$append(server = paste0('opal', i), url = opal_urls[[i]],
                 user = opal_users[[i]], password = opal_passwords[[i]],
                 table = table_paths[[i]], driver = 'OpalDriver')
}
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = 'D')
```

```{r federated-fit, eval=FALSE}
targets <- trimws(strsplit(row$target, ',', fixed = TRUE)[[1]])
features <- paste0('x', seq_len(row$n_features))

fit <- ds.flower.fit(
  conns,
  symbol = 'D',
  target = targets,
  features = features,
  model = row$method,
  strategy = 'fedavg',
  privacy = 'sandbox_open',
  rounds = row$rounds,
  task = row$task
)

post_caps <- DSI::datashield.aggregate(
  conns, as.symbol('flowerGetCapabilitiesDS')
)
DSI::datashield.logout(conns)
```

```{r acceptance-check, eval=FALSE}
history <- fit$history
federated_loss <- tail(stats::na.omit(as.numeric(history$loss)), 1)
federated_failures <- max(as.numeric(history$n_failures), na.rm = TRUE)
acceptable <- federated_failures == 0 &&
  federated_loss <= row$centralized_loss * row$acceptance_max_loss_ratio +
                    row$acceptance_max_loss_margin
acceptable
```

```{r}
if (nzchar(row$error)) cat(row$error)
```
