---
title: "dsFlower Validation Evidence Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dsFlower Validation Evidence Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_validation_results.json'),
  system.file('extdata', 'dsflower_method_validation_results.json', package = 'dsFlowerClient')
)
method_result_path <- method_paths[nzchar(method_paths) & file.exists(method_paths)][1]
method_validation <- jsonlite::fromJSON(method_result_path)
method_results <- method_validation$results
vision_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
vision_result_path <- vision_paths[nzchar(vision_paths) & file.exists(vision_paths)][1]
vision_validation <- if (!is.na(vision_result_path)) jsonlite::fromJSON(vision_result_path) else NULL
vision_results <- if (!is.null(vision_validation)) vision_validation$results else NULL
family_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_family_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_family_results.json'),
  system.file('extdata', 'dsflower_method_family_results.json', package = 'dsFlowerClient')
)
family_result_path <- family_paths[nzchar(family_paths) & file.exists(family_paths)][1]
family_validation <- if (!is.na(family_result_path)) jsonlite::fromJSON(family_result_path) else NULL
family_results <- if (!is.null(family_validation)) family_validation$results else NULL
family_dp_paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_method_family_dp_results.json'),
  file.path('inst', 'extdata', 'dsflower_method_family_dp_results.json'),
  system.file('extdata', 'dsflower_method_family_dp_results.json', package = 'dsFlowerClient')
)
family_dp_result_path <- family_dp_paths[nzchar(family_dp_paths) & file.exists(family_dp_paths)][1]
family_dp_validation <- if (!is.na(family_dp_result_path)) jsonlite::fromJSON(family_dp_result_path) else NULL
family_dp_results <- if (!is.null(family_dp_validation)) family_dp_validation$results else NULL
find_extdata <- function(file) {
  paths <- c(
    file.path('..', 'inst', 'extdata', file),
    file.path('inst', 'extdata', file),
    system.file('extdata', file, package = 'dsFlowerClient')
  )
  paths[nzchar(paths) & file.exists(paths)][1]
}
read_extdata <- function(file) {
  path <- find_extdata(file)
  if (is.na(path)) return(NULL)
  jsonlite::fromJSON(path, simplifyVector = FALSE)
}
count_results <- function(x) {
  if (is.null(x) || is.null(x$results)) return(NA_integer_)
  length(x$results)
}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0 || (length(x) == 1 && is.na(x))) y else x
}
count_failures <- function(x) {
  if (is.null(x) || is.null(x$results)) return(NA_integer_)
  sum(vapply(x$results, function(r) {
    sum(vapply(r$history %||% list(), function(h) h$n_failures %||% 0L,
               integer(1)))
  }, integer(1)))
}
clinical_trusted <- read_extdata('dsflower_clinical_algorithm_results.json')
clinical_secagg <- read_extdata('dsflower_clinical_secagg_results.json')
clinical_dp <- read_extdata('dsflower_clinical_dp_results.json')
xgb_noise <- read_extdata('dsflower_xgboost_histogram_privacy_results.json')
lung1_rad <- read_extdata('dsflower_lung1_radiomics_results.json')
lung1_img <- read_extdata('dsflower_lung1_direct_image_results.json')
pathmnist <- read_extdata('dsflower_pathmnist_direct_image_results.json')
column_or <- function(frame, name, default) {
  if (name %in% names(frame)) frame[[name]] else rep(default, nrow(frame))
}
method_display <- data.frame(
  suite = 'tabular_sequence_survival',
  method = method_results$method,
  task = method_results$task,
  centralized_metric = method_results$centralized_metric,
  centralized_loss = method_results$centralized_loss,
  federated_status = method_results$federated_status,
  federated_loss = method_results$federated_loss,
  delta_loss = method_results$delta_loss,
  acceptable_loss = method_results$acceptable_loss,
  validation_status = method_results$validation_status,
  stringsAsFactors = FALSE
)
vision_display <- if (!is.null(vision_results)) data.frame(
  suite = 'vision',
  method = vision_results$method,
  task = vision_results$task,
  centralized_metric = vision_results$centralized_metric,
  centralized_loss = vision_results$centralized_loss,
  federated_status = vision_results$federated_status,
  federated_loss = vision_results$federated_loss,
  delta_loss = vision_results$delta_loss,
  acceptable_loss = column_or(vision_results, 'acceptable_loss',
                              vision_results$validation_status == 'pass'),
  validation_status = vision_results$validation_status,
  stringsAsFactors = FALSE
) else method_display[0, ]
family_display <- if (!is.null(family_results)) data.frame(
  suite = 'clinical_support2_secagg',
  method = family_results$method,
  task = family_results$task,
  centralized_metric = family_results$centralized_metric,
  centralized_loss = family_results$centralized_loss,
  federated_status = family_results$federated_status,
  federated_loss = family_results$federated_loss,
  delta_loss = family_results$delta_loss,
  acceptable_loss = family_results$acceptable_loss,
  validation_status = family_results$validation_status,
  stringsAsFactors = FALSE
) else method_display[0, ]
family_dp_display <- if (!is.null(family_dp_results)) data.frame(
  suite = 'clinical_support2_dp_sgd',
  method = family_dp_results$method,
  task = family_dp_results$task,
  centralized_metric = family_dp_results$centralized_metric,
  centralized_loss = family_dp_results$centralized_loss,
  federated_status = family_dp_results$federated_status,
  federated_loss = family_dp_results$federated_loss,
  delta_loss = family_dp_results$delta_loss,
  acceptable_loss = family_dp_results$acceptable_loss,
  validation_status = family_dp_results$validation_status,
  stringsAsFactors = FALSE
) else method_display[0, ]
results <- rbind(method_display, vision_display, family_display, family_dp_display)
```

This macro-vignette summarizes the committed dsFlower validation evidence. The
current presentation layer is the public biomedical evidence: clinical
benchmarks, SUPPORT2 method-family tasks, LUNG1 imaging handoffs and the
PathMNIST direct-image benchmark. The synthetic method and vision suites provide
historical catalogue coverage: they record which templates in that retired
architecture could be staged, executed, aggregated, persisted and cleaned up
across three Opal/Rock servers. They are not a current server allowlist.
They complement the biomedical reports by checking the full template surface
under deterministic conditions.

The clinical and method-family benchmarks use `clinical_default`,
`clinical_update_noise` or `high_sensitivity_dp`, so Secure Aggregation,
profile-level disclosure guards and, where applicable, DP-SGD epsilon curves or
bounded histogram-noise curves are active. The SUPPORT2 DP-SGD evidence is
interpreted over the four templates whose losses decompose by sample. CoxPH is
still validated in the Secure Aggregation run; because its risk-set partial
likelihood is not a per-example loss, `high_sensitivity_dp` is not the matching
privacy profile for that route.

```{r current-evidence}
thesis_evidence <- data.frame(
  layer = c(
    'clinical trusted benchmark',
    'clinical SecAgg benchmark',
    'clinical DP-SGD benchmark',
    'XGBoost histogram update-noise',
    'SUPPORT2 SecAgg families',
    'SUPPORT2 DP-SGD families',
    'LUNG1 radiomics handoff',
    'LUNG1 direct-image handoff',
    'PathMNIST direct-image benchmark'
  ),
  data = c(
    'Breast Cancer, Heart Disease, Pima, CDC Diabetes',
    'Breast Cancer, Pima, CDC Diabetes',
    'CDC Diabetes',
    'CDC Diabetes',
    'SUPPORT2',
    'SUPPORT2',
    paste0(lung1_rad$dimensions[['dimensions of rad in combined studies']][[1]], ' LUNG1-derived rows'),
    paste0(sum(as.integer(unlist(lung1_img$metadata_n))), ' LUNG1 NIfTI assets'),
    paste0(pathmnist$n_total, ' PathMNIST images')
  ),
  profile = c(
    clinical_trusted$privacy_profile,
    clinical_secagg$privacy_profile,
    clinical_dp$privacy_profile,
    xgb_noise$privacy_profile,
    if (!is.null(family_validation)) family_validation$privacy_profile else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$privacy_profile else NA,
    lung1_rad$privacy,
    lung1_img$privacy,
    'trusted_internal + consortium_internal'
  ),
  runs_or_routes = c(
    count_results(clinical_trusted),
    count_results(clinical_secagg),
    count_results(clinical_dp),
    count_results(xgb_noise),
    if (!is.null(family_results)) nrow(family_results) else NA,
    if (!is.null(family_dp_results)) nrow(family_dp_results) else NA,
    length(lung1_rad$history),
    length(lung1_img$history),
    length(pathmnist$profile_results)
  ),
  failures = c(
    count_failures(clinical_trusted),
    count_failures(clinical_secagg),
    count_failures(clinical_dp),
    count_failures(xgb_noise),
    if (!is.null(family_results)) sum(family_results$federated_n_failures) else NA,
    if (!is.null(family_dp_results)) sum(family_dp_results$federated_n_failures) else NA,
    sum(vapply(lung1_rad$history, `[[`, numeric(1), 'n_failures')),
    lung1_img$local_vs_federated$federated$n_failures,
    pathmnist$trusted_internal_comparison$federated_failures +
      pathmnist$consortium_internal_secagg$federated_failures
  ),
  stringsAsFactors = FALSE
)
knitr::kable(thesis_evidence, row.names = FALSE)
```

```{r}
summary_table <- data.frame(
  suite = c('tabular_sequence_survival', 'vision', 'clinical_support2_secagg',
            'clinical_support2_dp_sgd'),
  generated_at = c(
    method_validation$generated_at,
    if (!is.null(vision_validation)) vision_validation$generated_at else NA,
    if (!is.null(family_validation)) family_validation$generated_at else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$generated_at else NA
  ),
  privacy_profile = c(
    method_validation$privacy_profile,
    if (!is.null(vision_validation)) vision_validation$privacy_profile else NA,
    if (!is.null(family_validation)) family_validation$privacy_profile else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$privacy_profile else NA
  ),
  n_sites = c(
    method_validation$n_sites,
    if (!is.null(vision_validation)) vision_validation$n_sites else NA,
    if (!is.null(family_validation)) family_validation$n_sites else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$n_sites else NA
  ),
  n_total = c(
    method_validation$n_total,
    if (!is.null(vision_validation)) vision_validation$n_total else NA,
    if (!is.null(family_validation)) family_validation$n_total else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$n_total else NA
  ),
  secagg_supported = c(
    method_validation$secagg_supported,
    NA,
    if (!is.null(family_validation)) family_validation$secagg_supported else NA,
    if (!is.null(family_dp_validation)) family_dp_validation$secagg_supported else NA
  ),
  stringsAsFactors = FALSE
)
knitr::kable(summary_table)
```

```{r}
knitr::kable(results)
```

```{r, fig.width=8, fig.height=5, fig.alt="Horizontal bar chart of federated minus centralized loss by method."}
plot_df <- results[is.finite(results$delta_loss), , drop = FALSE]
if (nrow(plot_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  plot_df$label <- paste(plot_df$suite, plot_df$method, sep = ': ')
  ggplot2::ggplot(plot_df, ggplot2::aes(x = stats::reorder(label, delta_loss), y = delta_loss, fill = task)) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = 'grey40') +
    ggplot2::geom_col(width = 0.7) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = 'Federated loss - centralized loss') +
    ggplot2::theme_minimal(base_size = 11)
}
```

Per-method historical vignettes (evidence only; not a current API index):

- [clinical method-family benchmarks](clinical-method-family-benchmarks.html)

- [sklearn_logreg](validation-sklearn-logreg.html)
- [sklearn_ridge](validation-sklearn-ridge.html)
- [sklearn_sgd](validation-sklearn-sgd.html)
- [sklearn_svm](validation-sklearn-svm.html)
- [sklearn_elastic_net](validation-sklearn-elastic-net.html)
- [pytorch_logreg](validation-pytorch-logreg.html)
- [pytorch_mlp](validation-pytorch-mlp.html)
- [pytorch_linear_regression](validation-pytorch-linear-regression.html)
- [pytorch_multiclass](validation-pytorch-multiclass.html)
- [pytorch_multilabel](validation-pytorch-multilabel.html)
- [pytorch_poisson](validation-pytorch-poisson.html)
- [pytorch_lstm](validation-pytorch-lstm.html)
- [pytorch_tcn](validation-pytorch-tcn.html)
- [xgboost](validation-xgboost.html)

Historical, unsupported in the current enforced-DP runtime:

- [pytorch_coxph](validation-pytorch-coxph.html)
- [pytorch_lognormal_aft](validation-pytorch-lognormal-aft.html)
- [pytorch_cause_specific_cox](validation-pytorch-cause-specific-cox.html)
- [pytorch_unet2d](validation-vision-pytorch-unet2d.html)

These four records do not correspond to registered models or runnable public
constructors. Cox risk-set losses lack the per-example decomposition required by
the current DP-SGD track; survival and dense segmentation are rejected early.

Vision method historical vignettes:

- [pytorch_resnet18](validation-vision-pytorch-resnet18.html)
- [pytorch_densenet121](validation-vision-pytorch-densenet121.html)

The vision-specific overview remains available at
[Vision Method Validation Overview](validation-vision-overview.html).

The full suite is produced by repeating the inline Opal/DataSHIELD/Flower
workflow shown in each per-method vignette for every method listed above.
The persisted JSON artifacts are the audit trail consumed by this
pkgdown site.
