---
title: "Vision Validation: pytorch_resnet18"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vision Validation: pytorch_resnet18}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "pytorch_resnet18"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records a live validation of a dsFlower vision template.
Synthetic PNG images were split across three Opal/Rock nodes, trained
federatively through DataSHIELD, and compared with a centralized PyTorch
baseline trained on the pooled synthetic cohort. This is a path validation
for image staging, template execution, result persistence, and cleanup;
it is not a clinical imaging benchmark.

```{r}
overview <- data.frame(
  Field = c(
    'method', 'task', 'target', 'n_total', 'image_size', 'rounds',
    'centralized_metric', 'centralized_loss', 'centralized_accuracy',
    'centralized_dice', 'federated_status', 'federated_loss',
    'federated_n_failures', 'delta_loss',
    'acceptance_max_loss_ratio', 'acceptance_max_loss_margin',
    'acceptable_loss', 'validation_status'
  ),
  Value = c(
    row$method, row$task, row$target, row$n_total, validation$image_size, row$rounds,
    row$centralized_metric, fmt(row$centralized_loss), fmt(row$centralized_accuracy),
    fmt(row$centralized_dice), row$federated_status, fmt(row$federated_loss),
    fmt(row$federated_n_failures), fmt(row$delta_loss),
    fmt(row$acceptance_max_loss_ratio), fmt(row$acceptance_max_loss_margin),
    row$acceptable_loss, row$validation_status
  )
)
knitr::kable(overview)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for pytorch_resnet18."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```

## Step-By-Step Run Output

The validation artifact stores the output of the live centralized and
federated runs. The chunks below render that output explicitly so the
reader can inspect the full path instead of seeing only a summary table.

```{r input-trace}
input_trace <- data.frame(
  step = c('generate_images', 'split_sites', 'upload_metadata', 'federated_rounds'),
  output = c(
    paste(row$n_total, 'synthetic PNG samples at', validation$image_size, 'x', validation$image_size),
    paste(row$n_sites, 'Opal/Rock nodes with', row$n_per_site, 'rows each'),
    'metadata tables only: id, relative_path, mask_path, label',
    paste(row$rounds, 'round(s) through DataSHIELD + Flower')
  )
)
knitr::kable(input_trace)
```

```{r centralized-run-output}
centralized_output <- data.frame(
  metric = c('status', row$centralized_metric, 'accuracy', 'dice', 'samples'),
  value = c(
    row$centralized_status,
    fmt(row$centralized_loss),
    fmt(row$centralized_accuracy),
    fmt(row$centralized_dice),
    row$n_total
  )
)
knitr::kable(centralized_output)
cat('[centralized] engine: local PyTorch baseline\n')
cat('[centralized] method:', row$method, '\n')
cat('[centralized] metric:', row$centralized_metric, '\n')
cat('[centralized] loss:', fmt(row$centralized_loss), '\n')
cat('[centralized] accuracy:', fmt(row$centralized_accuracy), '\n')
cat('[centralized] dice:', fmt(row$centralized_dice), '\n')
```

```{r federated-run-output}
dsflower_output <- data.frame(
  metric = c('status', 'loss', 'client_failures', 'accepted', 'validation_status'),
  value = c(
    row$federated_status,
    fmt(row$federated_loss),
    fmt(row$federated_n_failures),
    row$acceptable_loss,
    row$validation_status
  )
)
knitr::kable(dsflower_output)
cat('[DataSHIELD] connected nodes:', row$n_sites, '\n')
cat('[dsFlower] model template:', row$method, '\n')
cat('[Flower] final federated loss:', fmt(row$federated_loss), '\n')
cat('[Flower] client failures:', fmt(row$federated_n_failures), '\n')
cat('[acceptance] max loss ratio:', fmt(row$acceptance_max_loss_ratio), '\n')
cat('[acceptance] max loss margin:', fmt(row$acceptance_max_loss_margin), '\n')
cat('[acceptance] PASS:', row$validation_status == 'pass', '\n')
```

## Inline Execution Path

The validation row above was generated by running this vision template
through the inline workflow shown below. The vignette does not delegate to
an opaque benchmark wrapper; the rendered tables and console trace above
are loaded from the JSON artifact written by that live run.

```{r opal-config, eval=FALSE}
library(DSI)
library(DSOpal)
library(dsFlowerClient)
library(opalr)

opal_urls <- trimws(strsplit(Sys.getenv(
  'DSFLOWER_OPAL_URLS',
  'https://localhost:8443,https://localhost:8444,https://localhost:8445'
), ',', fixed = TRUE)[[1]])
opal_users <- rep(Sys.getenv('OPAL_USER', 'administrator'), length(opal_urls))
opal_passwords <- rep(Sys.getenv('OPAL_PASSWORD', 'admin123'), length(opal_urls))
project <- Sys.getenv('DSFLOWER_DEMO_PROJECT', 'dsflower_demo')
image_root <- Sys.getenv('DSFLOWER_IMAGE_ROOT', '/tmp/dsflower_vision_method_validation')
```

```{r make-images, eval=FALSE}
make_png <- function(path, label, size = validation$image_size) {
  dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)
  grid <- seq(0, 1, length.out = size)
  img <- outer(grid, grid, function(x, y) if (label == 1L) x else y)
  grDevices::png(path, width = size, height = size)
  old_par <- graphics::par(mar = c(0, 0, 0, 0))
  graphics::plot.new()
  graphics::rasterImage(grDevices::as.raster(img), 0, 0, 1, 1)
  graphics::par(old_par)
  grDevices::dev.off()
}

site_tables <- vector('list', length(opal_urls))
for (site in seq_along(opal_urls)) {
  rows <- vector('list', validation$n_per_site)
  for (i in seq_len(validation$n_per_site)) {
    label <- as.integer(i %% 2L)
    rel <- file.path(paste0('site', site), 'images', sprintf('img_%03d.png', i))
    mask <- file.path(paste0('site', site), 'masks', sprintf('mask_%03d.png', i))
    make_png(file.path(image_root, rel), label)
    make_png(file.path(image_root, mask), label)
    rows[[i]] <- data.frame(id = sprintf('vision_site%d_%03d', site, i),
                            relative_path = rel, mask_path = mask,
                            label = label)
  }
  site_tables[[site]] <- do.call(rbind, rows)
}
```

```{r upload-metadata, eval=FALSE}
table_paths <- character(length(opal_urls))
for (i in seq_along(opal_urls)) {
  opal <- opalr::opal.login(
    username = opal_users[[i]], password = opal_passwords[[i]],
    url = opal_urls[[i]],
    opts = list(ssl_verifyhost = 0L, ssl_verifypeer = 0L)
  )
  opalr::dsadmin.set_option(opal, 'dsflower.image_data_root',
                            image_root, profile = 'default')
  table_name <- paste0('vision_validation_', method_id, '_site', i)
  opalr::opal.table_save(opal, site_tables[[i]], project = project,
                         table = table_name, id.name = 'id',
                         policy = 'generate', overwrite = TRUE, force = TRUE)
  table_paths[[i]] <- paste(project, table_name, sep = '.')
  opalr::opal.logout(opal)
}

builder <- DSI::newDSLoginBuilder()
for (i in seq_along(opal_urls)) {
  builder$append(server = paste0('opal', i), url = opal_urls[[i]],
                 user = opal_users[[i]], password = opal_passwords[[i]],
                 table = table_paths[[i]], driver = 'OpalDriver')
}
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = 'D')
```

```{r federated-fit, eval=FALSE}
model <- switch(
  method_id,
  pytorch_resnet18 = ds.flower.model.pytorch_resnet18(
    n_classes = 2L, image_size = validation$image_size,
    batch_size = 2L, local_epochs = 1L
  ),
  pytorch_densenet121 = ds.flower.model.pytorch_densenet121(
    n_classes = 2L, image_size = validation$image_size,
    batch_size = 2L, local_epochs = 1L
  ),
  pytorch_unet2d = ds.flower.model.pytorch_unet2d(
    n_classes = 1L, image_size = validation$image_size,
    batch_size = 1L, local_epochs = 1L, base_channels = 8L
  )
)

recipe <- ds.flower.recipe(
  model = model,
  strategy = ds.flower.strategy.fedavg(),
  privacy = ds.flower.privacy.sandbox_open(),
  target = row$target,
  num_rounds = row$rounds
)

symbol <- paste0('flower_', method_id)
ds.flower.nodes.init(conns, data = 'D', symbol = symbol)
ds.flower.nodes.prepare(
  conns, symbol = symbol, target_column = row$target,
  run_config = c(model$params, list(data_type = 'image')),
  privacy = recipe$privacy, template_name = model$template
)
ds.flower.nodes.ensure(conns, symbol = symbol, template_name = model$template)
fit <- ds.flower.run.start(recipe, conns = conns, verbose = TRUE)
ds.flower.nodes.cleanup(conns, symbol = symbol)
DSI::datashield.logout(conns)
```

```{r acceptance-check, eval=FALSE}
history <- fit$history
federated_loss <- tail(stats::na.omit(as.numeric(history$loss)), 1)
federated_failures <- max(as.numeric(history$n_failures), na.rm = TRUE)
acceptable <- federated_failures == 0 &&
  federated_loss <= row$centralized_loss * row$acceptance_max_loss_ratio +
                    row$acceptance_max_loss_margin
acceptable
```

```{r}
if (nzchar(row$error)) cat(row$error)
```
