---
title: "Vision Validation: pytorch_densenet121"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vision Validation: pytorch_densenet121}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
method_id <- "pytorch_densenet121"
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
row <- results[results$method == method_id, , drop = FALSE]
fmt <- function(x) {
  if (length(x) == 0 || is.null(x) || is.na(x)) return('NA')
  if (is.numeric(x)) return(format(round(x, 4), nsmall = 4, trim = TRUE))
  as.character(x)
}
```

This vignette records a live validation of a dsFlower vision template.
Synthetic PNG images were split across three Opal/Rock nodes, trained
federatively through DataSHIELD, and compared with a centralized PyTorch
baseline trained on the pooled synthetic cohort. This is a path validation
for image staging, template execution, result persistence, and cleanup;
it is not a clinical imaging benchmark.

```{r}
overview <- data.frame(
  Field = c(
    'method', 'task', 'target', 'n_total', 'image_size', 'rounds',
    'centralized_metric', 'centralized_loss', 'centralized_accuracy',
    'centralized_dice', 'federated_status', 'federated_loss',
    'federated_n_failures', 'delta_loss', 'validation_status'
  ),
  Value = c(
    row$method, row$task, row$target, row$n_total, validation$image_size, row$rounds,
    row$centralized_metric, fmt(row$centralized_loss), fmt(row$centralized_accuracy),
    fmt(row$centralized_dice), row$federated_status, fmt(row$federated_loss),
    fmt(row$federated_n_failures), fmt(row$delta_loss), row$validation_status
  )
)
knitr::kable(overview)
```

```{r, fig.width=5.5, fig.height=3.2, fig.alt="Bar chart comparing centralized and federated validation loss for pytorch_densenet121."}
loss_df <- data.frame(
  mode = c('Centralized', 'Federated'),
  loss = c(row$centralized_loss, row$federated_loss)
)
loss_df <- loss_df[is.finite(loss_df$loss), , drop = FALSE]
if (nrow(loss_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(loss_df, ggplot2::aes(x = mode, y = loss, fill = mode)) +
    ggplot2::geom_col(width = 0.65, show.legend = FALSE) +
    ggplot2::labs(x = NULL, y = row$centralized_metric[[1]], title = method_id) +
    ggplot2::theme_minimal(base_size = 11)
} else {
  plot.new()
  text(0.5, 0.5, 'No federated loss available for this method')
}
```

To reproduce only this method against configured Opal servers:

```sh
DSFLOWER_VALIDATE_VISION_METHODS=pytorch_densenet121 Rscript inst/demos/validate_vision_methods.R
```

```{r}
if (nzchar(row$error)) cat(row$error)
```
