---
title: "Vision Method Validation Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vision Method Validation Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> **Historical validation artifact (pre-0.3).** The recorded results and API
> snippets below describe the retired client privacy-profile/Secure
> Aggregation architecture. They do not validate the current v0.3 lifetime-DP
> contract and must not be used as deployment or security instructions. See
> `README.md` and the server `ARCHITECTURE.md` for the enforced contract.

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
paths <- c(
  file.path('..', 'inst', 'extdata', 'dsflower_vision_validation_results.json'),
  file.path('inst', 'extdata', 'dsflower_vision_validation_results.json'),
  system.file('extdata', 'dsflower_vision_validation_results.json', package = 'dsFlowerClient')
)
result_path <- paths[nzchar(paths) & file.exists(paths)][1]
validation <- jsonlite::fromJSON(result_path)
results <- validation$results
```

This macro-vignette summarizes the catalogue coverage check for the
dsFlower vision templates. The experiment uses synthetic image and mask
files staged on the Rock servers, while metadata remains in Opal tables.
Each federated run is compared with a centralized PyTorch baseline trained
over the pooled files. The current biomedical direct-image evidence is
reported separately in the PathMNIST and LUNG1 dsImaging vignettes.

This suite was generated on `r validation$generated_at` under the
`r validation$privacy_profile` profile, across `r validation$n_sites`
Opal/Rock nodes holding `r validation$n_per_site` samples each
(`r validation$n_total` total) at
`r validation$image_size`x`r validation$image_size` resolution.

```{r}
display <- results[, c(
  'method', 'task', 'centralized_metric', 'centralized_loss',
  'federated_status', 'federated_loss', 'delta_loss', 'validation_status'
)]
knitr::kable(display, digits = 4)
```

```{r, fig.width=7, fig.height=3.5, fig.alt="Horizontal bar chart of vision-template federated minus centralized loss."}
plot_df <- results[is.finite(results$delta_loss), , drop = FALSE]
if (nrow(plot_df) > 0 && requireNamespace('ggplot2', quietly = TRUE)) {
  ggplot2::ggplot(plot_df, ggplot2::aes(x = stats::reorder(method, delta_loss), y = delta_loss, fill = task)) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.3, color = 'grey40') +
    ggplot2::geom_col(width = 0.7) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = NULL, y = 'Federated loss - centralized loss') +
    ggplot2::theme_minimal(base_size = 11)
}
```

Each vision template has its own per-method vignette with the full
DataSHIELD/Flower commands and recorded results. The committed evidence
artifact is the audit trail consumed by this pkgdown page.
