---
title: "Federated Machine Learning across 3 hospitals"
subtitle: "Training a model on data that never leaves the building — from your laptop, over Tor"
author: "dsFlower"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## The problem

Three hospitals each hold part of a breast-cancer dataset. Each tumour is
described by 30 imaging measurements and labelled **malignant** or **benign**.
Together the data would train an excellent diagnostic model — but **the data
cannot leave each hospital** for privacy and governance reasons.

**Federated learning** solves this: the *model* travels to the data, trains
locally at each site, and only model updates (never patient rows) come back.
`dsFlower` runs the [Flower](https://flower.ai) federated-learning engine inside
[DataSHIELD](https://datashield.org)/Opal, so the whole thing is driven from a
single R session on your laptop.

```
   your laptop (this R session)
        │  orchestrates rounds, never sees a patient row
        ▼
  ┌───────────┬───────────┬───────────┐
  │ Nairobi   │  Dakar    │  Douala   │   3 Opal hospitals
  │ 190 pts   │  190 pts  │  189 pts  │   data stays put
  └───────────┴───────────┴───────────┘
```

```{r libs}
library(DSI)
library(DSOpal)
library(dsFlowerClient)
```

## 1. Connect to the three hospitals

```{r login}
builder <- DSI::newDSLoginBuilder()
for (h in c("nairobi", "dakar", "douala")) {
  builder$append(
    server   = h,
    url      = paste0("https://", h, ".datashield.live"),
    user     = "administrator",
    password = "password",
    table    = "dsflower_demo.breast_cancer",
    driver   = "OpalDriver",
    options  = "list(ssl_verifyhost=0L, ssl_verifypeer=0L)"
  )
}
conns <- DSI::datashield.login(builder$build(), assign = TRUE, symbol = "D")
```

## 2. The data is distributed — and we never see it

Each site loaded its own slice into the symbol `D`. We can compute on it, but
individual rows stay on the servers.

```{r explore}
# How many patients per hospital?
DSI::datashield.aggregate(conns, quote(dimDS("D")))

# A federated mean of one feature — pooled across all 569 patients,
# computed without any row leaving a hospital:
dsBaseClient::ds.mean("D$mean_radius", type = "combine", datasources = conns)

# Class balance per site:
dsBaseClient::ds.table("D$malignant", datasources = conns)
```

> Notice: we asked questions of **569 patients across three countries** and got
> answers, yet not a single patient record was transferred to this laptop.

## 3. Open the federated-learning link (zero setup)

The Flower training engine needs the hospitals' compute nodes to talk to a
coordinator (SuperLink) on *this* laptop — but the laptop is behind NAT and the
hospitals' servers are locked down. `dsFlower` carries that entire conversation
**through the DataSHIELD channel itself**: a local SuperLink is started and each
node's SuperNode reaches it over the same authenticated DataSHIELD connection you
already logged in with. No public server, no account, no VPN key, no open ports —
nothing to configure.

```{r transport, eval=FALSE}
ds.flower.link.up(conns)   # local SuperLink; nodes connect over the DataSHIELD channel
```

## 4. Train a federated logistic regression

One call trains across all three hospitals. The model visits each site, trains
on local tumours, and only **aggregated weights** (protected by secure
aggregation) come back each round.

```{r train, eval=FALSE}
features <- c(
  "mean_radius", "mean_texture", "mean_perimeter", "mean_area",
  "mean_smoothness", "mean_compactness", "mean_concavity",
  "mean_concave_points", "mean_symmetry", "mean_fractal_dimension",
  "radius_error", "texture_error", "perimeter_error", "area_error",
  "smoothness_error", "compactness_error", "concavity_error",
  "concave_points_error", "symmetry_error", "fractal_dimension_error",
  "worst_radius", "worst_texture", "worst_perimeter", "worst_area",
  "worst_smoothness", "worst_compactness", "worst_concavity",
  "worst_concave_points", "worst_symmetry", "worst_fractal_dimension"
)

fit <- ds.flower.fit(
  conns,
  symbol   = "D",
  target   = "malignant",        # 1 = malignant, 0 = benign
  features = features,
  model    = ds.flower.model.sklearn_logreg(C = 1.0, max_iter = 100L),
  strategy = ds.flower.strategy.fedavg(),
  privacy  = ds.flower.privacy.clinical_default(),  # Secure Aggregation
  rounds   = 5
)
# clinical_default uses Secure Aggregation: all 3 hospitals must stay online for
# the whole round (min 3 clients). If a site flaps, lower `rounds`.
```

### How the model improved each round

```{r metrics, eval=FALSE}
ds.flower.metrics(fit)
ds.flower.plot(fit)   # accuracy / AUC across federated rounds
```

## 5. Predict on a new patient

The trained model can now score a new tumour locally — no server needed.

```{r predict, eval=FALSE}
new_patient <- data.frame(
  mean_radius = 17.99, mean_texture = 10.38, mean_perimeter = 122.8,
  mean_area = 1001, mean_smoothness = 0.1184, mean_compactness = 0.2776,
  mean_concavity = 0.3001, mean_concave_points = 0.1471, mean_symmetry = 0.2419,
  mean_fractal_dimension = 0.0787,
  radius_error = 1.095, texture_error = 0.9053, perimeter_error = 8.589,
  area_error = 153.4, smoothness_error = 0.0064, compactness_error = 0.049,
  concavity_error = 0.0537, concave_points_error = 0.0159,
  symmetry_error = 0.03, fractal_dimension_error = 0.0062,
  worst_radius = 25.38, worst_texture = 17.33, worst_perimeter = 184.6,
  worst_area = 2019, worst_smoothness = 0.1622, worst_compactness = 0.6656,
  worst_concavity = 0.7119, worst_concave_points = 0.2654,
  worst_symmetry = 0.4601, worst_fractal_dimension = 0.1189
)

ds.flower.predict(fit, new_patient, type = "prob")      # P(malignant)
ds.flower.predict(fit, new_patient, type = "response")  # class label
```

## 6. Clean up

```{r cleanup, eval=FALSE}
ds.flower.link.down(conns)
DSI::datashield.logout(conns)
```

---

### What just happened

- A diagnostic model was trained on **569 patients across three hospitals**.
- **No patient record ever left a hospital** — only model weights crossed the
  wire, protected by secure aggregation.
- The whole Flower SuperNode↔SuperLink conversation travelled **inside the
  DataSHIELD channel**, so it worked from a laptop behind NAT with **no servers,
  accounts, keys or open ports to set up**.
